### 2020-11-10 scripts/readingdatatrial.Rmd
A start to the BC sandbox

---
title: "Sandbox BC Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r sandbox}
library(here)
library(tidyverse)
library("readxl")

BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCdatastripped <- BCdata %>% 
  select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT") %>% 
  filter(DATA_LEVEL == "PROVINCE LEVEL", 
         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")

# view(BCdatastripped)
 ggplot(BCdatastripped, aes(SCHOOL_YEAR, NUMBER_OF_STUDENTS)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
   facet_wrap(~SPECIAL_NEEDS_CODE)

```

### 2020-11-12 scripts/draft_play.Rmd
Just an APA style draft outline.

---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "University of Oregon"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
library("papaja")
r_refs("r-references.bib")
library(here)
library(tidyverse)
library("readxl")

BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCdatastripped <- BCdata %>% 
  select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT") %>% 
  filter(DATA_LEVEL == "PROVINCE LEVEL", 
         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-11-12 scripts/readingdatatrial.Rmd
Updated Sandbox for BC Data

@@
          "PUBLIC_OR_INDEPENDENT") %>% 
+  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
+                          "Physically Dependent" = "A",
+                          "Deafblind" = "B",
+                          "Moderate+ Intellectual Disability" = "C",
+                          "Chronic Health Impairment" = "D",
+                          "Visual Impairment" = "E",
+                          "Deaf or Hard of Hearing" = "F",
+                          "Autism" = "G",
+                          "Intensive Behavioural Intervention" = "H",
+                          "Mild Intellectual Disability" = "K",
+                          "Learning Disability" = "Q",
+                          "Gifted" = "P",
+                          "Mild Behavioural Intervention" = "R"),
+         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
+         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
   filter(DATA_LEVEL == "PROVINCE LEVEL", 
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
+         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SPECIAL_NEEDS_CODE != "M",
+         SPECIAL_NEEDS_CODE != "J",
+         SPECIAL_NEEDS_CODE != "N")
 
-# view(BCdatastripped)
- ggplot(BCdatastripped, aes(SCHOOL_YEAR, NUMBER_OF_STUDENTS)) +
-  geom_col() +
+ ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+  geom_line(color = "cornflowerblue") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +

### 2020-11-12 scripts/SchoolDistrictCategoriesSandbox_DF.Rmd
Sandbox for School Distric-level Data

--- 2020-11-10 scripts/readingdatatrial.Rmd
@@
          "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT") %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL", 
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER") %>% 
+  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
+                          "Physically Dependent" = "A",
+                          "Deafblind" = "B",
+                          "Moderate+ Intellectual Disability" = "C",
+                          "Chronic Health Impairment" = "D",
+                          "Visual Impairment" = "E",
+                          "Deaf or Hard of Hearing" = "F",
+                          "Autism" = "G",
+                          "Intensive Behavioural Intervention" = "H",
+                          "Mild Intellectual Disability" = "K",
+                          "Learning Disability" = "Q",
+                          "Gifted" = "P",
+                          "Mild Behavioural Intervention" = "R"),
+         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
+         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
+  filter(DATA_LEVEL == "DISTRICT LEVEL", 
+         DISTRICT_NUMBER == "070",
+         #PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SPECIAL_NEEDS_CODE != "M",
+         SPECIAL_NEEDS_CODE != "J",
+         SPECIAL_NEEDS_CODE != "N")
 
-# view(BCdatastripped)
- ggplot(BCdatastripped, aes(SCHOOL_YEAR, NUMBER_OF_STUDENTS)) +
-  geom_col() +
+ ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+  geom_line(color = "cornflowerblue") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +

### 2020-11-15 scripts/draft_play.Rmd
added urban v. rural data
still a little messy

@@
 
+```{r urban vs. rural}
+
+bc_data_district_classification <- read_excel("C:/Users/makay/Dropbox (University of Oregon)/University of Oregon/Fall 2020/EDLD 651 Data Science/Final Project/bc_data_district_classification.xlsx")
+
+district_class <- bc_data_district_classification %>%
+  select("SCHOOL_YEAR",
+         "DATA_LEVEL",
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER",
+         "DISTRICT_NAME",
+         "DISTRICT_POPULATION_2016",
+         "DISTRICT_CLASSIFICATION",
+         "SPECIAL_NEEDS_CODE",
+         "NUMBER_OF_STUDENTS") %>%
+  filter(DATA_LEVEL == "DISTRICT LEVEL",
+         PUBLIC_OR_INDEPENDENT == "BC Public School")
+
+```
+
 # Methods

### 2020-11-15 scripts/draft_play.Rmd
Added some text and words

@@
 
+The district classification data was scaled down to include only public schools, while excluding private institutions.
+The school districts of Souteast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gqaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga's, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information.
+The urban vs.
+rural classifications were made based on the district's population on the 2016 census.
+If the population was above 100,000 individuals, it is classified as urban.
+If the population was below 99,999 individuals, then it is classified as rural.
+
+``` {r District Census Population Table}
+
+```
+
+This table displays the census results from 2011 and 2016 for our school districts.
+Many of the populations have stayed consistent within their urban or rural category.
+Three districts to note from the table are Nanaimo, Kamloops, and Chilliwack.
+In 2011, they were rural, but for our dataset they have been classified as urban due to their population increase in 2016.
+
 # Methods

### 2020-11-16 scripts/draft_play.Rmd
Adding Oregon data and code

@@
 
+```{r}
+library(tidyverse)
+library(janitor)
+library(here)
+library("readxl")
+primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
+cleanest_primo <-primodata %>% 
+  remove_empty(which = c("rows", "cols")) %>% 
+  clean_names()
+
+```
+
+```{r}
+cleanest_primo <- cleanest_primo %>% 
+  select(-six_to_twentyone_total, -x24)
+
+```
+
+```{r}
+cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
+cleanest_primo2
+
+```
+
+```{r}
+fresh_primo <- cleanest_primo2 %>% 
+  pivot_longer(starts_with("x"), names_to = "age", values_to = "frequency") %>% 
+  mutate(age = parse_number(age), 
+         year = parse_number(year))
+
+```
+
+```{r}
+fresh_primo %>% 
+  group_by(year, disability) %>% 
+  summarize(total=sum(frequency, na.rm=TRUE))
+```
+
+```{r}
+ggplot(fresh_primo, aes(age, frequency))+
+  geom_line(aes(color=disability))+
+  facet_wrap(~year)
+
+```
+
+```{r}
+#hiding this for now
+#fresh_primo %>% 
+ # filter(disability=="AUTISM") %>%
+
+#ggplot(aes(age, frequency))+
+#  geom_line(aes(color=disability))+
+ # facet_wrap(~year)
+
+```
+```{r}
+#this is more in line with the BC graphing
+#ggplot(fresh_primo, aes(year, frequency))+
+# geom_smooth(color = "cornflowerblue") +
+# theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+#  facet_wrap(~disability)
+
+```
+
 The district classification data was scaled down to include only public schools, while excluding private institutions.

### 2020-11-16 scripts/draft_play.Rmd
Moved Joe's code around

--- 2020-11-15 scripts/draft_play.Rmd
@@
 library("readxl")
+library(janitor)
 
@@
   filter(DATA_LEVEL == "DISTRICT LEVEL",
-         PUBLIC_OR_INDEPENDENT == "BC Public School")
+         PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
+  remove_empty_rows()
 
@@
 In 2011, they were rural, but for our dataset they have been classified as urban due to their population increase in 2016.
+
+```{r cleaning Oregon data}
+primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
+cleanest_primo <-primodata %>% 
+  remove_empty(which = c("rows", "cols")) %>% 
+  clean_names()
+
+cleanest_primo <- cleanest_primo %>% 
+  select(-six_to_twentyone_total, -x24)
+
+cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
+cleanest_primo2
+
+fresh_primo <- cleanest_primo2 %>% 
+  pivot_longer(starts_with("x"), names_to = "age", values_to = "frequency") %>% 
+  mutate(age = parse_number(age), 
+         year = parse_number(year))
+
+fresh_primo %>% 
+  group_by(year, disability) %>% 
+  summarize(total=sum(frequency, na.rm=TRUE))
+
+```
+
+```{r Oregon plot}
+
+ggplot(fresh_primo, aes(age, frequency))+
+  geom_line(aes(color=disability))+
+  facet_wrap(~year)
+
+```
+
+```{r}
+#hiding this for now
+#fresh_primo %>% 
+ # filter(disability=="AUTISM") %>%
+
+#ggplot(aes(age, frequency))+
+#  geom_line(aes(color=disability))+
+ # facet_wrap(~year)
+
+```
+```{r}
+#this is more in line with the BC graphing
+#ggplot(fresh_primo, aes(year, frequency))+
+# geom_smooth(color = "cornflowerblue") +
+# theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+#  facet_wrap(~disability)
+
+```
 

### 2020-11-16 scripts/draft_play.Rmd
Showing the cornflower blue for you...

@@
 #this is more in line with the BC graphing
-#ggplot(fresh_primo, aes(year, frequency))+
-# geom_smooth(color = "cornflowerblue") +
-# theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-#  facet_wrap(~disability)
+ggplot(fresh_primo, aes(year, frequency))+
+ geom_smooth(color = "cornflowerblue") +
+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+ facet_wrap(~disability)
 

### 2020-11-16 scripts/draft_play.Rmd
District class pivot longer

@@
 
-bc_data_district_classification <- read_excel("C:/Users/makay/Dropbox (University of Oregon)/University of Oregon/Fall 2020/EDLD 651 Data Science/Final Project/bc_data_district_classification.xlsx")
+bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))
 
@@
          PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
-  remove_empty_rows()
+  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE)
 

### 2020-11-16 scripts/draft_play.Rmd
Fixed district names
woo-hoo

@@
 The district classification data was scaled down to include only public schools, while excluding private institutions.
-The school districts of Souteast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gqaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga's, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information.
+The school districts of Southeast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gwaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga'a, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information.
 The urban vs.

### 2020-11-16 scripts/draft_play.Rmd
Updated with first view of BC Data

@@
 
+-- more introductory text here --
+Children designated with Special Needs categories have predominantly increased at different rates in British Colubmia over time.
+The figure below demonstrates growth of 12 potential designations over an 18 year time period:
+
+```{r sandbox}
+library(here)
+library(tidyverse)
+library("readxl")
+
+BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
+
+BCdatastripped <- BCdata %>% 
+  select("SCHOOL_YEAR", 
+         "DATA_LEVEL", 
+         "SPECIAL_NEEDS_CODE", 
+         "NUMBER_OF_STUDENTS", 
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER") %>% 
+  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
+                          "Physically Dependent" = "A",
+                          "Deafblind" = "B",
+                          "Moderate+ Intellectual Disability" = "C",
+                          "Chronic Health Impairment" = "D",
+                          "Visual Impairment" = "E",
+                          "Deaf or Hard of Hearing" = "F",
+                          "Autism" = "G",
+                          "Intensive Behavioural Intervention" = "H",
+                          "Mild Intellectual Disability" = "K",
+                          "Learning Disability" = "Q",
+                          "Gifted" = "P",
+                          "Mild Behavioural Intervention" = "R"),
+         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
+         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
+  filter(DATA_LEVEL == "DISTRICT LEVEL", 
+         DISTRICT_NUMBER == "070",
+         #PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SPECIAL_NEEDS_CODE != "M",
+         SPECIAL_NEEDS_CODE != "J",
+         SPECIAL_NEEDS_CODE != "N")
+
+ ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+  geom_line(color = "cornflowerblue") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+   facet_wrap(~SPECIAL_NEEDS_CODE)
+
+```
+
 ```{r urban vs. rural}
@@
          PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
-  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE)
+  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
+    select(-"name")
 
@@
 #hiding this for now
-#fresh_primo %>% 
- # filter(disability=="AUTISM") %>%
+fresh_primo %>%
+filter(disability=="AUTISM") %>%
 
-#ggplot(aes(age, frequency))+
-#  geom_line(aes(color=disability))+
- # facet_wrap(~year)
+ggplot(aes(age, frequency))+
+ geom_line(aes(color=disability))+
+facet_wrap(~year)
 

### 2020-11-16 scripts/draft_play.Rmd
Fixed spelling of Columbia

@@
 -- more introductory text here --
-Children designated with Special Needs categories have predominantly increased at different rates in British Colubmia over time.
+Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
 The figure below demonstrates growth of 12 potential designations over an 18 year time period:

### 2020-11-16 scripts/draft_play.Rmd
Author change

@@
 author: 
-  - name          : "First Author"
+  - name          : "Makayla Whitney"
     affiliation   : "1"

### 2020-11-16 scripts/draft_play.Rmd
More author changes

--- 2020-11-16 scripts/draft_play.Rmd
@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "Do We Have A Title Yet?"
+shorttitle        : "not quite"
 
 author: 
-  - name          : "First Author"
+  - name          : "David Fainstein"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
+  - name          : "Janette Avelar"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
+  - name          : "Joe Swinehart"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Ernst-August Doelle"
+  - name          : "Makayla Whitney"
     affiliation   : "1"

### 2020-11-16 scripts/draft_play.Rmd
Start of Intro

--- 2020-11-16 scripts/draft_play.Rmd
@@
 ```
+
+# Introduction
+
+Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
+The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon.
+The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
+The OR head counts include years 2002-2020.
+Within the BC dataset, data acquisition is part of formal student status reporting that is required within BC public schools, typically occurring in October of each school year.
+Levels of categorization include school- district- and provincial-level head counts.
+The OR data set includes state-wide head counts that are not aggregated by school or district.
+Etc.
 

### 2020-11-17 scripts/draft_play.Rmd
Added description of data/intro

--- 2020-11-16 scripts/draft_play.Rmd
@@
 The OR head counts include years 2002-2020.
-Within the BC dataset, data acquisition is part of formal student status reporting that is required within BC public schools, typically occurring in October of each school year.
-Levels of categorization include school- district- and provincial-level head counts.
+Levels of categorization include school- district- and provincial-level head counts for BC.
 The OR data set includes state-wide head counts that are not aggregated by school or district.
-Etc.
+
+We intend to explore several questions regarding longitudinal trends.
+Firstly, we hope to analyze trends in disability prevalence over time.
+We will engage in a discussion on developmental trajectories by studying how trends shift from static/linear to increasing linear trends based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data.
+In studying the differences and similarities between the two datasets we will also engage in a discussion on diagnosis terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM), in part as a response to a challenge set by differing terminology between BC/OR categorization.
+
+Secondly, we hope to more closely analyze any changes, or lack thereof, within the BC data following the year 2016 during which a BC Supreme Court ruled in favor of limiting the number of special needs students in classrooms and expanding the number of specialist teachers schools are required to hire.
+
+Finally, we will also explore differences between urban and rural school districts in BC.
+Districts are categorized by population size and proximity to metropolitan areas, as recorded and defined by the Statistics Canada census from 2016.
+Two fringe rural districts exhibiting high populations will be picked out and  high-incidence diagnoses will be compared to those of other regions over time.
 

### 2020-11-17 scripts/draft_play.Rmd
Spruce up for BC ggplot

@@
                           "Deafblind" = "B",
-                          "Moderate+ Intellectual Disability" = "C",
+                          "Moderate+ ID" = "C",
                           "Chronic Health Impairment" = "D",
@@
                           "Autism" = "G",
-                          "Intensive Behavioural Intervention" = "H",
-                          "Mild Intellectual Disability" = "K",
+                          "Intensive Behavioural Need" = "H",
+                          "Mild ID" = "K",
                           "Learning Disability" = "Q",
                           "Gifted" = "P",
-                          "Mild Behavioural Intervention" = "R"),
+                          "Mild Behavioural Need" = "R"),
          SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
@@
          SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(DATA_LEVEL == "DISTRICT LEVEL", 
-         DISTRICT_NUMBER == "070",
-         #PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+  filter(DATA_LEVEL == "PROVINCE LEVEL",
+         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SCHOOL_YEAR >= 7,
          SPECIAL_NEEDS_CODE != "M",
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-   facet_wrap(~SPECIAL_NEEDS_CODE)
+   facet_wrap(~SPECIAL_NEEDS_CODE) +
+   labs(title = "BC Provincial Headcount of Students with Exceptionalities by Category",
+        x = "School Year",
+        y = "Headcount")
 

### 2020-11-18 scripts/draft_play.Rmd
Added some text for paper and cleaned up BC data code

--- 2020-11-17 scripts/draft_play.Rmd
@@
 
--- more introductory text here --
+# Problem Statement and RQs
+In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
+While Canadian K-12 academic achievement outcomes are viewed as favourable on the world stage, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education.
+Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur.
+The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.
+
+With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
+(1) Do student prevalence rates of disability or disorder change after the supreme court ruling of 2016?
+(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+
+# Results
+
 Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
@@
 
-```{r sandbox}
+```{r sandbox, warning=FALSE}
 library(here)
@@
 
-BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
-
-BCdatastripped <- BCdata %>% 
-  select("SCHOOL_YEAR", 
-         "DATA_LEVEL", 
-         "SPECIAL_NEEDS_CODE", 
-         "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT",
-         "DISTRICT_NUMBER") %>% 
+BCdatastripped_spedcat <- BCdatastripped %>% 
   mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
@@
                           "Autism" = "G",
-                          "Intensive Behavioural Intervention" = "H",
-                          "Mild Intellectual Disability" = "K",
+                          "Intensive Behavioural Need" = "H",
+                          "Mild ID" = "K",
                           "Learning Disability" = "Q",
                           "Gifted" = "P",
-                          "Mild Behavioural Intervention" = "R"),
+                          "Mild Behavioural Need" = "R"),
          SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
          NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
-         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(DATA_LEVEL == "DISTRICT LEVEL", 
-         DISTRICT_NUMBER == "070",
-         #PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SCHOOL_YEAR = as.integer(SCHOOL_YEAR)) %>% 
+  filter(SCHOOL_YEAR >= 7,
          SPECIAL_NEEDS_CODE != "M",
@@
 
- ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+ ggplot(BCdatastripped_spedcat, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
   geom_line(color = "cornflowerblue") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-   facet_wrap(~SPECIAL_NEEDS_CODE)
+   facet_wrap(~SPECIAL_NEEDS_CODE) +
+   labs(title = "BC Provincial Headcount of Students with Exceptionalities by Category",
+        x = "School Year",
+        y = "Headcount")
 
@@
   geom_line(aes(color=disability))+
-  facet_wrap(~year)
+  facet_wrap(~year)+
+    labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
 
@@
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
- facet_wrap(~disability)
+ facet_wrap(~disability)+
+    labs(color="frequency", x= "Year", y="Student Count", title = "Trends in Frequency")
 
 ```
-
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Participants
-
-## Material
 

### 2020-11-19 scripts/draft_play.Rmd
comments and code edits
I included some general comments and also added some plots that you might find useful.

@@
 
+<!-- I'm sure you already know this, but you should fix the YAML. -->
+
 # Introduction
@@
 Two fringe rural districts exhibiting high populations will be picked out and  high-incidence diagnoses will be compared to those of other regions over time.
+
+<!-- It seems like you have a lot of interesting questions to answer with many different analyses.  Super cool.   -->
 
@@
 
+<!-- Personal preference suggestion; I would rather see the x-axis labels not tilted 90 degrees. -->
+
 ```{r urban vs. rural}
@@
 If the population was below 99,999 individuals, then it is classified as rural.
+
+<!-- That is an interesting distinction between rural and urban.  I had not thought to classify those distinctions by population.   -->
 
@@
 ```
+<!-- I think the previous code chunk would be a good place to include comments for your code.
+This would help a reader, or someone who may also work with this dataset or your code specifically.
+It isn't clear to me as a naive reader how and why you are tidying the way you are.
+Include some comments to help walk us through the good work you are doing here.
+-->
+
+<!-- Also, be sure to use the chunk options to indicate if you want the code to show when you knit. -->
+
+<!-- When I look into the fresh_primo data I notice that there are some observations which have a frequency of -9.
+What does that mean?
+Do you need to deal with that before generating summary statistics?
+-->
+
+<!-- This also might be a good place to include an APA table or kable.  It seems like the "fresh_primo" df is what you want to work with, or show the reader. -->
+
+```{r}
+
+# Here is a simple kable
+
+fresh_primo %>% 
+  group_by(year, disability) %>% 
+  summarize(total=sum(frequency, na.rm=TRUE)) %>% 
+  knitr::kable()
+
+# Here is the apa_table usage. https://www.rdocumentation.org/packages/papaja/versions/0.1.0.9997/topics/apa_table
+
+primo_apa <- apa_table(fresh_primo, caption = "Learning Disabilities in Oregon, 2002-2019", escape = TRUE)
+
+```
 
@@
 ```
+
+<!-- AWESOME PLOT!!!
+Super informative and nice to look at.
+Another nice option might be to facet on Age and use year as the independent variable.
+I think it would be interesting to see the trends across time.
+-->
+
+```{r}
+
+ggplot(fresh_primo, aes(year, frequency))+
+  geom_line(aes(color=disability))+
+  facet_wrap(~age)+
+    labs(color="Disability", x= "Year", y="Student Count", title ="Disability Prevalence By Year")
+```
+
+<!-- One additional thought I have about the plots.
+The age data is not continuous, it is discrete, (i.e.
+age 6, 7, 8, not 6, 6.1, 6,2, 6,3...) but you used geom_line here.
+I think it is effective, but it does suggest that there is data for ages between the discrete years.
+Maybe a geom_point + geom_smooth would be more accurate.
+However, with 11 disabilities, 11 sets of points and 11 fit lines might get messy.
+Again, I think your choice is effective in displaying the trend of the data, so if you end up using these plots you should make a note about the continuous/discrete nature of the data.
+-->
+
+<!-- It might also be nice to see the totals across time without faceting on age. -->
+
+```{r}
+by_age <- fresh_primo %>% 
+  group_by(year, disability) %>% 
+  summarise(frequency = sum(frequency))
+
+by_age %>% 
+  ggplot(aes(year, frequency))+
+  geom_point()+
+  geom_smooth()+
+  facet_wrap(~disability, scales = "free_y")+
+    labs(color="Disability", x= "Year", y="Student Count", title ="Disability Prevalence By Year", subtitle = "Data displays sums of all age categories")
+
+```
+
+<!-- It seems like there are some weird points in the data for Developmental Delay and Multiple Disabilities.
+It seems related to some of those negative terms, or maybe some NA's.
+If you end up using plots like this, it is obviously something you want to look into.
+Also, play around with eh x-axis, it would be nice to see all the years displayed instead of at 5 year intervals.
+-->
 
@@
 ```
+
+<!-- Another really nice plot.
+Some of my research relates to Autism so I find this very interesting.
+It seems like there is an obvious age related trend in this data.
+Perhaps you will comment on that in the analysis or discussion section.
+-->
+
+<!-- Sorry for reiterating here, but again, this data is not technically continuous.
+You might also use some of the other presentations I used about to combine the different age groups into a total, and plot across time with year as the independent variable.
+-->
+
+```{r}
+by_age %>% 
+filter(disability=="AUTISM") %>%
+  ggplot(aes(year, frequency)) +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(~disability, scales = "free_y") +
+    labs(color="Disability", x= "Year", y="Student Count", title ="Prevalence of Autism", subtitle = "Data displays the sum of all age categories.")
+```
+
 ```{r}
@@
 ```
+
+<!-- OK cool!
+It looks like y'all had the same idea I expressed above.
+Obviously this is a great way to display the data as we both came to that conclusion separately.
+I think these plots are accurate, but check out the code I used in lines 260-270, it includes a few extra bells and whistles to make it look pretty.
+-->
+
+<!-- At this point, I went back to check your intro for the research questions you mentioned.
+You mentioned being interested in the data trends with respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers.
+In seeing this, I think it is important to highlight the 2016 year on your plots.  You can include a line, maybe a dashed line to show the cutoff of the 2016 court ruling. -->
+
+```{r}
+
+# TRYING TO ADD ABLINE at year 2016
+
+by_age %>% 
+filter(disability=="AUTISM") %>%
+  ggplot(aes(year, frequency)) +
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 2016) + # without any line parameters you get a simple black line, check below for prettifying options
+  facet_wrap(~disability, scales = "free_y") +
+    labs(color="Disability", x= "Year", y="Student Count", title ="Prevalence of Autism", subtitle = "Data displays the sum of all age categories.")
+
+by_age %>% 
+  ggplot(aes(year, frequency))+
+  geom_point() +
+  geom_smooth() +
+  geom_vline(xintercept = 2016, col = "green", lty=3, lwd= 1) + # you can change the vertical line type with lty, the width with lwd, and the color with col
+  facet_wrap(~disability, scales = "free_y") +
+    labs(color="Disability", x= "Year", y="Student Count", title ="Disability Prevalence By Year", subtitle = "Data displays sums of all age categories")
+
+```
+
+<!-- I think y'all have done a really nice job here!
+Your analysis and discussion of the plots will be interesting to hear about.
+It's not obvious to me as a naive reader how your displays of the data help expose the answers to your research questions of interest.
+But I think your writeup will help make that more clear.
+-->
 

### 2020-11-19 scripts/draft_play.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The title},
  pdfauthor={Makayla Whitney1, Ernst-August Doelle1, Ernst-August Doelle1, \& Ernst-August Doelle1},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Title}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[shorthands=off,main=english]{babel}
\fi

\title{The title}
\author{Makayla Whitney\textsuperscript{1}, Ernst-August Doelle\textsuperscript{1}, Ernst-August Doelle\textsuperscript{1}, \& Ernst-August Doelle\textsuperscript{1}}
\date{}

\authornote{

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

The authors made the following contributions. Makayla Whitney: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Makayla Whitney, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Oregon}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background}, comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general context}.

Two or three sentences to provide a \textbf{broader perspective}, readily comprehensible to a scientist in any discipline.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21. The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon. The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020. The OR head counts include years 2002-2020. Levels of categorization include school- district- and provincial-level head counts for BC. The OR data set includes state-wide head counts that are not aggregated by school or district.

We intend to explore several questions regarding longitudinal trends. Firstly, we hope to analyze trends in disability prevalence over time. We will engage in a discussion on developmental trajectories by studying how trends shift from static/linear to increasing linear trends based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data. In studying the differences and similarities between the two datasets we will also engage in a discussion on diagnosis terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM), in part as a response to a challenge set by differing terminology between BC/OR categorization.

Secondly, we hope to more closely analyze any changes, or lack thereof, within the BC data following the year 2016 during which a BC Supreme Court ruled in favor of limiting the number of special needs students in classrooms and expanding the number of specialist teachers schools are required to hire.

Finally, we will also explore differences between urban and rural school districts in BC. Districts are categorized by population size and proximity to metropolitan areas, as recorded and defined by the Statistics Canada census from 2016. Two fringe rural districts exhibiting high populations will be picked out and high-incidence diagnoses will be compared to those of other regions over time.

\hypertarget{problem-statement-and-rqs}{%
\section{Problem Statement and RQs}\label{problem-statement-and-rqs}}

In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker \& Bergmann, 2013). While Canadian K-12 academic achievement outcomes are viewed as favourable on the world stage, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education. Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur. The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.

With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
(1) Do student prevalence rates of disability or disorder change after the supreme court ruling of 2016?
(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?

\hypertarget{results}{%
\section{Results}\label{results}}

Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time. The figure below demonstrates growth of 12 potential designations over an 18 year time period:

\includegraphics{draft_play_files/figure-latex/sandbox-1.pdf}

The district classification data was scaled down to include only public schools, while excluding private institutions. The school districts of Southeast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gwaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga'a, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information. The urban vs.~rural classifications were made based on the district's population on the 2016 census. If the population was above 100,000 individuals, it is classified as urban. If the population was below 99,999 individuals, then it is classified as rural.

This table displays the census results from 2011 and 2016 for our school districts. Many of the populations have stayed consistent within their urban or rural category. Three districts to note from the table are Nanaimo, Kamloops, and Chilliwack. In 2011, they were rural, but for our dataset they have been classified as urban due to their population increase in 2016.

\begin{verbatim}
## # A tibble: 198 x 18
##    year  disability    x6    x7    x8    x9   x10   x11   x12   x13   x14   x15
##    <chr> <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
##  1 2002~ INTELLECT~   135   172   220   261   279   303   370   364   388   378
##  2 2002~ HEARING I~    46    66    69    67    79    78    69    70    53    72
##  3 2002~ SPEECH OR~  2502  2466  2443  2205  1894  1282   874   677   493   340
##  4 2002~ VISUAL IM~    18    27    25    16    12    26    19    24    31    19
##  5 2002~ EMOTIONAL~    81   178   241   292   347   433   504   487   549   548
##  6 2002~ ORTHOPEDI~    65    77    64    49    64    71    50    48    44    43
##  7 2002~ DEAF-BLIN~     0     2     0     0     1     0     2     0     4     2
##  8 2002~ MULTIPLE ~     0     0     0     0     0     0     0     0     0     0
##  9 2002~ AUTISM       267   289   290   334   325   316   290   256   243   219
## 10 2002~ TRAUMATIC~     6    11    17    23    22    19    21    23    36    36
## # ... with 188 more rows, and 6 more variables: x16 <dbl>, x17 <dbl>,
## #   x18 <dbl>, x19 <dbl>, x20 <dbl>, x21 <dbl>
\end{verbatim}

\begin{verbatim}
## # A tibble: 198 x 3
## # Groups:   year [18]
##     year disability                    total
##    <dbl> <chr>                         <dbl>
##  1  2002 AUTISM                         3339
##  2  2002 DEAF-BLINDNESS                   17
##  3  2002 DEVELOPMENTAL DELAY               0
##  4  2002 EMOTIONAL DISTURBANCE          4736
##  5  2002 HEARING IMPAIRMENT              873
##  6  2002 INTELLECTUAL DISABILITY        4387
##  7  2002 MULTIPLE DISABILITIES             0
##  8  2002 ORTHOPEDIC IMPAIRMENT           754
##  9  2002 SPEECH OR LANGUAGE IMPAIRMENT 15784
## 10  2002 TRAUMATIC BRAIN INJURY          306
## # ... with 188 more rows
\end{verbatim}

\begin{tabular}{r|l|r}
\hline
year & disability & total\\
\hline
2002 & AUTISM & 3339\\
\hline
2002 & DEAF-BLINDNESS & 17\\
\hline
2002 & DEVELOPMENTAL DELAY & 0\\
\hline
2002 & EMOTIONAL DISTURBANCE & 4736\\
\hline
2002 & HEARING IMPAIRMENT & 873\\
\hline
2002 & INTELLECTUAL DISABILITY & 4387\\
\hline
2002 & MULTIPLE DISABILITIES & 0\\
\hline
2002 & ORTHOPEDIC IMPAIRMENT & 754\\
\hline
2002 & SPEECH OR LANGUAGE IMPAIRMENT & 15784\\
\hline
2002 & TRAUMATIC BRAIN INJURY & 306\\
\hline
2002 & VISUAL IMPAIRMENT & 297\\
\hline
2003 & AUTISM & 3759\\
\hline
2003 & DEAF-BLINDNESS & 17\\
\hline
2003 & DEVELOPMENTAL DELAY & 0\\
\hline
2003 & EMOTIONAL DISTURBANCE & 4597\\
\hline
2003 & HEARING IMPAIRMENT & 811\\
\hline
2003 & INTELLECTUAL DISABILITY & 4351\\
\hline
2003 & MULTIPLE DISABILITIES & 0\\
\hline
2003 & ORTHOPEDIC IMPAIRMENT & 749\\
\hline
2003 & SPEECH OR LANGUAGE IMPAIRMENT & 15642\\
\hline
2003 & TRAUMATIC BRAIN INJURY & 300\\
\hline
2003 & VISUAL IMPAIRMENT & 295\\
\hline
2004 & AUTISM & 4341\\
\hline
2004 & DEAF-BLINDNESS & 12\\
\hline
2004 & DEVELOPMENTAL DELAY & 0\\
\hline
2004 & EMOTIONAL DISTURBANCE & 4661\\
\hline
2004 & HEARING IMPAIRMENT & 809\\
\hline
2004 & INTELLECTUAL DISABILITY & 4334\\
\hline
2004 & MULTIPLE DISABILITIES & 0\\
\hline
2004 & ORTHOPEDIC IMPAIRMENT & 741\\
\hline
2004 & SPEECH OR LANGUAGE IMPAIRMENT & 15627\\
\hline
2004 & TRAUMATIC BRAIN INJURY & 286\\
\hline
2004 & VISUAL IMPAIRMENT & 325\\
\hline
2005 & AUTISM & 4810\\
\hline
2005 & DEAF-BLINDNESS & 14\\
\hline
2005 & DEVELOPMENTAL DELAY & 0\\
\hline
2005 & EMOTIONAL DISTURBANCE & 4661\\
\hline
2005 & HEARING IMPAIRMENT & 799\\
\hline
2005 & INTELLECTUAL DISABILITY & 4218\\
\hline
2005 & MULTIPLE DISABILITIES & 0\\
\hline
2005 & ORTHOPEDIC IMPAIRMENT & 738\\
\hline
2005 & SPEECH OR LANGUAGE IMPAIRMENT & 15833\\
\hline
2005 & TRAUMATIC BRAIN INJURY & 273\\
\hline
2005 & VISUAL IMPAIRMENT & 301\\
\hline
2006 & AUTISM & 5459\\
\hline
2006 & DEAF-BLINDNESS & 12\\
\hline
2006 & DEVELOPMENTAL DELAY & 0\\
\hline
2006 & EMOTIONAL DISTURBANCE & 4662\\
\hline
2006 & HEARING IMPAIRMENT & 812\\
\hline
2006 & INTELLECTUAL DISABILITY & 4193\\
\hline
2006 & MULTIPLE DISABILITIES & 0\\
\hline
2006 & ORTHOPEDIC IMPAIRMENT & 742\\
\hline
2006 & SPEECH OR LANGUAGE IMPAIRMENT & 15956\\
\hline
2006 & TRAUMATIC BRAIN INJURY & 269\\
\hline
2006 & VISUAL IMPAIRMENT & 302\\
\hline
2007 & AUTISM & 6071\\
\hline
2007 & DEAF-BLINDNESS & 21\\
\hline
2007 & DEVELOPMENTAL DELAY & 0\\
\hline
2007 & EMOTIONAL DISTURBANCE & 4631\\
\hline
2007 & HEARING IMPAIRMENT & 835\\
\hline
2007 & INTELLECTUAL DISABILITY & 4121\\
\hline
2007 & MULTIPLE DISABILITIES & 0\\
\hline
2007 & ORTHOPEDIC IMPAIRMENT & 745\\
\hline
2007 & SPEECH OR LANGUAGE IMPAIRMENT & 16046\\
\hline
2007 & TRAUMATIC BRAIN INJURY & 258\\
\hline
2007 & VISUAL IMPAIRMENT & 325\\
\hline
2008 & AUTISM & 6571\\
\hline
2008 & DEAF-BLINDNESS & 15\\
\hline
2008 & DEVELOPMENTAL DELAY & 0\\
\hline
2008 & EMOTIONAL DISTURBANCE & 4671\\
\hline
2008 & HEARING IMPAIRMENT & 823\\
\hline
2008 & INTELLECTUAL DISABILITY & 4012\\
\hline
2008 & MULTIPLE DISABILITIES & 0\\
\hline
2008 & ORTHOPEDIC IMPAIRMENT & 746\\
\hline
2008 & SPEECH OR LANGUAGE IMPAIRMENT & 16116\\
\hline
2008 & TRAUMATIC BRAIN INJURY & 279\\
\hline
2008 & VISUAL IMPAIRMENT & 317\\
\hline
2009 & AUTISM & 7012\\
\hline
2009 & DEAF-BLINDNESS & 8\\
\hline
2009 & DEVELOPMENTAL DELAY & 0\\
\hline
2009 & EMOTIONAL DISTURBANCE & 4669\\
\hline
2009 & HEARING IMPAIRMENT & 830\\
\hline
2009 & INTELLECTUAL DISABILITY & 4000\\
\hline
2009 & MULTIPLE DISABILITIES & 0\\
\hline
2009 & ORTHOPEDIC IMPAIRMENT & 713\\
\hline
2009 & SPEECH OR LANGUAGE IMPAIRMENT & 16132\\
\hline
2009 & TRAUMATIC BRAIN INJURY & 283\\
\hline
2009 & VISUAL IMPAIRMENT & 314\\
\hline
2010 & AUTISM & 7357\\
\hline
2010 & DEAF-BLINDNESS & 8\\
\hline
2010 & DEVELOPMENTAL DELAY & -144\\
\hline
2010 & EMOTIONAL DISTURBANCE & 4638\\
\hline
2010 & HEARING IMPAIRMENT & 827\\
\hline
2010 & INTELLECTUAL DISABILITY & 3926\\
\hline
2010 & MULTIPLE DISABILITIES & -144\\
\hline
2010 & ORTHOPEDIC IMPAIRMENT & 718\\
\hline
2010 & SPEECH OR LANGUAGE IMPAIRMENT & 16354\\
\hline
2010 & TRAUMATIC BRAIN INJURY & 271\\
\hline
2010 & VISUAL IMPAIRMENT & 312\\
\hline
2011 & AUTISM & 7618\\
\hline
2011 & DEAF-BLINDNESS & 12\\
\hline
2011 & DEVELOPMENTAL DELAY & -36\\
\hline
2011 & EMOTIONAL DISTURBANCE & 4567\\
\hline
2011 & HEARING IMPAIRMENT & 810\\
\hline
2011 & INTELLECTUAL DISABILITY & 3830\\
\hline
2011 & MULTIPLE DISABILITIES & -144\\
\hline
2011 & ORTHOPEDIC IMPAIRMENT & 738\\
\hline
2011 & SPEECH OR LANGUAGE IMPAIRMENT & 16206\\
\hline
2011 & TRAUMATIC BRAIN INJURY & 268\\
\hline
2011 & VISUAL IMPAIRMENT & 313\\
\hline
2012 & AUTISM & 7888\\
\hline
2012 & DEAF-BLINDNESS & 10\\
\hline
2012 & DEVELOPMENTAL DELAY & -54\\
\hline
2012 & EMOTIONAL DISTURBANCE & 4524\\
\hline
2012 & HEARING IMPAIRMENT & 810\\
\hline
2012 & INTELLECTUAL DISABILITY & 3847\\
\hline
2012 & MULTIPLE DISABILITIES & -144\\
\hline
2012 & ORTHOPEDIC IMPAIRMENT & 705\\
\hline
2012 & SPEECH OR LANGUAGE IMPAIRMENT & 15938\\
\hline
2012 & TRAUMATIC BRAIN INJURY & 256\\
\hline
2012 & VISUAL IMPAIRMENT & 300\\
\hline
2013 & AUTISM & 8021\\
\hline
2013 & DEAF-BLINDNESS & 9\\
\hline
2013 & DEVELOPMENTAL DELAY & -36\\
\hline
2013 & EMOTIONAL DISTURBANCE & 4526\\
\hline
2013 & HEARING IMPAIRMENT & 830\\
\hline
2013 & INTELLECTUAL DISABILITY & 3847\\
\hline
2013 & MULTIPLE DISABILITIES & -144\\
\hline
2013 & ORTHOPEDIC IMPAIRMENT & 700\\
\hline
2013 & SPEECH OR LANGUAGE IMPAIRMENT & 16077\\
\hline
2013 & TRAUMATIC BRAIN INJURY & 254\\
\hline
2013 & VISUAL IMPAIRMENT & 288\\
\hline
2014 & AUTISM & 8364\\
\hline
2014 & DEAF-BLINDNESS & 5\\
\hline
2014 & DEVELOPMENTAL DELAY & -36\\
\hline
2014 & EMOTIONAL DISTURBANCE & 4577\\
\hline
2014 & HEARING IMPAIRMENT & 851\\
\hline
2014 & INTELLECTUAL DISABILITY & 3874\\
\hline
2014 & MULTIPLE DISABILITIES & -144\\
\hline
2014 & ORTHOPEDIC IMPAIRMENT & 673\\
\hline
2014 & SPEECH OR LANGUAGE IMPAIRMENT & 16236\\
\hline
2014 & TRAUMATIC BRAIN INJURY & 248\\
\hline
2014 & VISUAL IMPAIRMENT & 296\\
\hline
2015 & AUTISM & 8665\\
\hline
2015 & DEAF-BLINDNESS & 5\\
\hline
2015 & DEVELOPMENTAL DELAY & -36\\
\hline
2015 & EMOTIONAL DISTURBANCE & 4704\\
\hline
2015 & HEARING IMPAIRMENT & 844\\
\hline
2015 & INTELLECTUAL DISABILITY & 3951\\
\hline
2015 & MULTIPLE DISABILITIES & -144\\
\hline
2015 & ORTHOPEDIC IMPAIRMENT & 649\\
\hline
2015 & SPEECH OR LANGUAGE IMPAIRMENT & 16294\\
\hline
2015 & TRAUMATIC BRAIN INJURY & 240\\
\hline
2015 & VISUAL IMPAIRMENT & 291\\
\hline
2016 & AUTISM & 8940\\
\hline
2016 & DEAF-BLINDNESS & 8\\
\hline
2016 & DEVELOPMENTAL DELAY & 0\\
\hline
2016 & EMOTIONAL DISTURBANCE & 4944\\
\hline
2016 & HEARING IMPAIRMENT & 829\\
\hline
2016 & INTELLECTUAL DISABILITY & 4058\\
\hline
2016 & MULTIPLE DISABILITIES & -144\\
\hline
2016 & ORTHOPEDIC IMPAIRMENT & 624\\
\hline
2016 & SPEECH OR LANGUAGE IMPAIRMENT & 16075\\
\hline
2016 & TRAUMATIC BRAIN INJURY & 261\\
\hline
2016 & VISUAL IMPAIRMENT & 306\\
\hline
2017 & AUTISM & 9288\\
\hline
2017 & DEAF-BLINDNESS & 12\\
\hline
2017 & DEVELOPMENTAL DELAY & 0\\
\hline
2017 & EMOTIONAL DISTURBANCE & 5122\\
\hline
2017 & HEARING IMPAIRMENT & 841\\
\hline
2017 & INTELLECTUAL DISABILITY & 4103\\
\hline
2017 & MULTIPLE DISABILITIES & -144\\
\hline
2017 & ORTHOPEDIC IMPAIRMENT & 612\\
\hline
2017 & SPEECH OR LANGUAGE IMPAIRMENT & 16087\\
\hline
2017 & TRAUMATIC BRAIN INJURY & 282\\
\hline
2017 & VISUAL IMPAIRMENT & 282\\
\hline
2018 & AUTISM & 9752\\
\hline
2018 & DEAF-BLINDNESS & 11\\
\hline
2018 & DEVELOPMENTAL DELAY & -36\\
\hline
2018 & EMOTIONAL DISTURBANCE & 5286\\
\hline
2018 & HEARING IMPAIRMENT & 848\\
\hline
2018 & INTELLECTUAL DISABILITY & 4167\\
\hline
2018 & MULTIPLE DISABILITIES & -144\\
\hline
2018 & ORTHOPEDIC IMPAIRMENT & 602\\
\hline
2018 & SPEECH OR LANGUAGE IMPAIRMENT & 16357\\
\hline
2018 & TRAUMATIC BRAIN INJURY & 294\\
\hline
2018 & VISUAL IMPAIRMENT & 269\\
\hline
2019 & AUTISM & 10290\\
\hline
2019 & DEAF-BLINDNESS & 21\\
\hline
2019 & DEVELOPMENTAL DELAY & 266\\
\hline
2019 & EMOTIONAL DISTURBANCE & 5568\\
\hline
2019 & HEARING IMPAIRMENT & 870\\
\hline
2019 & INTELLECTUAL DISABILITY & 4199\\
\hline
2019 & MULTIPLE DISABILITIES & -144\\
\hline
2019 & ORTHOPEDIC IMPAIRMENT & 562\\
\hline
2019 & SPEECH OR LANGUAGE IMPAIRMENT & 16423\\
\hline
2019 & TRAUMATIC BRAIN INJURY & 327\\
\hline
2019 & VISUAL IMPAIRMENT & 282\\
\hline
\end{tabular}

\begin{verbatim}
## Warning: Removed 12 row(s) containing missing values (geom_path).
\end{verbatim}

\includegraphics{draft_play_files/figure-latex/Oregon plot-1.pdf}

\includegraphics{draft_play_files/figure-latex/unnamed-chunk-2-1.pdf}

\begin{verbatim}
## Warning: Removed 190 rows containing non-finite values (stat_smooth).
\end{verbatim}

\includegraphics{draft_play_files/figure-latex/unnamed-chunk-3-1.pdf}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We used R (Version 3.6.1; R Core Team, 2020) and the R-package \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2020) for all our analyses.

\hypertarget{results-1}{%
\section{Results}\label{results-1}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-papaja}{}%
Aust, F., \& Barth, M. (2020). \emph{papaja: Create APA manuscripts with R Markdown}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\endgroup

\end{document}

### 2020-11-23 scripts/draft_play.Rmd
Xiaoyu Liu_peer review

--- 2020-11-18 scripts/draft_play.Rmd
@@
         y = "Headcount")
-
+#strength 1: I really like the data visualization. It is very clear! 
 ```
@@
 
+# strength 2: I really like the data cleaning part. They used pivot_longer to make the data more clear! 
 ```
@@
 
+# Strength 3: Remove_empty: this function is great to remove NAs.
 cleanest_primo <- cleanest_primo %>% 
@@
 
+# This is the part the I learned.
+I learned the parse_number function to .This drops any non-numeric characters before or after the first number.
+The grouping mark specified by the locale is ignored inside the number.
 ```
@@
 
+#improvement: I am not sure what does the second ```r``` includes. maybe you could delete. 
 ```

### 2020-11-23 scripts/draft_play.Rmd
Peer Review

--- 2020-11-18 scripts/draft_play.Rmd
@@
 shorttitle        : "Title"
-
 author: 
@@
       - Writing - Review & Editing
-
 affiliation:
@@
     institution   : "University of Oregon"
-
 authornote: |
   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
   Enter author note here.
-
 abstract: |
@@
   <!-- https://tinyurl.com/ybremelq -->
-
 keywords          : "keywords"
 wordcount         : "X"
-
 bibliography      : ["r-references.bib"]
-
 floatsintext      : no
@@
 draft             : no
-
 documentclass     : "apa6"
@@
 output            : papaja::apa6_pdf
+editor_options: 
+  chunk_output_type: console
 ---

### 2020-11-23 scripts/draft_play.Rmd

--- 2020-11-23 scripts/draft_play.Rmd
@@
 ```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
-library("papaja")
-r_refs("r-references.bib")
+
+library(papaja)
 library(here)
 library(tidyverse)
-library("readxl")
+library(readxl)
 library(janitor)
+r_refs('r-references.bib')
+#avoid using quotation marks when using library() command
+
+library(ggplot2)
+#library("ggplot2")
+#why leave the two same commands?
 
 BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
+#the best practice would be to use clean_names() and as_tibble() here, in particularly because janitor package was loaded
 
@@
 library("readxl")
+#no need to load the same packages, except for "readxl"
 
@@
          SPECIAL_NEEDS_CODE != "N")
+#factoring was nicely done here!
 
@@
         y = "Headcount")
-#strength 1: I really like the data visualization. It is very clear! 
+
+#nice visualization! It's interesting to observe the yearly trend of students with exceptionalities changes by year differently
+#using the ratio of students by category instead of headcount will be more intuitive. use some nice-looking themes as well
+
+BCdatastripped_spedcat2 <- BCdatastripped_spedcat %>%
+  group_by(SCHOOL_YEAR) %>%
+  mutate(total = sum(NUMBER_OF_STUDENTS)) %>% 
+  mutate(ratio = NUMBER_OF_STUDENTS / total)
+
+ggplot(BCdatastripped_spedcat2, aes(x = SCHOOL_YEAR, y = ratio)) +
+  geom_line(color = "cornflowerblue") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+   facet_wrap(~SPECIAL_NEEDS_CODE) +
+   labs(title = "BC Provincial Ratio of Students with Exceptionalities by Category",
+        x = "School Year",
+        y = "Ratio") + 
+   theme_minimal()
+
 ```
@@
 bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))
+#clean_names and as_tibble?
 
@@
          PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
-  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
+  pivot_longer(DISTRICT_CLASSIFICATION, 
+               values_drop_na = TRUE, 
+               values_to = "DISTRICT_CLASSIFICATION") %>% 
     select(-"name")
 
-# strength 2: I really like the data cleaning part. They used pivot_longer to make the data more clear! 
 ```
@@
 
-# Strength 3: Remove_empty: this function is great to remove NAs.
 cleanest_primo <- cleanest_primo %>% 
@@
 
-# This is the part the I learned.
-I learned the parse_number function to .This drops any non-numeric characters before or after the first number.
-The grouping mark specified by the locale is ignored inside the number.
 ```
@@
 ```{r Oregon plot}
-
 ggplot(fresh_primo, aes(age, frequency))+
@@
     labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
+#nicely done as well, but it using 4 variables in one plot can be tricky. showing disability prevalence by age, but for each year would be one way to solve it
+
+fresh_primo %>% 
+  filter(year == 2002) %>% 
+  ggplot(aes(age, frequency))+
+  geom_line(aes(color=disability))+
+  labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age in 2002") + 
+  theme_mi
 
@@
   filter(disability=="AUTISM") %>%
-
-ggplot(aes(age, frequency))+
- geom_line(aes(color=disability))+
+  ggplot(aes(age, frequency)) +
+  geom_line(aes(color=disability)) + 
   facet_wrap(~year)
@@
 ggplot(fresh_primo, aes(year, frequency))+
- geom_smooth(color = "cornflowerblue") +
+ geom_smooth(color = "cornflowerblue", se = FALSE) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
@@
     labs(color="frequency", x= "Year", y="Student Count", title = "Trends in Frequency")
+#geom_smooth was used instead of geom_line because of yearly fluctuation. Were only even/odd years got recorded?
 
-#improvement: I am not sure what does the second ```r``` includes. maybe you could delete. 
 ```

### 2020-11-23 scripts/draft_play.tex

@@
 
-\begin{tabular}{r|l|r}
-\hline
-year & disability & total\\
-\hline
-2002 & AUTISM & 3339\\
-\hline
-2002 & DEAF-BLINDNESS & 17\\
-\hline
-2002 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2002 & EMOTIONAL DISTURBANCE & 4736\\
-\hline
-2002 & HEARING IMPAIRMENT & 873\\
-\hline
-2002 & INTELLECTUAL DISABILITY & 4387\\
-\hline
-2002 & MULTIPLE DISABILITIES & 0\\
-\hline
-2002 & ORTHOPEDIC IMPAIRMENT & 754\\
-\hline
-2002 & SPEECH OR LANGUAGE IMPAIRMENT & 15784\\
-\hline
-2002 & TRAUMATIC BRAIN INJURY & 306\\
-\hline
-2002 & VISUAL IMPAIRMENT & 297\\
-\hline
-2003 & AUTISM & 3759\\
-\hline
-2003 & DEAF-BLINDNESS & 17\\
-\hline
-2003 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2003 & EMOTIONAL DISTURBANCE & 4597\\
-\hline
-2003 & HEARING IMPAIRMENT & 811\\
-\hline
-2003 & INTELLECTUAL DISABILITY & 4351\\
-\hline
-2003 & MULTIPLE DISABILITIES & 0\\
-\hline
-2003 & ORTHOPEDIC IMPAIRMENT & 749\\
-\hline
-2003 & SPEECH OR LANGUAGE IMPAIRMENT & 15642\\
-\hline
-2003 & TRAUMATIC BRAIN INJURY & 300\\
-\hline
-2003 & VISUAL IMPAIRMENT & 295\\
-\hline
-2004 & AUTISM & 4341\\
-\hline
-2004 & DEAF-BLINDNESS & 12\\
-\hline
-2004 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2004 & EMOTIONAL DISTURBANCE & 4661\\
-\hline
-2004 & HEARING IMPAIRMENT & 809\\
-\hline
-2004 & INTELLECTUAL DISABILITY & 4334\\
-\hline
-2004 & MULTIPLE DISABILITIES & 0\\
-\hline
-2004 & ORTHOPEDIC IMPAIRMENT & 741\\
-\hline
-2004 & SPEECH OR LANGUAGE IMPAIRMENT & 15627\\
-\hline
-2004 & TRAUMATIC BRAIN INJURY & 286\\
-\hline
-2004 & VISUAL IMPAIRMENT & 325\\
-\hline
-2005 & AUTISM & 4810\\
-\hline
-2005 & DEAF-BLINDNESS & 14\\
-\hline
-2005 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2005 & EMOTIONAL DISTURBANCE & 4661\\
-\hline
-2005 & HEARING IMPAIRMENT & 799\\
-\hline
-2005 & INTELLECTUAL DISABILITY & 4218\\
-\hline
-2005 & MULTIPLE DISABILITIES & 0\\
-\hline
-2005 & ORTHOPEDIC IMPAIRMENT & 738\\
-\hline
-2005 & SPEECH OR LANGUAGE IMPAIRMENT & 15833\\
-\hline
-2005 & TRAUMATIC BRAIN INJURY & 273\\
-\hline
-2005 & VISUAL IMPAIRMENT & 301\\
-\hline
-2006 & AUTISM & 5459\\
-\hline
-2006 & DEAF-BLINDNESS & 12\\
-\hline
-2006 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2006 & EMOTIONAL DISTURBANCE & 4662\\
-\hline
-2006 & HEARING IMPAIRMENT & 812\\
-\hline
-2006 & INTELLECTUAL DISABILITY & 4193\\
-\hline
-2006 & MULTIPLE DISABILITIES & 0\\
-\hline
-2006 & ORTHOPEDIC IMPAIRMENT & 742\\
-\hline
-2006 & SPEECH OR LANGUAGE IMPAIRMENT & 15956\\
-\hline
-2006 & TRAUMATIC BRAIN INJURY & 269\\
-\hline
-2006 & VISUAL IMPAIRMENT & 302\\
-\hline
-2007 & AUTISM & 6071\\
-\hline
-2007 & DEAF-BLINDNESS & 21\\
-\hline
-2007 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2007 & EMOTIONAL DISTURBANCE & 4631\\
-\hline
-2007 & HEARING IMPAIRMENT & 835\\
-\hline
-2007 & INTELLECTUAL DISABILITY & 4121\\
-\hline
-2007 & MULTIPLE DISABILITIES & 0\\
-\hline
-2007 & ORTHOPEDIC IMPAIRMENT & 745\\
-\hline
-2007 & SPEECH OR LANGUAGE IMPAIRMENT & 16046\\
-\hline
-2007 & TRAUMATIC BRAIN INJURY & 258\\
-\hline
-2007 & VISUAL IMPAIRMENT & 325\\
-\hline
-2008 & AUTISM & 6571\\
-\hline
-2008 & DEAF-BLINDNESS & 15\\
-\hline
-2008 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2008 & EMOTIONAL DISTURBANCE & 4671\\
-\hline
-2008 & HEARING IMPAIRMENT & 823\\
-\hline
-2008 & INTELLECTUAL DISABILITY & 4012\\
-\hline
-2008 & MULTIPLE DISABILITIES & 0\\
-\hline
-2008 & ORTHOPEDIC IMPAIRMENT & 746\\
-\hline
-2008 & SPEECH OR LANGUAGE IMPAIRMENT & 16116\\
-\hline
-2008 & TRAUMATIC BRAIN INJURY & 279\\
-\hline
-2008 & VISUAL IMPAIRMENT & 317\\
-\hline
-2009 & AUTISM & 7012\\
-\hline
-2009 & DEAF-BLINDNESS & 8\\
-\hline
-2009 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2009 & EMOTIONAL DISTURBANCE & 4669\\
-\hline
-2009 & HEARING IMPAIRMENT & 830\\
-\hline
-2009 & INTELLECTUAL DISABILITY & 4000\\
-\hline
-2009 & MULTIPLE DISABILITIES & 0\\
-\hline
-2009 & ORTHOPEDIC IMPAIRMENT & 713\\
-\hline
-2009 & SPEECH OR LANGUAGE IMPAIRMENT & 16132\\
-\hline
-2009 & TRAUMATIC BRAIN INJURY & 283\\
-\hline
-2009 & VISUAL IMPAIRMENT & 314\\
-\hline
-2010 & AUTISM & 7357\\
-\hline
-2010 & DEAF-BLINDNESS & 8\\
-\hline
-2010 & DEVELOPMENTAL DELAY & -144\\
-\hline
-2010 & EMOTIONAL DISTURBANCE & 4638\\
-\hline
-2010 & HEARING IMPAIRMENT & 827\\
-\hline
-2010 & INTELLECTUAL DISABILITY & 3926\\
-\hline
-2010 & MULTIPLE DISABILITIES & -144\\
-\hline
-2010 & ORTHOPEDIC IMPAIRMENT & 718\\
-\hline
-2010 & SPEECH OR LANGUAGE IMPAIRMENT & 16354\\
-\hline
-2010 & TRAUMATIC BRAIN INJURY & 271\\
-\hline
-2010 & VISUAL IMPAIRMENT & 312\\
-\hline
-2011 & AUTISM & 7618\\
-\hline
-2011 & DEAF-BLINDNESS & 12\\
-\hline
-2011 & DEVELOPMENTAL DELAY & -36\\
-\hline
-2011 & EMOTIONAL DISTURBANCE & 4567\\
-\hline
-2011 & HEARING IMPAIRMENT & 810\\
-\hline
-2011 & INTELLECTUAL DISABILITY & 3830\\
-\hline
-2011 & MULTIPLE DISABILITIES & -144\\
-\hline
-2011 & ORTHOPEDIC IMPAIRMENT & 738\\
-\hline
-2011 & SPEECH OR LANGUAGE IMPAIRMENT & 16206\\
-\hline
-2011 & TRAUMATIC BRAIN INJURY & 268\\
-\hline
-2011 & VISUAL IMPAIRMENT & 313\\
-\hline
-2012 & AUTISM & 7888\\
-\hline
-2012 & DEAF-BLINDNESS & 10\\
-\hline
-2012 & DEVELOPMENTAL DELAY & -54\\
-\hline
-2012 & EMOTIONAL DISTURBANCE & 4524\\
-\hline
-2012 & HEARING IMPAIRMENT & 810\\
-\hline
-2012 & INTELLECTUAL DISABILITY & 3847\\
-\hline
-2012 & MULTIPLE DISABILITIES & -144\\
-\hline
-2012 & ORTHOPEDIC IMPAIRMENT & 705\\
-\hline
-2012 & SPEECH OR LANGUAGE IMPAIRMENT & 15938\\
-\hline
-2012 & TRAUMATIC BRAIN INJURY & 256\\
-\hline
-2012 & VISUAL IMPAIRMENT & 300\\
-\hline
-2013 & AUTISM & 8021\\
-\hline
-2013 & DEAF-BLINDNESS & 9\\
-\hline
-2013 & DEVELOPMENTAL DELAY & -36\\
-\hline
-2013 & EMOTIONAL DISTURBANCE & 4526\\
-\hline
-2013 & HEARING IMPAIRMENT & 830\\
-\hline
-2013 & INTELLECTUAL DISABILITY & 3847\\
-\hline
-2013 & MULTIPLE DISABILITIES & -144\\
-\hline
-2013 & ORTHOPEDIC IMPAIRMENT & 700\\
-\hline
-2013 & SPEECH OR LANGUAGE IMPAIRMENT & 16077\\
-\hline
-2013 & TRAUMATIC BRAIN INJURY & 254\\
-\hline
-2013 & VISUAL IMPAIRMENT & 288\\
-\hline
-2014 & AUTISM & 8364\\
-\hline
-2014 & DEAF-BLINDNESS & 5\\
-\hline
-2014 & DEVELOPMENTAL DELAY & -36\\
-\hline
-2014 & EMOTIONAL DISTURBANCE & 4577\\
-\hline
-2014 & HEARING IMPAIRMENT & 851\\
-\hline
-2014 & INTELLECTUAL DISABILITY & 3874\\
-\hline
-2014 & MULTIPLE DISABILITIES & -144\\
-\hline
-2014 & ORTHOPEDIC IMPAIRMENT & 673\\
-\hline
-2014 & SPEECH OR LANGUAGE IMPAIRMENT & 16236\\
-\hline
-2014 & TRAUMATIC BRAIN INJURY & 248\\
-\hline
-2014 & VISUAL IMPAIRMENT & 296\\
-\hline
-2015 & AUTISM & 8665\\
-\hline
-2015 & DEAF-BLINDNESS & 5\\
-\hline
-2015 & DEVELOPMENTAL DELAY & -36\\
-\hline
-2015 & EMOTIONAL DISTURBANCE & 4704\\
-\hline
-2015 & HEARING IMPAIRMENT & 844\\
-\hline
-2015 & INTELLECTUAL DISABILITY & 3951\\
-\hline
-2015 & MULTIPLE DISABILITIES & -144\\
-\hline
-2015 & ORTHOPEDIC IMPAIRMENT & 649\\
-\hline
-2015 & SPEECH OR LANGUAGE IMPAIRMENT & 16294\\
-\hline
-2015 & TRAUMATIC BRAIN INJURY & 240\\
-\hline
-2015 & VISUAL IMPAIRMENT & 291\\
-\hline
-2016 & AUTISM & 8940\\
-\hline
-2016 & DEAF-BLINDNESS & 8\\
-\hline
-2016 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2016 & EMOTIONAL DISTURBANCE & 4944\\
-\hline
-2016 & HEARING IMPAIRMENT & 829\\
-\hline
-2016 & INTELLECTUAL DISABILITY & 4058\\
-\hline
-2016 & MULTIPLE DISABILITIES & -144\\
-\hline
-2016 & ORTHOPEDIC IMPAIRMENT & 624\\
-\hline
-2016 & SPEECH OR LANGUAGE IMPAIRMENT & 16075\\
-\hline
-2016 & TRAUMATIC BRAIN INJURY & 261\\
-\hline
-2016 & VISUAL IMPAIRMENT & 306\\
-\hline
-2017 & AUTISM & 9288\\
-\hline
-2017 & DEAF-BLINDNESS & 12\\
-\hline
-2017 & DEVELOPMENTAL DELAY & 0\\
-\hline
-2017 & EMOTIONAL DISTURBANCE & 5122\\
-\hline
-2017 & HEARING IMPAIRMENT & 841\\
-\hline
-2017 & INTELLECTUAL DISABILITY & 4103\\
-\hline
-2017 & MULTIPLE DISABILITIES & -144\\
-\hline
-2017 & ORTHOPEDIC IMPAIRMENT & 612\\
-\hline
-2017 & SPEECH OR LANGUAGE IMPAIRMENT & 16087\\
-\hline
-2017 & TRAUMATIC BRAIN INJURY & 282\\
-\hline
-2017 & VISUAL IMPAIRMENT & 282\\
-\hline
-2018 & AUTISM & 9752\\
-\hline
-2018 & DEAF-BLINDNESS & 11\\
-\hline
-2018 & DEVELOPMENTAL DELAY & -36\\
-\hline
-2018 & EMOTIONAL DISTURBANCE & 5286\\
-\hline
-2018 & HEARING IMPAIRMENT & 848\\
-\hline
-2018 & INTELLECTUAL DISABILITY & 4167\\
-\hline
-2018 & MULTIPLE DISABILITIES & -144\\
-\hline
-2018 & ORTHOPEDIC IMPAIRMENT & 602\\
-\hline
-2018 & SPEECH OR LANGUAGE IMPAIRMENT & 16357\\
-\hline
-2018 & TRAUMATIC BRAIN INJURY & 294\\
-\hline
-2018 & VISUAL IMPAIRMENT & 269\\
-\hline
-2019 & AUTISM & 10290\\
-\hline
-2019 & DEAF-BLINDNESS & 21\\
-\hline
-2019 & DEVELOPMENTAL DELAY & 266\\
-\hline
-2019 & EMOTIONAL DISTURBANCE & 5568\\
-\hline
-2019 & HEARING IMPAIRMENT & 870\\
-\hline
-2019 & INTELLECTUAL DISABILITY & 4199\\
-\hline
-2019 & MULTIPLE DISABILITIES & -144\\
-\hline
-2019 & ORTHOPEDIC IMPAIRMENT & 562\\
-\hline
-2019 & SPEECH OR LANGUAGE IMPAIRMENT & 16423\\
-\hline
-2019 & TRAUMATIC BRAIN INJURY & 327\\
-\hline
-2019 & VISUAL IMPAIRMENT & 282\\
-\hline
-\end{tabular}
-
 \begin{verbatim}
@@
 
-\includegraphics{draft_play_files/figure-latex/unnamed-chunk-2-1.pdf}
+\includegraphics{draft_play_files/figure-latex/unnamed-chunk-1-1.pdf}
 
@@
 
-\includegraphics{draft_play_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{draft_play_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
-We used R (Version 3.6.1; R Core Team, 2020) and the R-package \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2020) for all our analyses.
+We used R (Version 4.0.2; R Core Team, 2020) and the R-packages \emph{dplyr} (Version 1.0.2; Wickham et al., 2020), \emph{forcats} (Version 0.5.0; Wickham, 2020a), \emph{ggplot2} (Version 3.3.2; Wickham, 2016), \emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 2.0.1; Firke, 2020), \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2020), \emph{purrr} (Version 0.3.4; Henry \& Wickham, 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, \& Francois, 2018), \emph{readxl} (Version 1.3.1; Wickham \& Bryan, 2019), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version 3.0.3; Müller \& Wickham, 2020), \emph{tidyr} (Version 1.1.2; Wickham, 2020b), and \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al., 2019) for all our analyses.
 
@@
 
+\leavevmode\hypertarget{ref-R-janitor}{}%
+Firke, S. (2020). \emph{Janitor: Simple tools for examining and cleaning dirty data}. Retrieved from \url{https://CRAN.R-project.org/package=janitor}
+
+\leavevmode\hypertarget{ref-R-purrr}{}%
+Henry, L., \& Wickham, H. (2020). \emph{Purrr: Functional programming tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
+
+\leavevmode\hypertarget{ref-R-here}{}%
+Müller, K. (2017). \emph{Here: A simpler way to find your files}. Retrieved from \url{https://CRAN.R-project.org/package=here}
+
+\leavevmode\hypertarget{ref-R-tibble}{}%
+Müller, K., \& Wickham, H. (2020). \emph{Tibble: Simple data frames}. Retrieved from \url{https://CRAN.R-project.org/package=tibble}
+
 \leavevmode\hypertarget{ref-R-base}{}%
 R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
+
+\leavevmode\hypertarget{ref-R-ggplot2}{}%
+Wickham, H. (2016). \emph{Ggplot2: Elegant graphics for data analysis}. Springer-Verlag New York. Retrieved from \url{https://ggplot2.tidyverse.org}
+
+\leavevmode\hypertarget{ref-R-stringr}{}%
+Wickham, H. (2019). \emph{Stringr: Simple, consistent wrappers for common string operations}. Retrieved from \url{https://CRAN.R-project.org/package=stringr}
+
+\leavevmode\hypertarget{ref-R-forcats}{}%
+Wickham, H. (2020a). \emph{Forcats: Tools for working with categorical variables (factors)}. Retrieved from \url{https://CRAN.R-project.org/package=forcats}
+
+\leavevmode\hypertarget{ref-R-tidyr}{}%
+Wickham, H. (2020b). \emph{Tidyr: Tidy messy data}. Retrieved from \url{https://CRAN.R-project.org/package=tidyr}
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the tidyverse.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
+
+\leavevmode\hypertarget{ref-R-readxl}{}%
+Wickham, H., \& Bryan, J. (2019). \emph{Readxl: Read excel files}. Retrieved from \url{https://CRAN.R-project.org/package=readxl}
+
+\leavevmode\hypertarget{ref-R-dplyr}{}%
+Wickham, H., François, R., Henry, L., \& Müller, K. (2020). \emph{Dplyr: A grammar of data manipulation}. Retrieved from \url{https://CRAN.R-project.org/package=dplyr}
+
+\leavevmode\hypertarget{ref-R-readr}{}%
+Wickham, H., Hester, J., \& Francois, R. (2018). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
 

### 2020-11-25 scripts/draft_play.Rmd
Working on Endgame
Wrote in some hypotheses to match the RQs and started an another section for descriptive analysis of growth per category.

--- 2020-11-23 scripts/draft_play.Rmd
@@
 
-# Problem Statement and RQs
+# Problem Statement and Research Questions
 In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
@@
 (2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+
+## Hypotheses
+(1) We anticipate that prevalence rates of disability or disorder will not change for diagnostically rigid categories (e.g., Autism Spectrum Disorder) and may plateau or decrease for diagnostically malleable categories (e.g., mild behavioural need).
+(2) Allied professional resource support in the BC public education system points to challenges in rural settings.
+Therefore, we anticipate that the longitudinal provincial trend will largely be driven by urban school districts and rural districts may have highly variable outcomes post-2016 supreme court ruling.
 
@@
         y = "Headcount")
-#strength 1: I really like the data visualization. It is very clear! 
+
+```
+As observed in the figure above, low-incidence categories of need (e.g., deafblindness, visual impairment) remain constant over time.
+Both categorizations of intellectual disability saw a slight linear decline over time, while intensive behavioural need and learning disability categories had a slight linear increase over time.
+Chronic health impairment appeared to have a linear increase; though, no other category increased as rapidly as Autism throughout the more than two decades observed.
+
+``` {r growthrates}
+
+# library(growthrates)
+# 
+# growthrate1 <- BCdatastripped_spedcat %>% 
+#   filter(SPECIAL_NEEDS_CODE == "Autism")
+# 
+# all_easylinear(BCdatastripped_spedcat, grouping = "SPECIAL_NEEDS_CODE", time = "SCHOOL_YEAR", y= "NUMBER_OF_STUDENTS")
+
+ ggplot(BCdatastripped_spedcat, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+  geom_line(color = "cornflowerblue") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+   facet_wrap(~SPECIAL_NEEDS_CODE) +
+   labs(title = "BC Provincial Headcount of Students with Exceptionalities by Category",
+        x = "School Year",
+        y = "Headcount")
+
 ```
@@
 
-# strength 2: I really like the data cleaning part. They used pivot_longer to make the data more clear! 
 ```
@@
 
-#improvement: I am not sure what does the second ```r``` includes. maybe you could delete. 
 ```

### 2020-11-25 scripts/draft_play.Rmd

--- 2020-11-23 scripts/draft_play.Rmd
@@
 ---
-title             : "The title"
-shorttitle        : "Title"
+title             : "Trends in Pacific Northwest Special Education Population"
+shorttitle        : "PACNW SpEd Trends"
+
 author: 
-  - name          : "Makayla Whitney"
+  - name          : "David Fainstein"
     affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
+    address       : "1 David Ave, Eugene, Oregon, USA"
+    email         : "dfainste@uoregon.edu"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
       - Writing - Review & Editing
+
 affiliation:
@@
     institution   : "University of Oregon"
+
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-  Enter author note here.
+  Janette Avelar, University of Oregon College of Education
+  David Fainstein, University of Oregon College of Education
+  Joe Swinehart, Behavioral Research & Teaching
+  Makayla Whitney, University of Oregon College of Education
+
+  Project completed using RStudio, GitKraken, GitHub, and papaja
+
+  Contact: dfainste@uoregon.edu
 abstract: |
-  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
+  This paper will examine the changes in public school special education population in Vancouver, B.C. and Oregon over the last 20 years.
 
@@
   <!-- https://tinyurl.com/ybremelq -->
-keywords          : "keywords"
-wordcount         : "X"
+
+keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
+wordcount         : "0<n<1,000,000"
+
 bibliography      : ["r-references.bib"]
+
 floatsintext      : no
@@
 draft             : no
+
 documentclass     : "apa6"
@@
 output            : papaja::apa6_pdf
-editor_options: 
-  chunk_output_type: console
 ---
@@
 
-# Problem Statement and RQs
-In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
-While Canadian K-12 academic achievement outcomes are viewed as favourable on the world stage, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education.
-Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur.
-The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.
-
-With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
-(1) Do student prevalence rates of disability or disorder change after the supreme court ruling of 2016?
-(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
-
-# Results
-
+-- more introductory text here --
 Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
@@
 
-```{r sandbox, warning=FALSE}
+```{r sandbox}
 library(here)
@@
 
-BCdatastripped_spedcat <- BCdatastripped %>% 
+BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
+
+BCdatastripped <- BCdata %>% 
+  select("SCHOOL_YEAR", 
+         "DATA_LEVEL", 
+         "SPECIAL_NEEDS_CODE", 
+         "NUMBER_OF_STUDENTS", 
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER") %>% 
   mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
@@
          NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
-         SCHOOL_YEAR = as.integer(SCHOOL_YEAR)) %>% 
-  filter(SCHOOL_YEAR >= 7,
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
+  filter(DATA_LEVEL == "PROVINCE LEVEL",
+         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
+         SCHOOL_YEAR >= 7,
          SPECIAL_NEEDS_CODE != "M",
@@
 
- ggplot(BCdatastripped_spedcat, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+ ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
   geom_line(color = "cornflowerblue") +
@@
  facet_wrap(~disability)+
-    labs(color="frequency", x= "Year", y="Student Count", title = "Trends in Frequency")
+    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")
 
 ```
+
+# Methods
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+
+## Participants
+
+## Material
 

### 2020-11-28 scripts/j_play.Rmd
Added a table, made columns in urban/rural data numeric

--- 2020-11-25 scripts/draft_play.Rmd
@@
 ---
-title             : "Trends in Pacific Northwest Special Education Population"
-shorttitle        : "PACNW SpEd Trends"
-
-author: 
-  - name          : "David Fainstein"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "1 David Ave, Eugene, Oregon, USA"
-    email         : "dfainste@uoregon.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
-  - name          : "Janette Avelar"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
-  - name          : "Makayla Whitney"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
-  - name          : "Joe Swinehart"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
-
-affiliation:
-  - id            : "1"
-    institution   : "University of Oregon"
-
-authornote: |
-  Janette Avelar, University of Oregon College of Education
-  David Fainstein, University of Oregon College of Education
-  Joe Swinehart, Behavioral Research & Teaching
-  Makayla Whitney, University of Oregon College of Education
-
-  Project completed using RStudio, GitKraken, GitHub, and papaja
-
-  Contact: dfainste@uoregon.edu
-abstract: |
-  This paper will examine the changes in public school special education population in Vancouver, B.C. and Oregon over the last 20 years.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
-wordcount         : "0<n<1,000,000"
-
-bibliography      : ["r-references.bib"]
-
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
+title: "j_play"
+output: html_document
 ---
 
-```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 library("papaja")
-r_refs("r-references.bib")
 library(here)
@@
 
-BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
-
-BCdatastripped <- BCdata %>% 
-  select("SCHOOL_YEAR", 
-         "DATA_LEVEL", 
-         "SPECIAL_NEEDS_CODE", 
-         "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT") %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL", 
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
+#Put library calls you add for your file below here:
+library(kableExtra)
 ```
 
-# Introduction
-
-Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
-The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon.
-The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
-The OR head counts include years 2002-2020.
-Levels of categorization include school- district- and provincial-level head counts for BC.
-The OR data set includes state-wide head counts that are not aggregated by school or district.
-
-We intend to explore several questions regarding longitudinal trends.
-Firstly, we hope to analyze trends in disability prevalence over time.
-We will engage in a discussion on developmental trajectories by studying how trends shift from static/linear to increasing linear trends based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data.
-In studying the differences and similarities between the two datasets we will also engage in a discussion on diagnosis terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM), in part as a response to a challenge set by differing terminology between BC/OR categorization.
-
-Secondly, we hope to more closely analyze any changes, or lack thereof, within the BC data following the year 2016 during which a BC Supreme Court ruled in favor of limiting the number of special needs students in classrooms and expanding the number of specialist teachers schools are required to hire.
-
-Finally, we will also explore differences between urban and rural school districts in BC.
-Districts are categorized by population size and proximity to metropolitan areas, as recorded and defined by the Statistics Canada census from 2016.
-Two fringe rural districts exhibiting high populations will be picked out and  high-incidence diagnoses will be compared to those of other regions over time.
-
--- more introductory text here --
-Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
-The figure below demonstrates growth of 12 potential designations over an 18 year time period:
-
-```{r sandbox}
-library(here)
-library(tidyverse)
-library("readxl")
-
+```{r bc sandbox}
 BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
@@
         y = "Headcount")
-
 ```
 
-```{r urban vs. rural}
-
+```{r urban vs rural}
 bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))
@@
 
+#Attempting to make columns numeric
+view(district_class) #Need cols 6 & 8
+
+x <- c(6,8)
+district_class[ , x] <- apply(district_class[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
+
 ```
@@
 ``` {r District Census Population Table}
+# as_tibble(census_data)
+census_data <- district_class %>% 
+  janitor::clean_names() %>% 
+  group_by(district_name, district_number) %>% 
+  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c("school_year",
+            "data_level",
+            "public_or_independent",
+            "special_needs_code",
+            "number_of_students")) %>% 
+  distinct()
 
+apa_table(
+  census_data,
+    digits = 0,
+  caption = "District Classification by Population According to 2016 Census",
+  note = "Number of students is estimated and does not include masked data.",
+  col.names = c("District Number",
+            "District Name",
+            "Population",
+            "Classification",
+            "Number of Students")
+)
 ```
 
-This table displays the census results from 2011 and 2016 for our school districts.
-Many of the populations have stayed consistent within their urban or rural category.
-Three districts to note from the table are Nanaimo, Kamloops, and Chilliwack.
-In 2011, they were rural, but for our dataset they have been classified as urban due to their population increase in 2016.
-
-```{r cleaning Oregon data}
+```{r clean OR data}
 primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
@@
     labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
-
 ```
-
-```{r}
-#hiding this for now
-fresh_primo %>%
-filter(disability=="AUTISM") %>%
-
-ggplot(aes(age, frequency))+
- geom_line(aes(color=disability))+
-facet_wrap(~year)
-
-```
-```{r}
-#this is more in line with the BC graphing
-ggplot(fresh_primo, aes(year, frequency))+
- geom_smooth(color = "cornflowerblue") +
- theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
- facet_wrap(~disability)+
-    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")
-
-```
-
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Participants
-
-## Material
-
-## Procedure
-
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
-
-# Results
-
-# Discussion
-
-\newpage
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup

### 2020-11-28 scripts/draft_play.Rmd
Added chunks and numericalized numbers.

@@
 
+ # Making student number column numeric
+x <- c(10)
+BCdata[ , x] <- apply(BCdata[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
+#
+```
+
+```{r table: headcount by year}
+#This is a work in progress no touchy
+
+#Data View
+#view(BCdatastripped)
+#view(BCdata)
+#as_tibble(count_by_year)
+
+#Data Prep
+count_by_year <- BCdata %>% 
+  janitor::clean_names() %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select("school_year",
+         "special_needs_desc",
+         "total_students") %>% 
+  drop_na()
+
+#Table
+apa_table(
+  count_by_year,
+  digits = 0,
+  caption = "Headcount by Classification and District, Select Years 1996-2020",
+  note = "Number of students is estimated and does not include masked data.",
+  col.names = c("School Year",
+                "Special Needs Description",
+                "Number of Students")
+)
+
+#Table is huge - pivot_wider by year and maybe select specific years to look at
+
+unique(count_by_year$school_year)
 ```
@@
 
+#Attempting to make columns numeric
+view(district_class) #Need cols 6 & 8
+
+x <- c(6,8)
+district_class[ , x] <- apply(district_class[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
+#
 ```
@@
 ``` {r District Census Population Table}
+#Data Prep
+census_data <- district_class %>% 
+  janitor::clean_names() %>% 
+  group_by(district_name, district_number) %>% 
+  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c("school_year",
+            "data_level",
+            "public_or_independent",
+            "special_needs_code",
+            "number_of_students")) %>% 
+  distinct()
 
+#Table
+apa_table(
+  census_data,
+    digits = 0,
+  caption = "District Classification by Population According to 2016 Census",
+  note = "Number of students is estimated and does not include masked data.",
+  col.names = c("District Number",
+            "District Name",
+            "Population",
+            "Classification",
+            "Number of Students")
+)
 ```

### 2020-11-28 scripts/j_play.Rmd
Added chunks and numericalized numbers.

@@
 #Put library calls you add for your file below here:
-library(kableExtra)
 ```
@@
 
+# Making student number column numeric
+x <- c(10)
+BCdata[ , x] <- apply(BCdata[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
+#
+
  ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
@@
         y = "Headcount")
+```
+
+```{r table: headcount by year}
+#Data View
+#view(BCdatastripped)
+#view(BCdata)
+#as_tibble(count_by_year)
+
+#Data Prep
+count_by_year <- BCdata %>% 
+  janitor::clean_names() %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select("school_year",
+         "special_needs_desc",
+         "total_students") %>% 
+  drop_na()
+
+#Table
+apa_table(
+  count_by_year,
+  digits = 0,
+  caption = "Headcount by Classification and District, Select Years 1996-2020",
+  note = "Number of students is estimated and does not include masked data.",
+  col.names = c("School Year",
+                "Special Needs Description",
+                "Number of Students")
+)
+
+#Table is huge - pivot_wider by year and maybe select specific years to look at
+
+unique(count_by_year$school_year)
 ```
@@
                                function(x)as.numeric(as.character(x)))
-
+#
 ```
@@
 ``` {r District Census Population Table}
-# as_tibble(census_data)
+#Data Prep
 census_data <- district_class %>% 
@@
 
+#Table
 apa_table(

### 2020-11-28 scripts/draft_play.Rmd
Hid table code so it doesn't kill your computers

@@
 #Table
-apa_table(
-  count_by_year,
-  digits = 0,
-  caption = "Headcount by Classification and District, Select Years 1996-2020",
-  note = "Number of students is estimated and does not include masked data.",
-  col.names = c("School Year",
-                "Special Needs Description",
-                "Number of Students")
-)
+#apa_table(
+#  count_by_year,
+#  digits = 0,
+#  caption = "Headcount by Classification and District, Select Years 1996-2020",
+#  note = "Number of students is estimated and does not include masked data.",
+#  col.names = c("School Year",
+#                "Special Needs Description",
+#                "Number of Students")
+#)
 
-#Table is huge - pivot_wider by year and maybe select specific years to look at
+#Table is huuuge - do not run table code yet or your computer will be frozen for 5 minutes.
 

### 2020-11-29 scripts/BCOR_combo.Rmd
Big Clean
Cleaned up my data to Joe's suggestions (added years, readin new file, cleaned population data, added in prevalence rates) and started on a BC/OR comparison

--- 2020-11-25 scripts/draft_play.Rmd
@@
 
+BCpopulation <- read_excel(here("data", "BCpopdata.xlsx")) %>% 
+  select("SCHOOL_YEAR",
+         "TOTAL_ENROLMENT",
+         "PUBLIC_OR_INDEPENDENT") %>% 
+  filter(PUBLIC_OR_INDEPENDENT == "BC PUBLIC SCHOOL") %>% 
+  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)
+
 BCdatastripped <- BCdata %>% 
+    #Makayla & Janette - may want to include more columns here with the select for our global "BCdatastripped" df
     select("SCHOOL_YEAR", 
@@
          "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT") %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL", 
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER") %>% 
+  filter(SPECIAL_NEEDS_CODE != "M",
+         SPECIAL_NEEDS_CODE != "J",
+         SPECIAL_NEEDS_CODE != "N") %>% 
+    mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
+                          "Physically Dependent" = "A",
+                          "Deafblind" = "B",
+                          "Moderate+ ID" = "C",
+                          "Chronic Health Impairment" = "D",
+                          "Visual Impairment" = "E",
+                          "Deaf or Hard of Hearing" = "F",
+                          "Autism" = "G",
+                          "Intensive Behavioural Need" = "H",
+                          "Mild ID" = "K",
+                          "Learning Disability" = "Q",
+                          "Gifted" = "P",
+                          "Mild Behavioural Need" = "R"),
+           SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
+           NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
+           SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
+  filter(SCHOOL_YEAR >= 9)
+
 ```
@@
 Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
-The figure below demonstrates growth of 12 potential designations over an 18 year time period:
+The figure below demonstrates growth of 12 potential designations over 15 years:
 
-```{r sandbox}
-library(here)
-library(tidyverse)
-library("readxl")
+```{r Provinical_Data_Overview}
 
-BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
+BCprovincial <- BCdatastripped %>% 
+  filter(DATA_LEVEL == "PROVINCE LEVEL",
+         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")  %>% 
+  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
+         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
+  mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995)
 
-BCdatastripped <- BCdata %>% 
-  select("SCHOOL_YEAR", 
-         "DATA_LEVEL", 
-         "SPECIAL_NEEDS_CODE", 
-         "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT",
-         "DISTRICT_NUMBER") %>% 
-  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
-                          "Physically Dependent" = "A",
-                          "Deafblind" = "B",
-                          "Moderate+ ID" = "C",
-                          "Chronic Health Impairment" = "D",
-                          "Visual Impairment" = "E",
-                          "Deaf or Hard of Hearing" = "F",
-                          "Autism" = "G",
-                          "Intensive Behavioural Need" = "H",
-                          "Mild ID" = "K",
-                          "Learning Disability" = "Q",
-                          "Gifted" = "P",
-                          "Mild Behavioural Need" = "R"),
-         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
-         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
-         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL",
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
-         SCHOOL_YEAR >= 7,
-         SPECIAL_NEEDS_CODE != "M",
-         SPECIAL_NEEDS_CODE != "J",
-         SPECIAL_NEEDS_CODE != "N")
-
- ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+ ggplot(BCprovincial, aes(x = SCHOOL_YEAR, y = prevalence)) +
   geom_line(color = "cornflowerblue") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(~SPECIAL_NEEDS_CODE) +
-   labs(title = "BC Provincial Headcount of Students with Exceptionalities by Category",
+   labs(title = "BC Provincial Prevalence of Students with Exceptionalities by Category",
         x = "School Year",
-        y = "Headcount")
+        y = "Prevalence Rate")
 
@@
 
+``` {r BCOR combo}
+
+combined <- fresh_primo %>% 
+  as_tibble() %>% 
+    mutate(ORheadcount = fct_recode(disability,
+                          "Deafblind" = "DEAF-BLINDNESS",
+                          "Moderate+ ID" = "INTELLECTUAL DISABILITY",
+                          "Chronic Health Impairment" = "EMOTIONAL DISTURBANCE",
+                          "Visual Impairment" = "VISUAL IMPAIRMENT",
+                          "Deaf or Hard of Hearing" = "HEARING IMPAIRMENT",
+                          "Autism" = "AUTISM",
+                          "Mild ID" = "INTELLECTUAL DISABILITY")) %>% 
+  filter(ORheadcount == "Autism") %>% 
+  group_by(year) %>% 
+  summarise(frequency = sum(frequency))
+
+BCAutism <- BCprovincial %>% 
+  filter(SPECIAL_NEEDS_CODE == "Autism")
+
+# BCprovincial <- BCdatastripped %>% 
+#   filter(DATA_LEVEL == "PROVINCE LEVEL",
+#          PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")  %>% 
+#   mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
+#          prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
+#   mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995)
+
+ggplot() +
+  geom_smooth(data = combined, aes(x = year, y = frequency), color = "cornflowerblue") +
+  geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS), color = "red") +
+  labs(title = "Visualizing BC and Oregon Headcount of Autism Spectrum Disorder",
+        x = "School Year",
+        y = "Headcount")
+
+```
+
 # Methods

### 2020-11-30 scripts/draft_play.Rmd
Urban/Rural exploration chunks

@@
 
+```{r rural play}
+#Data Prep
+rural_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Rural") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+#Viz - Totals by Code
+ggplot(rural_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality",
+       x = "Exceptionality",
+       y = "Number of Students")
+
+#Data for Rural Totals (No categories)
+rural_totals <- rural_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#Data Viz Totals/Time
+ggplot(rural_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+#What's going on with Prince George? Why is there such a steep drop in 2006?
+pg <- rural_districts %>% 
+  filter(district_name == "Prince George") %>% 
+  as_tibble()
+
+pg_sums <- pg %>% 
+  group_by(school_year) %>% 
+  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#What's happening by Dx?
+pg_dx_sum <- pg %>% 
+  group_by(school_year, special_needs_code) %>% 
+  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(district_number, district_classification, number_of_students)) %>% 
+  distinct()
+
+ggplot(pg_dx_sum, aes(school_year, student)) +
+  geom_col() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  facet_wrap(~special_needs_code)
+
+#What's going on with gifted?????
+pg_gifted <- pg %>% 
+  filter(special_needs_code == "P") %>% 
+  select(-c(district_number, district_classification))
+
+ggplot(pg_gifted, aes(school_year, number_of_students)) +
+  geom_col() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+```{r urban play}
+#Data Prep
+urban_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Urban") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+urban_totals <- urban_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#Data Viz Totals/Time
+ggplot(urban_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+#Viz - Totals by Code
+ggplot(urban_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality",
+       x = "Exceptionality",
+       y = "Number of Students")
+```
+
 ```{r cleaning Oregon data}

### 2020-11-30 scripts/j_play.Rmd
Urban/Rural exploration chunks

@@
 
+count_by_year <- BCdata %>% 
+  janitor::clean_names() %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select("school_year",
+         "special_needs_desc",
+         "total_students") %>% 
+  pivot_wider(names_from = school_year,
+              values_from = total_students) %>% 
+  drop_na()
+
 #Table
@@
 
+```{r rural play}
+#Data Prep
+rural_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Rural") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+#Viz - Totals by Code
+ggplot(rural_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality",
+       x = "Exceptionality",
+       y = "Number of Students") +
+  scale_color_manual(name = "Special Needs Description",
+                    labels = c("A - Physically Dependent",
+                               "B - Deafblind",
+                               "C - Moderate to Profound Intellectual Disability",
+                               "D - Physical Disability or Chronic Health Impairment",
+                               "E - Visual Impairment",
+                               "F - Deaf or Hard of Hearing",
+                               "G - Autism Spectrum Disorder",
+                               "H - Inten Behav Inter/Serious Mental Illness",
+                               "J - Severe Learning Disabilities",
+                               "K - Mild Intellectual Disability",
+                               "M - Behavior Disorder (Moderate)",
+                               "N - Behavior Disorder (Rehabilitation)",
+                               "P - Gifted",
+                               "Q - Learning Disability",
+                               "R - Moderate Behaviour Support/Mental Illness"))
+
+rlang::last_error()
+```
+
+```{r codes and descs}
+code_key <- bc_data_district_classification %>% 
+  janitor::clean_names() %>% 
+  select(special_needs_code, special_needs_desc) %>% 
+  distinct()
+```
+
 ```{r clean OR data}

### 2020-12-02 scripts/draft_play.Rmd
Work in Progress - abstract & Sooke case study

@@
 abstract: |
-  This paper will examine the changes in public school special education population in Vancouver, B.C. and Oregon over the last 20 years.
+  This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
+The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministrty of Education in British Columbia.
+This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
+The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020.
+This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year.
+Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
 
@@
 library(janitor)
+library(dplyr)
 
@@
 Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
-The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon.
+The datasets detail the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon.
 The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
@@
     theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+##Looking into Sooke (SD 62 - pop. 9,269) categorized under Victoria so urban but technically rural
+## In progress
+
+sooke <- district_class %>%
+  filter(DISTRICT_NAME == "Sooke") %>%
+  as_tibble()
+
+ggplot(sooke, aes(SPECIAL_NEEDS_CODE)) +
+  facet_wrap(~SCHOOL_YEAR)
+  geom_col()
 ```

### 2020-12-02 scripts/draft_play.Rmd
Added some labels

@@
 Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
-The datasets detail the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon.
+The datasets detail the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon (OR).
 The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
@@
 Firstly, we hope to analyze trends in disability prevalence over time.
-We will engage in a discussion on developmental trajectories by studying how trends shift from static/linear to increasing linear trends based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data.
+We will engage in a discussion on developmental trajectories by studying how trends shift from static/linear to increasing linear trends based on age of diagnosis for the OR data, which can serve as a springboard to make inferences about BC data.
 In studying the differences and similarities between the two datasets we will also engage in a discussion on diagnosis terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM), in part as a response to a challenge set by differing terminology between BC/OR categorization.
@@
 #Attempting to make columns numeric
-view(district_class) #Need cols 6 & 8
+#view(district_class) #Need cols 6 & 8
 
@@
   geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
 
@@
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  facet_wrap(~special_needs_code)
+  facet_wrap(~special_needs_code) +
+  labs(title = "Students with Exceptionalities in Prince George 1999-2020",
+       x = "School Year",
+       y = "Number of Students")
+# Need some explanation here for explaining facet wrap
 
@@
   geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Gifted Students in Prince George 1999-2020",
+       subtitle = "Note: Totals under 10 are masked by district in reporting.",
+       x = "School Year",
+       y = "Total Students")
 
@@
   geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
 

### 2020-12-02 scripts/BCOR_combo.Rmd
Effect Size - First Structure is there!

@@
 library(janitor)
+library(esc)
 
@@
                           "Moderate+ ID" = "C",
-                          "Chronic Health Impairment" = "D",
+                          "Chronic Health" = "D",
                           "Visual Impairment" = "E",
@@
                           "Autism" = "G",
-                          "Intensive Behavioural Need" = "H",
+                          "High Behavioural Need" = "H",
                           "Mild ID" = "K",
@@
                           "Moderate+ ID" = "INTELLECTUAL DISABILITY",
-                          "Chronic Health Impairment" = "EMOTIONAL DISTURBANCE",
+                          "High Behavioural Need" = "EMOTIONAL DISTURBANCE",
                           "Visual Impairment" = "VISUAL IMPAIRMENT",
@@
                           "Mild ID" = "INTELLECTUAL DISABILITY")) %>% 
-  filter(ORheadcount == "Autism") %>% 
+  filter(ORheadcount == "High Behavioural Need") %>% 
   group_by(year) %>% 
@@
 BCAutism <- BCprovincial %>% 
-  filter(SPECIAL_NEEDS_CODE == "Autism")
-
-# BCprovincial <- BCdatastripped %>% 
-#   filter(DATA_LEVEL == "PROVINCE LEVEL",
-#          PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")  %>% 
-#   mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
-#          prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
-#   mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995)
+  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need")
 
 ggplot() +
-  geom_smooth(data = combined, aes(x = year, y = frequency), color = "cornflowerblue") +
-  geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS), color = "red") +
-  labs(title = "Visualizing BC and Oregon Headcount of Autism Spectrum Disorder",
+  geom_smooth(data = combined, aes(x = year, y = frequency, color = "Oregon")) +
+  geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS, color = "British Columbia"), se = FALSE) +
+  labs(title = "BC and Oregon Headcount of Students with Chronic Health Conditions",
         x = "School Year",
-        y = "Headcount")
+        y = "Headcount",
+       color = "Legend",
+        scale_color_manual(labels = c("Oregon", "British Columbia"), values = c("blue", "red")))
+
+```
+
+``` {r calculatingES}
+
+ESdata <- BCdatastripped %>% 
+  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need",
+         DATA_LEVEL == "DISTRICT LEVEL",
+         PUBLIC_OR_INDEPENDENT == "BC Public School",
+         DISTRICT_NUMBER == "036")  %>% 
+  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
+         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
+  mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995,
+         pre16 = ifelse(SCHOOL_YEAR <= 2016, "before16", "after16")) %>%
+  group_by(pre16) %>% 
+  pivot_wider(
+    names_from = pre16,
+    values_from = prevalence
+  )
+
+presup <- ESdata %>% 
+  drop_na(before16) %>% 
+  select(before16)
+as.numeric(unlist(presup))
+
+postsup <- ESdata %>% 
+  drop_na(after16) %>% 
+  select(after16)
+as.numeric(unlist(postsup))
+
+presup_mean <- sapply(presup, mean)
+presup_sd <- sapply(presup, sd)
+postsup_mean <- sapply(postsup, mean)
+postsup_sd <- sapply(postsup, sd)
+
+esc_mean_sd(grp1m = presup_mean, grp1sd = presup_sd, grp1n = 13, grp2m = postsup_mean, grp2sd = postsup_sd, grp2n = 3, es.type = "g")
 

### 2020-12-02 scripts/j_play.Rmd

--- 2020-11-28 scripts/j_play.Rmd
@@
 
+count_by_year <- BCdata %>% 
+  janitor::clean_names() %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select("school_year",
+         "special_needs_desc",
+         "total_students") %>% 
+  pivot_wider(names_from = school_year,
+              values_from = total_students) %>% 
+  drop_na()
+
 #Table
-apa_table(
-  count_by_year,
-  digits = 0,
-  caption = "Headcount by Classification and District, Select Years 1996-2020",
-  note = "Number of students is estimated and does not include masked data.",
-  col.names = c("School Year",
-                "Special Needs Description",
-                "Number of Students")
-)
+#apa_table(
+#  count_by_year,
+#  digits = 0,
+#  caption = "Headcount by Classification and District, Select Years 1996-2020",
+#  note = "Number of students is estimated and does not include masked data.",
+#  col.names = c("School Year",
+#                "Special Needs Description",
+#                "Number of Students")
+#)
 
@@
 
-unique(count_by_year$school_year)
+#unique(count_by_year$school_year)
 ```
@@
   caption = "District Classification by Population According to 2016 Census",
-  note = "Number of students is estimated and does not include masked data.",
+  note = "Number of students does not include masked data.",
   col.names = c("District Number",
@@
 )
+```
+
+```{r rural play}
+#Data Prep
+rural_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Rural") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+#Viz - Totals by Code
+ggplot(rural_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality",
+       x = "Exceptionality",
+       y = "Number of Students")
+
+#Data for Rural Totals (NO categories)
+rural_totals <- rural_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#Data Viz Totals/Time
+ggplot(rural_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+#What's going on with Prince George? Why is there such a steep drop in 2006?
+pg <- rural_districts %>% 
+  filter(district_name == "Prince George") %>% 
+  as_tibble()
+
+pg_sums <- pg %>% 
+  group_by(school_year) %>% 
+  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#What's happening by Dx?
+pg_dx_sum <- pg %>% 
+  group_by(school_year, special_needs_code) %>% 
+  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(district_number, district_classification, number_of_students)) %>% 
+  distinct()
+
+ggplot(pg_dx_sum, aes(school_year, student)) +
+  geom_col() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  facet_wrap(~special_needs_code)
+
+#What's going on with gifted?????
+pg_gifted <- pg %>% 
+  filter(special_needs_code == "P") %>% 
+  select(-c(district_number, district_classification))
+
+ggplot(pg_gifted, aes(school_year, number_of_students)) +
+  geom_col() +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+```{r urban play}
+#Data Prep
+urban_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Urban") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+urban_totals <- urban_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#Data Viz Totals/Time
+ggplot(urban_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+#Viz - Totals by Code
+ggplot(urban_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality",
+       x = "Exceptionality",
+       y = "Number of Students")
+
+```
+
+```{r ASD}
+asd <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(special_needs_code == "G")
+```
+
+```{r codes and descs}
+code_key <- bc_data_district_classification %>% 
+  janitor::clean_names() %>% 
+  select(special_needs_code, special_needs_desc) %>% 
+  distinct()
 ```

### 2020-12-03 scripts/BCOR_combo.Rmd
Trying to Update Again
ES data calculated for Rural vs. Urban Settings

@@
 
+urban <- c("023", "034", "035", "036", "037", "038", "039", "040", "041", "042", "043", "044", "045", "046")
+
 ESdata <- BCdatastripped %>% 
-  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need",
+  filter(SPECIAL_NEEDS_CODE == "Chronic Health",
          DATA_LEVEL == "DISTRICT LEVEL",
          PUBLIC_OR_INDEPENDENT == "BC Public School",
-         DISTRICT_NUMBER == "036")  %>% 
+         !DISTRICT_NUMBER %in% urban)  %>% 
   mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
@@
 as.numeric(unlist(presup))
+presup_count <- count(presup)
 
@@
 as.numeric(unlist(postsup))
+postsup_count <- count(postsup)
 
@@
 
-esc_mean_sd(grp1m = presup_mean, grp1sd = presup_sd, grp1n = 13, grp2m = postsup_mean, grp2sd = postsup_sd, grp2n = 3, es.type = "g")
+esc_mean_sd(grp1m = presup_mean, grp1sd = presup_sd, grp1n = presup_count, grp2m = postsup_mean, grp2sd = postsup_sd, grp2n = postsup_count, es.type = "g")
 

### 2020-12-04 scripts/BCOR_combo.Rmd
Updated Data, Tidying, and Prevalence Calculations for ES

@@
 
+BCpop_district <- read_excel(here("data", "BCdistrictpop.xlsx")) %>% 
+  select("SCHOOL_YEAR",
+         "TOTAL_ENROLMENT",
+         "DISTRICT_NUMBER") %>% 
+  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)
+
 BCdatastripped <- BCdata %>% 
@@
   filter(SCHOOL_YEAR >= 9)
+
+BC_districtdata <- BCdatastripped %>% 
+  filter(DATA_LEVEL == "DISTRICT LEVEL")
+Prevalence_data <- inner_join(BCpop_district, BC_districtdata, by = c("SCHOOL_YEAR", "DISTRICT_NUMBER"), all.x = TRUE)
 
@@
 
-ESdata <- BCdatastripped %>% 
-  filter(SPECIAL_NEEDS_CODE == "Chronic Health",
-         DATA_LEVEL == "DISTRICT LEVEL",
-         PUBLIC_OR_INDEPENDENT == "BC Public School",
+ESdata <- Prevalence_data %>% 
+  filter(SPECIAL_NEEDS_CODE == "Mild Behavioural Need",
          !DISTRICT_NUMBER %in% urban)  %>% 
-  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
-         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
+  mutate(prevalence = NUMBER_OF_STUDENTS/TOTAL_ENROLMENT) %>% 
   mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995,
@@
   )
+
+# 1) group_by for all students with SPED, then run that against public population to sort out overall prevelance of SPED
+# 2)
 

### 2020-12-05 scripts/Final_project.Rmd
Final Project Doc for Combining our Work

---
title             : "Exploring Anomalies in Pacific Northwest Special Education Trends"
shorttitle        : "PACNW SpEd Trends"

author: 
  - name          : "David Fainstein"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "1655 Alder, Eugene, Oregon, USA  97403"
    email         : "dfainste@uoregon.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Janette Avelar"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Makayla Whitney"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
  - name          : "Joe Swinehart"
    affiliation   : "1"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "University of Oregon"

authornote: |
  Janette Avelar, University of Oregon College of Education
  David Fainstein, University of Oregon College of Education
  Joe Swinehart, Behavioral Research & Teaching
  Makayla Whitney, University of Oregon College of Education

  Project completed using RStudio, GitKraken, GitHub, and papaja

  Contact: dfainste@uoregon.edu
abstract: |
  This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years. The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministrty of Education in British Columbia. This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year. The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020. This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year. Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
wordcount         : "0<n<1,000,000"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
library("papaja")
r_refs("r-references.bib")
library(here)
library(tidyverse)
library("readxl")
library(janitor)
library(esc)

BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCpopulation <- read_excel(here("data", "BCpopdata.xlsx")) %>% 
  select("SCHOOL_YEAR",
         "TOTAL_ENROLMENT",
         "PUBLIC_OR_INDEPENDENT") %>% 
  filter(PUBLIC_OR_INDEPENDENT == "BC PUBLIC SCHOOL") %>% 
  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)

BCpop_district <- read_excel(here("data", "BCdistrictpop.xlsx")) %>% 
  select("SCHOOL_YEAR",
         "TOTAL_ENROLMENT",
         "DISTRICT_NUMBER") %>% 
  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)

BCdatastripped <- BCdata %>% 
    #Makayla & Janette - may want to include more columns here with the select for our global "BCdatastripped" df
    select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT",
         "DISTRICT_NUMBER") %>% 
  filter(SPECIAL_NEEDS_CODE != "M",
         SPECIAL_NEEDS_CODE != "J",
         SPECIAL_NEEDS_CODE != "N") %>% 
    mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
                          "Physically Dependent" = "A",
                          "Deafblind" = "B",
                          "Moderate+ ID" = "C",
                          "Chronic Health" = "D",
                          "Visual Impairment" = "E",
                          "Deaf or Hard of Hearing" = "F",
                          "Autism" = "G",
                          "High Behavioural Need" = "H",
                          "Mild ID" = "K",
                          "Learning Disability" = "Q",
                          "Gifted" = "P",
                          "Mild Behavioural Need" = "R"),
           SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
           NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
           SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
  filter(SCHOOL_YEAR >= 9)

BC_districtdata <- BCdatastripped %>% 
  filter(DATA_LEVEL == "DISTRICT LEVEL")
Prevalence_data <- inner_join(BCpop_district, BC_districtdata, by = c("SCHOOL_YEAR", "DISTRICT_NUMBER"), all.x = TRUE)

```

# Introduction
Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society (Ball, 1998). Educational policy exists as a mechanism to drive changes in school buildings that result in socially important outcomes (e.g., increasing literacy acquisition through improved instruction, reducing instances of bullying through prevention programs). However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical (Cohen et al., 2009). Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students (Swanson, 2007).

When policy is enacted, the intended outcomes are often not realized for months, years, or decades (Gu et al., 2018). For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke & McArdle, 2009). Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore & Dynarski, 2017).

##Current Study
Our project primarily examines two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21. The datasets offer insights to the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon (OR). The head counts from BC are collected from 2002/2003 to the most recent data from 2019/2020. The OR head counts are drawn from an annual dataset spanning 2002 through 2020. Levels of categorization include school-, district-, and provincial-level head counts for BC. The OR data set includes state-wide head counts that are not aggregated by school or district.

We intend to explore several questions regarding longitudinal trends where educational policy may have impacted special education practice. First, we will use descriptive statistics to describe trends in disability prevalence over time. The initial analyses will inform a discussion on developmental trajectories by illustrating shifts in prevalence growth from static to linearly increasing based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data (e.g., common age of onset for diagnosis per category and diagnostic patterns over time). In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM; American Psychological Association, 2013), in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).

Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017). Canadian supreme court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire. The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change. Theoretically, school district administrators made decisions based on current staffing levels, student designations, diagnostic capacity and malleability, as well as external pressures to adjust practice according to legal mandates.

Finally, we will explore differences between urban and rural school districts in BC. Districts are categorized by population size and proximity to metropolitan areas, as recorded and defined by the Statistics Canada census from 2016. School district level data is useful for examining longitudinal trends for individual administrative units (e.g., superintendent’s office and director(s) of inclusive education); however, this level of data analysis cannot speak to province-wide trends in the culture of superordinate legal directives which impact swift policy changes. Given this, we will run analyses on both exemplar districts to highlight localized policy impacts on practice and amalgamated districts (e.g., school districts defined as rural by Statistics Canada) to investigate cultural shifts in the policy-to-practice association.

##Problem Statement and RQs
In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013). While Canadian K-12 academic achievement outcomes are generally higher than a national-level comparison with the US, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education. Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur. The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.

With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
(1) Do student prevalence rates of disability or disorder decrease after the supreme court ruling of 2016 against existing trends?
(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?

#Methods

Datasets were extracted from publicly available sources. Specifically, the BC Ministry of Education, Oregon Department of Education, and Statistics Canada are the data sources for the current study. Note that the primary jurisdiction of interest is BC, and OR is selected as a reasonable basis of comparison (i.e., similar geography, population, and population growth rate). Further, BC has distinct educational policy from OR as Canadian education is administered entirely by provincial ministries of education and special education law is remarkably different between BC and OR. Together, these datasets help achieve three objectives for acquiring data appropriate to the research questions:
Headcount data at multiple levels (e.g., district and state or province) for students with exceptionalities, disaggregated by category of exceptionality, year, and location.
Population data to fit school district locations to a geographical classification (e.g., urban, suburban, rural) as specified in national documentation (e.g., urban centers have a minimum of 10,000 inhabitants based on census data), and calculate prevalence rates.
Data from multiple jurisdictions where educational policies do not align chronologically. For example, special education law in the United States is disjointed from BC special education law and policy changes are not the same, nor are they implemented simultaneously.

The nature of the current study is intentionally exploratory within a systematic approach of viewing data, testing a priori hypotheses against descriptive statistics, calculating effect sizes if appropriate, and, finally, exploring notable anomalies in prevalence trends that were not hypothesized a posteriori. As a first step, the authors took the approach of examining provincial-level data in BC. Then, trends in BC data were compared against trends in OR data to see if similar rates of students were designated with exceptionalities across jurisdictions. When anomalies were suspected in the BC trends of exceptionality prevalence, then effect size tests were conducted in line with the research questions. Finally, any anomalies that were suspected a posteriori were tested with effect size calculations and discussed in the final section of our current study.

The three datasets were used for analysis in R (cite R) for reproducibility and data visualization. All code for analysis and data representation is publicly available for reproducibility (https://github.com/JoeSwinehart/datascienceproject). There are several caveats for data use and cleaning in the current analysis. First, the BC and OR datasets were not a completely parallel comparison, because the OR dataset includes a broader age range of students (ages 5 - 21) as compared to the BC dataset (ages 5 - 19). Second, data masking necessitates a decision-point for analysis. Specifically, the BC Ministry of Education data masks all cells with headcounts below 10 students, therefore our analyses left out these cases when computing effect sizes. Data masking was most prominently an issue when examining district-level data in BC, as smaller school districts often masked low-incidence exceptionality data. When appropriate, masked data is assigned a convservative estimate of (n = 9) such that enough data units were included in effect size comparisons. Third, Canadian census data from 2016 is the basis of classification for district geographical location (e.g., urban, suburban, rural, exurban). Due to the geographical distribution of population in BC, only two categories are used for the current analyses: urban and rural.

##Analysis Plan
The majority of the analyses for the current study are descriptive in nature. The descriptive analysis includes a table or visual representation of prevalence rates or headcounts of students with exceptionalities over time. As the current study is examining anomalies in trend data, trends that break the longitudinal trendline (e.g., sudden drop in prevalence after a decade of linear increase in prevalence) may be examined further. That is, effect sizes between two groups (e.g., group one is three years before a policy change, and group two is three years after a policy change) may be examined to detect the magnitude of differences between groups if they are hypothesized to represent a change in practice stemming from policy implementation. More specifically on the calculation of effect sizes, the Hedge’s g statistic will be used because of the robustness against unequal group sizes.

$$
Hedges' g = \frac{M_1 + M_2}{SD_{pooled}} 
$$
#Results
Writing on results here...

First, we see that the BC trends do not show significant changes after the 2016 supreme court ruling.
```{r Provinical_Data_Overview}

BCprovincial <- BCdatastripped %>% 
  filter(DATA_LEVEL == "PROVINCE LEVEL",
         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")  %>% 
  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
  mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995)

 ggplot(BCprovincial, aes(x = SCHOOL_YEAR, y = prevalence)) +
  geom_line(color = "cornflowerblue") +
   facet_wrap(~SPECIAL_NEEDS_CODE) +
   labs(title = "BC Provincial Prevalence of Students with Exceptionalities by Category",
        x = "School Year",
        y = "Prevalence Rate")
```

```{r cleaning Oregon data}
primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
cleanest_primo <-primodata %>% 
  remove_empty(which = c("rows", "cols")) %>% 
  clean_names()

cleanest_primo <- cleanest_primo %>% 
  select(-six_to_twentyone_total, -x24)

cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
cleanest_primo2

fresh_primo <- cleanest_primo2 %>% 
  pivot_longer(starts_with("x"), names_to = "age", values_to = "frequency") %>% 
  mutate(age = parse_number(age), 
         year = parse_number(year))

fresh_primo %>% 
  group_by(year, disability) %>% 
  summarize(total=sum(frequency, na.rm=TRUE))

```

This is roughly in line with Oregon's designation data, showing that there were no state-wide anomalies after the year 2016.
```{r}
#this is more in line with the BC graphing
ggplot(fresh_primo, aes(year, frequency))+
 geom_smooth(color = "cornflowerblue") +
 theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
 facet_wrap(~disability)+
    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")

```

Even for diagnostically malleable categories, such as students classified in Oregon has having an "emotional disturbance," or in BC students requiring a "high behavioural need", Oregon and BC had highly similar trends.

``` {r BCOR combo}

combined <- fresh_primo %>% 
  as_tibble() %>% 
    mutate(ORheadcount = fct_recode(disability,
                          "Deafblind" = "DEAF-BLINDNESS",
                          "Moderate+ ID" = "INTELLECTUAL DISABILITY",
                          "High Behavioural Need" = "EMOTIONAL DISTURBANCE",
                          "Visual Impairment" = "VISUAL IMPAIRMENT",
                          "Deaf or Hard of Hearing" = "HEARING IMPAIRMENT",
                          "Autism" = "AUTISM",
                          "Mild ID" = "INTELLECTUAL DISABILITY")) %>% 
  filter(ORheadcount == "High Behavioural Need") %>% 
  group_by(year) %>% 
  summarise(frequency = sum(frequency))

BCAutism <- BCprovincial %>% 
  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need")

ggplot() +
  geom_smooth(data = combined, aes(x = year, y = frequency, color = "Oregon")) +
  geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS, color = "British Columbia"), se = FALSE) +
  labs(title = "BC and Oregon Headcount of Students with Chronic Health Conditions",
        x = "School Year",
        y = "Headcount",
       color = "Legend",
        scale_color_manual(labels = c("Oregon", "British Columbia"), values = c("blue", "red")))

```

Then, we examine the rural versus urban trends and notice a few things...

```{r urban vs. rural}

bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))

district_class <- bc_data_district_classification %>%
  select("SCHOOL_YEAR",
         "DATA_LEVEL",
         "PUBLIC_OR_INDEPENDENT",
         "DISTRICT_NUMBER",
         "DISTRICT_NAME",
         "DISTRICT_POPULATION_2016",
         "DISTRICT_CLASSIFICATION",
         "SPECIAL_NEEDS_CODE",
         "NUMBER_OF_STUDENTS") %>%
  filter(DATA_LEVEL == "DISTRICT LEVEL",
         PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
    select(-"name")

#Attempting to make columns numeric
#view(district_class) #Need cols 6 & 8

x <- c(6,8)
district_class[ , x] <- apply(district_class[ , x], 2,
                               function(x)as.numeric(as.character(x)))
#
```

``` {r District Census Population Table}
#Data Prep
census_data <- district_class %>% 
  janitor::clean_names() %>% 
  group_by(district_name, district_number) %>% 
  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c("school_year",
            "data_level",
            "public_or_independent",
            "special_needs_code",
            "number_of_students")) %>% 
  distinct()

#Table
apa_table(
  census_data,
    digits = 0,
  caption = "District Classification by Population According to 2016 Census",
  note = "Number of students is estimated and does not include masked data.",
  col.names = c("District Number",
            "District Name",
            "Population",
            "Classification",
            "Number of Students")
)
```

```{r rural play}
#Data Prep
rural_districts <- district_class %>% 
  janitor::clean_names() %>% 
  filter(district_classification == "Rural") %>% 
  select(-c("data_level",
            "public_or_independent",
            "district_population_2016"))

#Viz - Totals by Code
ggplot(rural_districts, 
       aes(special_needs_code, number_of_students,
           color = special_needs_code)) +
  geom_line() +
  facet_wrap(~school_year) +
  labs(title = "Total Students by Exceptionality",
       x = "Exceptionality",
       y = "Number of Students")

#Data for Rural Totals (No categories)
rural_totals <- rural_districts %>% 
  group_by(school_year, district_name, district_number) %>% 
  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#Data Viz Totals/Time
ggplot(rural_totals, aes(school_year, students)) +
  geom_col(aes(color = district_name)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
       x = "School Year",
       y = "Total Students")

#What's going on with Prince George? Why is there such a steep drop in 2006?
pg <- rural_districts %>% 
  filter(district_name == "Prince George") %>% 
  as_tibble()

pg_sums <- pg %>% 
  group_by(school_year) %>% 
  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#What's happening by Dx?
pg_dx_sum <- pg %>% 
  group_by(school_year, special_needs_code) %>% 
  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(district_number, district_classification, number_of_students)) %>% 
  distinct()

ggplot(pg_dx_sum, aes(school_year, student)) +
  geom_col() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~special_needs_code) +
  labs(title = "Students with Exceptionalities in Prince George 1999-2020",
       x = "School Year",
       y = "Number of Students")
# Need some explanation here for explaining facet wrap

#What's going on with gifted?????
pg_gifted <- pg %>% 
  filter(special_needs_code == "P") %>% 
  select(-c(district_number, district_classification))

ggplot(pg_gifted, aes(school_year, number_of_students)) +
  geom_col() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Gifted Students in Prince George 1999-2020",
       subtitle = "Note: Totals under 10 are masked by district in reporting.",
       x = "School Year",
       y = "Total Students")

##Looking into Sooke (SD 62 - pop. 9,269) categorized under Victoria so urban but technically rural
## In progress

sooke <- district_class %>%
  filter(DISTRICT_NAME == "Sooke") %>%
  as_tibble()

ggplot(sooke, aes(SPECIAL_NEEDS_CODE)) +
  facet_wrap(~SCHOOL_YEAR)
  geom_col()
```

```{r urban play}
#Data Prep
urban_districts <- district_class %>% 
  janitor::clean_names() %>% 
  filter(district_classification == "Urban") %>% 
  select(-c("data_level",
            "public_or_independent",
            "district_population_2016"))

urban_totals <- urban_districts %>% 
  group_by(school_year, district_name, district_number) %>% 
  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#Data Viz Totals/Time
ggplot(urban_totals, aes(school_year, students)) +
  geom_col(aes(color = district_name)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
       x = "School Year",
       y = "Total Students")

#Viz - Totals by Code
ggplot(urban_districts, 
       aes(special_needs_code, number_of_students,
           color = special_needs_code)) +
  geom_line() +
  facet_wrap(~school_year) +
  labs(title = "Total Students by Exceptionality",
       x = "Exceptionality",
       y = "Number of Students")
```

### 2020-12-06 scripts/Final_project.Rmd
Updated Methods

@@
 There are several caveats for data use and cleaning in the current analysis.
-First, the BC and OR datasets were not a completely parallel comparison, because the OR dataset includes a broader age range of students (ages 5 - 21) as compared to the BC dataset (ages 5 - 19).
+First, the BC and OR datasets were not a completely parallel comparison, because the OR data set includes a broader age range of students (ages 3 - 21) as compared to the BC data set (ages 5 - 19).
+The Oregon data set was pared down to include a similar age range to BC data; however, it is worth noting that differences in age range may represent different approaches to diagnostic categorization (e.g., intellectual disability in BC can be diagnosed at age 5; whereas the Oregon data follows IDEA where it is more likely young children will be categorized as developmentally delayed).
 Second, data masking necessitates a decision-point for analysis.
-Specifically, the BC Ministry of Education data masks all cells with headcounts below 10 students, therefore our analyses left out these cases when computing effect sizes.
-Data masking was most prominently an issue when examining district-level data in BC, as smaller school districts often masked low-incidence exceptionality data.
-When appropriate, masked data is assigned a convservative estimate of (n = 9) such that enough data units were included in effect size comparisons.
+Specifically, the BC Ministry of Education data masks all cells with headcounts below 10 students, therefore our analyses left out these cases when computing effect sizes for provincial-level data.
+Data masking was most prominently an issue when examining school district-level data in BC, as smaller school districts often masked low-incidence exceptionality data.
+When appropriate, masked data is assigned a conservative estimate of (n = 9) such that enough data units were included in effect size comparisons.
 Third, Canadian census data from 2016 is the basis of classification for district geographical location (e.g., urban, suburban, rural, exurban).
 Due to the geographical distribution of population in BC, only two categories are used for the current analyses: urban and rural.
+Fourth, due to low population numbers with independent schools in BC, the current study will only examine public school data for the years 2002 - 2019.
 

### 2020-12-06 scripts/Final_project.Rmd
Changed headings and YAML

@@
   Janette Avelar, University of Oregon College of Education
+
   David Fainstein, University of Oregon College of Education
+
   Joe Swinehart, Behavioral Research & Teaching
+
   Makayla Whitney, University of Oregon College of Education
@@
   This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
-The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministrty of Education in British Columbia.
+The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministry of Education in British Columbia.
 This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
@@
 Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
-
-  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
-
-  One sentence clearly stating the **general problem** being addressed by  this particular study.
-
-  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
-
-  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
-
-  One or two sentences to put the results into a more **general context**.
-
-  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
-
-  <!-- https://tinyurl.com/ybremelq -->
 
@@
 
-##Current Study
+## Current Study
 Our project primarily examines two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
@@
 
-##Problem Statement and RQs
+## Problem Statement and RQs
 In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
@@
 
-#Methods
+# Methods
 
@@
 
-##Analysis Plan
+## Analysis Plan
 The majority of the analyses for the current study are descriptive in nature.
@@
 $$
-#Results
+# Results
 Writing on results here...
@@
 ```
+
+# Discussion
+Regarding the first research question, it is apparent that the 2016 supreme court decision did not impact any of the 12 designation categories as hypothesized.
+The trendlines for each of the exceptionality categories remained consistent post-2016 in BC and this mapped on to Oregon trends in exceptionality designation reasonably closely.
+This finding suggests that the resources and practices around supporting student
+
+The rapid decrease in students identified as gifted in rural districts between the years of 2004 and 2007 cannot be explained by one sole factor in this exploratory analysis.
+With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts.
+In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky & Clelland, 2016).
+In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students.
+This 2016 shift in policy links to a higher degree of resources afforded to gifted students through “independent, guided education” (p.
+52).
+Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students.
+For example, Lo et al.
+(2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.

### 2020-12-06 scripts/Final_project.Rmd
Comprehensive data display for intro

@@
 library(esc)
+library(dplyr)
+install.packages("splom")
 
@@
 Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore & Dynarski, 2017).
+
+```{comprehensive display}
+
+g_vic <- district_class %>%
+  filter(DISTRICT_NAME == "Greater Victoria") %>%
+  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR),
+         SPECIAL_NEEDS_CODE = as.factor(SPECIAL_NEEDS_CODE),
+          SPECIAL_NEEDS_CODE = as.numeric(SPECIAL_NEEDS_CODE),
+         NUMBER_OF_STUDENTS = as.factor(NUMBER_OF_STUDENTS),
+         NUMBER_OF_STUDENTS = as.numeric(NUMBER_OF_STUDENTS)) %>%
+  select("SCHOOL_YEAR",
+         "SPECIAL_NEEDS_CODE",
+         "NUMBER_OF_STUDENTS") %>%
+  as_tibble()
+
+install.packages("GGally")
+library(GGally)
+
+#best one yet! Box-and-whisker plot
+ggpairs(g_vic,
+        upper = list(continuous = ggally_dot_and_box, combo = ggally_box_no_facet),
+        lower = list (continuous = "smooth",
+                      combo = "facetdensity",
+                      mapping = aes(color = SPECIAL_NEEDS_CODE)))
+
+#let's try bar graph
+ggpairs(g_vic,
+        upper = list(continuous = ggally_barDiag, combo = ggally_barDiag),
+        lower = list (continuous = "smooth",
+                      combo = "facetdensity",
+                      mapping = aes(color = SPECIAL_NEEDS_CODE)))
+
+ggally_dot_and_box(g_vic, 
+                   mapping = ggplot2::aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS),
+                   boxPlot = TRUE) +
+  theme(axis.text.x = element_blank())
+
+```
 

### 2020-12-06 scripts/Final_project.Rmd
Deleted splom

@@
 library(dplyr)
-install.packages("splom")
 

### 2020-12-06 scripts/Final_project.tex
Deleted splom

--- 2020-12-06 scripts/Final_project.Rmd
@@
----
-title             : "Exploring Anomalies in Pacific Northwest Special Education Trends"
-shorttitle        : "PACNW SpEd Trends"
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  english,
+  man]{apa6}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Exploring Anomalies in Pacific Northwest Special Education Trends},
+  pdfauthor={David Fainstein1, Janette Avelar1, Makayla Whitney1, \& Joe Swinehart1},
+  pdflang={en-EN},
+  pdfkeywords={Vancouver, British Columbia, Oregon, Special Education, SpEd},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+% Manuscript styling
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
 
-author: 
-  - name          : "David Fainstein"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "1655 Alder, Eugene, Oregon, USA  97403"
-    email         : "dfainste@uoregon.edu"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - Conceptualization
-      - Writing - Original Draft Preparation
-      - Writing - Review & Editing
-  - name          : "Janette Avelar"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
-  - name          : "Makayla Whitney"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
-  - name          : "Joe Swinehart"
-    affiliation   : "1"
-    role:
-      - Writing - Review & Editing
+% Table formatting
+\usepackage{longtable}
+\usepackage{lscape}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
 
-affiliation:
-  - id            : "1"
-    institution   : "University of Oregon"
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
 
-authornote: |
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
+
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
+\makeatother
+\shorttitle{PACNW SpEd Trends}
+\keywords{Vancouver, British Columbia, Oregon, Special Education, SpEd\newline\indent Word count: 0<n<1,000,000}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
+\usepackage{csquotes}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
+\fi
+
+\title{Exploring Anomalies in Pacific Northwest Special Education Trends}
+\author{David Fainstein\textsuperscript{1}, Janette Avelar\textsuperscript{1}, Makayla Whitney\textsuperscript{1}, \& Joe Swinehart\textsuperscript{1}}
+\date{}
+
+\authornote{
+
 Janette Avelar, University of Oregon College of Education
@@
 
-  Joe Swinehart, Behavioral Research & Teaching
+Joe Swinehart, Behavioral Research \& Teaching
 
@@
 
-  Contact: dfainste@uoregon.edu
-abstract: |
+Contact: \href{mailto:dfainste@uoregon.edu}{\nolinkurl{dfainste@uoregon.edu}}
+
+The authors made the following contributions.
+David Fainstein: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Janette Avelar: Writing - Review \& Editing; Makayla Whitney: Writing - Review \& Editing; Joe Swinehart: Writing - Review \& Editing.
+
+Correspondence concerning this article should be addressed to David Fainstein, 1655 Alder, Eugene, Oregon, USA 97403. E-mail: \href{mailto:dfainste@uoregon.edu}{\nolinkurl{dfainste@uoregon.edu}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Oregon}
+
+\abstract{
 This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
@@
 Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
+}
 
-keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
-wordcount         : "0<n<1,000,000"
+\begin{document}
+\maketitle
 
-bibliography      : ["r-references.bib"]
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
----
-
-```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
-library("papaja")
-r_refs("r-references.bib")
-library(here)
-library(tidyverse)
-library("readxl")
-library(janitor)
-library(esc)
-
-BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
-
-BCpopulation <- read_excel(here("data", "BCpopdata.xlsx")) %>% 
-  select("SCHOOL_YEAR",
-         "TOTAL_ENROLMENT",
-         "PUBLIC_OR_INDEPENDENT") %>% 
-  filter(PUBLIC_OR_INDEPENDENT == "BC PUBLIC SCHOOL") %>% 
-  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
-         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)
-
-BCpop_district <- read_excel(here("data", "BCdistrictpop.xlsx")) %>% 
-  select("SCHOOL_YEAR",
-         "TOTAL_ENROLMENT",
-         "DISTRICT_NUMBER") %>% 
-  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
-         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)
-
-BCdatastripped <- BCdata %>% 
-    #Makayla & Janette - may want to include more columns here with the select for our global "BCdatastripped" df
-    select("SCHOOL_YEAR", 
-         "DATA_LEVEL", 
-         "SPECIAL_NEEDS_CODE", 
-         "NUMBER_OF_STUDENTS", 
-         "PUBLIC_OR_INDEPENDENT",
-         "DISTRICT_NUMBER") %>% 
-  filter(SPECIAL_NEEDS_CODE != "M",
-         SPECIAL_NEEDS_CODE != "J",
-         SPECIAL_NEEDS_CODE != "N") %>% 
-    mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
-                          "Physically Dependent" = "A",
-                          "Deafblind" = "B",
-                          "Moderate+ ID" = "C",
-                          "Chronic Health" = "D",
-                          "Visual Impairment" = "E",
-                          "Deaf or Hard of Hearing" = "F",
-                          "Autism" = "G",
-                          "High Behavioural Need" = "H",
-                          "Mild ID" = "K",
-                          "Learning Disability" = "Q",
-                          "Gifted" = "P",
-                          "Mild Behavioural Need" = "R"),
-           SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
-           NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
-           SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(SCHOOL_YEAR >= 9)
-
-BC_districtdata <- BCdatastripped %>% 
-  filter(DATA_LEVEL == "DISTRICT LEVEL")
-Prevalence_data <- inner_join(BCpop_district, BC_districtdata, by = c("SCHOOL_YEAR", "DISTRICT_NUMBER"), all.x = TRUE)
-
-```
-
-# Introduction
 Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society (Ball, 1998).
@@
 When policy is enacted, the intended outcomes are often not realized for months, years, or decades (Gu et al., 2018).
-For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke & McArdle, 2009).
-Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore & Dynarski, 2017).
+For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke \& McArdle, 2009).
+Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore \& Dynarski, 2017).
 
-## Current Study
+\hypertarget{current-study}{%
+\subsection{Current Study}\label{current-study}}
+
 Our project primarily examines two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
@@
 Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017).
-Canadian supreme court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire.
+Canadian supreme court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers' that school districts are required to hire.
 The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change.
@@
 Districts are categorized by population size and proximity to metropolitan areas, as recorded and defined by the Statistics Canada census from 2016.
-School district level data is useful for examining longitudinal trends for individual administrative units (e.g., superintendent’s office and director(s) of inclusive education); however, this level of data analysis cannot speak to province-wide trends in the culture of superordinate legal directives which impact swift policy changes.
+School district level data is useful for examining longitudinal trends for individual administrative units (e.g., superintendent's office and director(s) of inclusive education); however, this level of data analysis cannot speak to province-wide trends in the culture of superordinate legal directives which impact swift policy changes.
 Given this, we will run analyses on both exemplar districts to highlight localized policy impacts on practice and amalgamated districts (e.g., school districts defined as rural by Statistics Canada) to investigate cultural shifts in the policy-to-practice association.
 
-## Problem Statement and RQs
-In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
+\hypertarget{problem-statement-and-rqs}{%
+\subsection{Problem Statement and RQs}\label{problem-statement-and-rqs}}
+
+In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker \& Bergmann, 2013).
 While Canadian K-12 academic achievement outcomes are generally higher than a national-level comparison with the US, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education.
@@
 
-# Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 The three datasets were used for analysis in R (cite R) for reproducibility and data visualization.
-All code for analysis and data representation is publicly available for reproducibility (https://github.com/JoeSwinehart/datascienceproject).
+All code for analysis and data representation is publicly available for reproducibility (\url{https://github.com/JoeSwinehart/datascienceproject}).
 There are several caveats for data use and cleaning in the current analysis.
@@
 
-## Analysis Plan
+\hypertarget{analysis-plan}{%
+\subsection{Analysis Plan}\label{analysis-plan}}
+
 The majority of the analyses for the current study are descriptive in nature.
@@
 That is, effect sizes between two groups (e.g., group one is three years before a policy change, and group two is three years after a policy change) may be examined to detect the magnitude of differences between groups if they are hypothesized to represent a change in practice stemming from policy implementation.
-More specifically on the calculation of effect sizes, the Hedge’s g statistic will be used because of the robustness against unequal group sizes.
+More specifically on the calculation of effect sizes, the Hedge's g statistic will be used because of the robustness against unequal group sizes.
 
-$$
+\[
 Hedges' g = \frac{M_1 + M_2}{SD_{pooled}} 
-$$
-# Results
-Writing on results here...
+\]
+\# Results
+Writing on results here\ldots{}
 
 First, we see that the BC trends do not show significant changes after the 2016 supreme court ruling.
-```{r Provinical_Data_Overview}
+\includegraphics{Final_project_files/figure-latex/Provinical_Data_Overview-1.pdf}
 
-BCprovincial <- BCdatastripped %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL",
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")  %>% 
-  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
-         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
-  mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995)
+\begin{verbatim}
+## # A tibble: 198 x 18
+##    year  disability    x6    x7    x8    x9   x10   x11   x12   x13   x14   x15
+##    <chr> <chr>      <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
+##  1 2002~ INTELLECT~   135   172   220   261   279   303   370   364   388   378
+##  2 2002~ HEARING I~    46    66    69    67    79    78    69    70    53    72
+##  3 2002~ SPEECH OR~  2502  2466  2443  2205  1894  1282   874   677   493   340
+##  4 2002~ VISUAL IM~    18    27    25    16    12    26    19    24    31    19
+##  5 2002~ EMOTIONAL~    81   178   241   292   347   433   504   487   549   548
+##  6 2002~ ORTHOPEDI~    65    77    64    49    64    71    50    48    44    43
+##  7 2002~ DEAF-BLIN~     0     2     0     0     1     0     2     0     4     2
+##  8 2002~ MULTIPLE ~     0     0     0     0     0     0     0     0     0     0
+##  9 2002~ AUTISM       267   289   290   334   325   316   290   256   243   219
+## 10 2002~ TRAUMATIC~     6    11    17    23    22    19    21    23    36    36
+## # ... with 188 more rows, and 6 more variables: x16 <dbl>, x17 <dbl>,
+## #   x18 <dbl>, x19 <dbl>, x20 <dbl>, x21 <dbl>
+\end{verbatim}
 
- ggplot(BCprovincial, aes(x = SCHOOL_YEAR, y = prevalence)) +
-  geom_line(color = "cornflowerblue") +
-   facet_wrap(~SPECIAL_NEEDS_CODE) +
-   labs(title = "BC Provincial Prevalence of Students with Exceptionalities by Category",
-        x = "School Year",
-        y = "Prevalence Rate")
-```
-
-```{r cleaning Oregon data}
-primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
-cleanest_primo <-primodata %>% 
-  remove_empty(which = c("rows", "cols")) %>% 
-  clean_names()
-
-cleanest_primo <- cleanest_primo %>% 
-  select(-six_to_twentyone_total, -x24)
-
-cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
-cleanest_primo2
-
-fresh_primo <- cleanest_primo2 %>% 
-  pivot_longer(starts_with("x"), names_to = "age", values_to = "frequency") %>% 
-  mutate(age = parse_number(age), 
-         year = parse_number(year))
-
-fresh_primo %>% 
-  group_by(year, disability) %>% 
-  summarize(total=sum(frequency, na.rm=TRUE))
-
-```
+\begin{verbatim}
+## # A tibble: 198 x 3
+## # Groups:   year [18]
+##     year disability                    total
+##    <dbl> <chr>                         <dbl>
+##  1  2002 AUTISM                         3339
+##  2  2002 DEAF-BLINDNESS                   17
+##  3  2002 DEVELOPMENTAL DELAY               0
+##  4  2002 EMOTIONAL DISTURBANCE          4736
+##  5  2002 HEARING IMPAIRMENT              873
+##  6  2002 INTELLECTUAL DISABILITY        4387
+##  7  2002 MULTIPLE DISABILITIES             0
+##  8  2002 ORTHOPEDIC IMPAIRMENT           754
+##  9  2002 SPEECH OR LANGUAGE IMPAIRMENT 15784
+## 10  2002 TRAUMATIC BRAIN INJURY          306
+## # ... with 188 more rows
+\end{verbatim}
 
 This is roughly in line with Oregon's designation data, showing that there were no state-wide anomalies after the year 2016.
-```{r}
-#this is more in line with the BC graphing
-ggplot(fresh_primo, aes(year, frequency))+
- geom_smooth(color = "cornflowerblue") +
- theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
- facet_wrap(~disability)+
-    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")
 
-```
+\begin{verbatim}
+## Warning: Removed 190 rows containing non-finite values (stat_smooth).
+\end{verbatim}
 
-Even for diagnostically malleable categories, such as students classified in Oregon has having an "emotional disturbance," or in BC students requiring a "high behavioural need", Oregon and BC had highly similar trends.
+\includegraphics{Final_project_files/figure-latex/unnamed-chunk-1-1.pdf}
 
-``` {r BCOR combo}
+Even for diagnostically malleable categories, such as students classified in Oregon has having an \enquote{emotional disturbance,} or in BC students requiring a \enquote{high behavioural need}, Oregon and BC had highly similar trends.
 
-combined <- fresh_primo %>% 
-  as_tibble() %>% 
-    mutate(ORheadcount = fct_recode(disability,
-                          "Deafblind" = "DEAF-BLINDNESS",
-                          "Moderate+ ID" = "INTELLECTUAL DISABILITY",
-                          "High Behavioural Need" = "EMOTIONAL DISTURBANCE",
-                          "Visual Impairment" = "VISUAL IMPAIRMENT",
-                          "Deaf or Hard of Hearing" = "HEARING IMPAIRMENT",
-                          "Autism" = "AUTISM",
-                          "Mild ID" = "INTELLECTUAL DISABILITY")) %>% 
-  filter(ORheadcount == "High Behavioural Need") %>% 
-  group_by(year) %>% 
-  summarise(frequency = sum(frequency))
+\includegraphics{Final_project_files/figure-latex/BCOR combo-1.pdf}
 
-BCAutism <- BCprovincial %>% 
-  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need")
+Then, we examine the rural versus urban trends and notice a few things\ldots{}
 
-ggplot() +
-  geom_smooth(data = combined, aes(x = year, y = frequency, color = "Oregon")) +
-  geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS, color = "British Columbia"), se = FALSE) +
-  labs(title = "BC and Oregon Headcount of Students with Chronic Health Conditions",
-        x = "School Year",
-        y = "Headcount",
-       color = "Legend",
-        scale_color_manual(labels = c("Oregon", "British Columbia"), values = c("blue", "red")))
+\begin{verbatim}
+## Warning in FUN(newX[, i], ...): NAs introduced by coercion
+\end{verbatim}
 
-```
+\begin{table}[tbp]
 
-Then, we examine the rural versus urban trends and notice a few things...
+\begin{center}
+\begin{threeparttable}
 
-```{r urban vs. rural}
+\caption{\label{tab:District Census Population Table}District Classification by Population According to 2016 Census}
 
-bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))
+\begin{tabular}{lllll}
+\toprule
+District Number & District Name & Population & Classification & Number of Students\\
+\midrule
+022 & Vernon & 61,334 & Rural & 16,768\\
+023 & Central Okanagan & 194,882 & Urban & 43,609\\
+028 & Quesnel & 23,146 & Rural & 9,036\\
+033 & Chilliwack & 101,512 & Urban & 28,997\\
+034 & Abbotsford & 180,518 & Urban & 38,369\\
+035 & Langley & 2,463,431 & Urban & 43,043\\
+036 & Surrey & 2,463,431 & Urban & 143,222\\
+037 & Delta & 2,463,431 & Urban & 43,897\\
+038 & Richmond & 2,463,431 & Urban & 40,221\\
+039 & Vancouver & 2,463,431 & Urban & 132,584\\
+040 & New Westminster & 2,463,431 & Urban & 11,714\\
+041 & Burnaby & 2,463,431 & Urban & 47,952\\
+042 & Maple Ridge-Pitt Meadows & 2,463,431 & Urban & 33,633\\
+043 & Coquitlam & 2,463,431 & Urban & 88,291\\
+044 & North Vancouver & 2,463,431 & Urban & 39,657\\
+045 & West Vancouver & 2,463,431 & Urban & 11,732\\
+046 & Sunshine Coast & 2,463,431 & Urban & 13,283\\
+047 & Powell River & 16,783 & Rural & 6,816\\
+052 & Prince Rupert & 12,687 & Rural & 7,222\\
+053 & Okanagan Similkameen & 43,432 & Rural & 6,596\\
+057 & Prince George & 86,622 & Rural & 39,443\\
+061 & Greater Victoria & 367,770 & Urban & 50,051\\
+062 & Sooke & 367,770 & Urban & 24,126\\
+063 & Saanich & 367,770 & Urban & 19,533\\
+067 & Okanagan Skaha & 43,432 & Rural & 19,156\\
+068 & Nanaimo-Ladysmith & 104,936 & Urban & 30,870\\
+070 & Port Alberni & 25,112 & Rural & 11,654\\
+073 & Kamloops-Thompson & 103,811 & Urban & 32,620\\
+075 & Mission & 33,261 & Rural & 13,858\\
+079 & Cowichan Valley & 44,451 & Rural & 21,479\\
+081 & Fort Nelson & 18,307 & Rural & 1,319\\
+083 & North Okanagan-Shuswap & 43,432 & Rural & 23,720\\
+\bottomrule
+\addlinespace
+\end{tabular}
 
-district_class <- bc_data_district_classification %>%
-  select("SCHOOL_YEAR",
-         "DATA_LEVEL",
-         "PUBLIC_OR_INDEPENDENT",
-         "DISTRICT_NUMBER",
-         "DISTRICT_NAME",
-         "DISTRICT_POPULATION_2016",
-         "DISTRICT_CLASSIFICATION",
-         "SPECIAL_NEEDS_CODE",
-         "NUMBER_OF_STUDENTS") %>%
-  filter(DATA_LEVEL == "DISTRICT LEVEL",
-         PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
-  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
-    select(-"name")
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Number of students is estimated and does not include masked data.}
+\end{tablenotes}
 
-#Attempting to make columns numeric
-#view(district_class) #Need cols 6 & 8
+\end{threeparttable}
+\end{center}
 
-x <- c(6,8)
-district_class[ , x] <- apply(district_class[ , x], 2,
-                               function(x)as.numeric(as.character(x)))
-#
-```
+\end{table}
 
-``` {r District Census Population Table}
-#Data Prep
-census_data <- district_class %>% 
-  janitor::clean_names() %>% 
-  group_by(district_name, district_number) %>% 
-  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c("school_year",
-            "data_level",
-            "public_or_independent",
-            "special_needs_code",
-            "number_of_students")) %>% 
-  distinct()
+\begin{verbatim}
+## Warning: Removed 129 row(s) containing missing values (geom_path).
+\end{verbatim}
 
-#Table
-apa_table(
-  census_data,
-    digits = 0,
-  caption = "District Classification by Population According to 2016 Census",
-  note = "Number of students is estimated and does not include masked data.",
-  col.names = c("District Number",
-            "District Name",
-            "Population",
-            "Classification",
-            "Number of Students")
-)
-```
+\includegraphics{Final_project_files/figure-latex/rural play-1.pdf} \includegraphics{Final_project_files/figure-latex/rural play-2.pdf} \includegraphics{Final_project_files/figure-latex/rural play-3.pdf}
 
-```{r rural play}
-#Data Prep
-rural_districts <- district_class %>% 
-  janitor::clean_names() %>% 
-  filter(district_classification == "Rural") %>% 
-  select(-c("data_level",
-            "public_or_independent",
-            "district_population_2016"))
+\begin{verbatim}
+## Warning: Removed 12 rows containing missing values (position_stack).
+\end{verbatim}
 
-#Viz - Totals by Code
-ggplot(rural_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
-       x = "Exceptionality",
-       y = "Number of Students")
+\includegraphics{Final_project_files/figure-latex/rural play-4.pdf} \includegraphics{Final_project_files/figure-latex/rural play-5.pdf}
 
-#Data for Rural Totals (No categories)
-rural_totals <- rural_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
+\begin{verbatim}
+## geom_col: width = NULL, na.rm = FALSE
+## stat_identity: na.rm = FALSE
+## position_stack
+\end{verbatim}
 
-#Data Viz Totals/Time
-ggplot(rural_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
+\includegraphics{Final_project_files/figure-latex/urban play-1.pdf}
 
-#What's going on with Prince George? Why is there such a steep drop in 2006?
-pg <- rural_districts %>% 
-  filter(district_name == "Prince George") %>% 
-  as_tibble()
+\begin{verbatim}
+## Warning: Removed 25 row(s) containing missing values (geom_path).
+\end{verbatim}
 
-pg_sums <- pg %>% 
-  group_by(school_year) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
+\includegraphics{Final_project_files/figure-latex/urban play-2.pdf}
 
-#What's happening by Dx?
-pg_dx_sum <- pg %>% 
-  group_by(school_year, special_needs_code) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, district_classification, number_of_students)) %>% 
-  distinct()
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
-ggplot(pg_dx_sum, aes(school_year, student)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  facet_wrap(~special_needs_code) +
-  labs(title = "Students with Exceptionalities in Prince George 1999-2020",
-       x = "School Year",
-       y = "Number of Students")
-# Need some explanation here for explaining facet wrap
-
-#What's going on with gifted?????
-pg_gifted <- pg %>% 
-  filter(special_needs_code == "P") %>% 
-  select(-c(district_number, district_classification))
-
-ggplot(pg_gifted, aes(school_year, number_of_students)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Gifted Students in Prince George 1999-2020",
-       subtitle = "Note: Totals under 10 are masked by district in reporting.",
-       x = "School Year",
-       y = "Total Students")
-
-##Looking into Sooke (SD 62 - pop. 9,269) categorized under Victoria so urban but technically rural
-## In progress
-
-sooke <- district_class %>%
-  filter(DISTRICT_NAME == "Sooke") %>%
-  as_tibble()
-
-ggplot(sooke, aes(SPECIAL_NEEDS_CODE)) +
-  facet_wrap(~SCHOOL_YEAR)
-  geom_col()
-```
-
-```{r urban play}
-#Data Prep
-urban_districts <- district_class %>% 
-  janitor::clean_names() %>% 
-  filter(district_classification == "Urban") %>% 
-  select(-c("data_level",
-            "public_or_independent",
-            "district_population_2016"))
-
-urban_totals <- urban_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#Data Viz Totals/Time
-ggplot(urban_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
-
-#Viz - Totals by Code
-ggplot(urban_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
-       x = "Exceptionality",
-       y = "Number of Students")
-```
-
-# Discussion
 Regarding the first research question, it is apparent that the 2016 supreme court decision did not impact any of the 12 designation categories as hypothesized.
@@
 With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts.
-In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky & Clelland, 2016).
+In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky \& Clelland, 2016).
 In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students.
-This 2016 shift in policy links to a higher degree of resources afforded to gifted students through “independent, guided education” (p.
-52).
+This 2016 shift in policy links to a higher degree of resources afforded to gifted students through \enquote{independent, guided education} (p.~52).
 Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students.
-For example, Lo et al.
-(2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+For example, Lo et al.~(2019) suggest that the British Columbia Ministry of Education's recent curriculum update makes amends for traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+
+\end{document}

### 2020-12-06 scripts/j_play.Rmd
Deleted splom

@@
 library(janitor)
+library(kableExtra)
 
@@
 
-``` {r District Census Population Table}
+``` {r District Level Population Table}
 #Data Prep
@@
   janitor::clean_names() %>% 
-  group_by(district_name, district_number) %>% 
+  group_by(district_name, district_number, school_year) %>% 
   mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
@@
 )
+
+#Ignore
+
 ```
@@
   geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+    theme(text = element_text(size = 6),
+      axis.text.x = element_text(angle = 90, hjust = 1)) +
   facet_wrap(~special_needs_code)

### 2020-12-06 scripts/BCOR_combo.Rmd
Updated BCOR_combo for Commenting Ideas

--- 2020-12-02 scripts/BCOR_combo.Rmd
@@
 
+#One data set for global use and tidying for code chunks
 BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))
 
+#reading in and initial tidying provincial-level population data
 BCpopulation <- read_excel(here("data", "BCpopdata.xlsx")) %>% 
@@
 
+#reading in and initial tidying district-level population data
+BCpop_district <- read_excel(here("data", "BCdistrictpop.xlsx")) %>% 
+  select("SCHOOL_YEAR",
+         "TOTAL_ENROLMENT",
+         "DISTRICT_NUMBER") %>% 
+  mutate(SCHOOL_YEAR = as.factor(SCHOOL_YEAR), 
+         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)+8)
+
+#reading in and initial tidying Provincial headcount data on SPED categories by district and province
 BCdatastripped <- BCdata %>% 
-    #Makayla & Janette - may want to include more columns here with the select for our global "BCdatastripped" df
     select("SCHOOL_YEAR", 
@@
            SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(SCHOOL_YEAR >= 9)
+  filter(SCHOOL_YEAR >= 9) #aligning data to fit chronologically with OR data
+
+#Initializing data set for calculating prevalence by joining student population data with SPED headcount data
+BC_districtdata <- BCdatastripped %>% 
+  filter(DATA_LEVEL == "DISTRICT LEVEL")
+Prevalence_data <- inner_join(BCpop_district, BC_districtdata, by = c("SCHOOL_YEAR", "DISTRICT_NUMBER"), all.x = TRUE)
 
@@
 ```{r Provinical_Data_Overview}
-
+#cleaning provincial-level data and calculating prevalence data
 BCprovincial <- BCdatastripped %>% 
@@
 
+#Initial plot showing each prevalence rates of designation categories
  ggplot(BCprovincial, aes(x = SCHOOL_YEAR, y = prevalence)) +
@@
 
+#preparing OR data to fit with BC data to map on to one ggplot. Exceptionality category head count is summed by year.
 combined <- fresh_primo %>% 
@@
 
+#selecting the exceptionality category of interest from BC data to match cleaned OR data
 BCAutism <- BCprovincial %>% 
@@
 
+#One figure that shows head counts in each of OR and BC for one exceptionality as an exemplar   
 ggplot() +
@@
 
-ESdata <- BCdatastripped %>% 
-  filter(SPECIAL_NEEDS_CODE == "High Behavioural Need",
-         DATA_LEVEL == "DISTRICT LEVEL",
-         PUBLIC_OR_INDEPENDENT == "BC Public School",
-         DISTRICT_NUMBER == "036")  %>% 
-  mutate(Public_population = BCpopulation$TOTAL_ENROLMENT[SCHOOL_YEAR-8],
-         prevalence = NUMBER_OF_STUDENTS/Public_population) %>% 
+#hard coding the 14 urban school districts
+urban <- c("023", "034", "035", "036", "037", "038", "039", "040", "041", "042", "043", "044", "045", "046")
+
+#Rural school districts prevalence for one exceptionality, then grouped 4 years preceding 2016, and 3 years post-2016.
+Code structure used for running multiple exceptionality categories (behavioural needs, learning disability, etc)
+ESdata <- Prevalence_data %>% 
+  filter(SPECIAL_NEEDS_CODE == "Mild Behavioural Need",
+         !DISTRICT_NUMBER %in% urban)  %>% 
+  mutate(prevalence = NUMBER_OF_STUDENTS/TOTAL_ENROLMENT) %>% 
   mutate(SCHOOL_YEAR = SCHOOL_YEAR+1995,
          pre16 = ifelse(SCHOOL_YEAR <= 2016, "before16", "after16")) %>%
+  filter(SCHOOL_YEAR >= 2013) %>% 
   group_by(pre16) %>% 
@@
 
+#assigning an object to hold pre-2016 group values
 presup <- ESdata %>% 
@@
 
+#assigning an object to hold pre-2016 group values
 postsup <- ESdata %>% 
@@
 
+#assigning variables as objects that are needed to calculate effect size
+presup_count <- count(presup)
 presup_mean <- sapply(presup, mean)
 presup_sd <- sapply(presup, sd)
+postsup_count <- count(postsup)
 postsup_mean <- sapply(postsup, mean)
@@
 
-esc_mean_sd(grp1m = presup_mean, grp1sd = presup_sd, grp1n = 13, grp2m = postsup_mean, grp2sd = postsup_sd, grp2n = 3, es.type = "g")
+#calculating Hedges' g
+esc_mean_sd(grp1m = presup_mean, grp1sd = presup_sd, grp1n = presup_count, grp2m = postsup_mean, grp2sd = postsup_sd, grp2n = postsup_count, es.type = "g")
+
+#running two-sample t-test for complete statistical test information
+t.test(presup, postsup)
 
@@
 We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
+
+Here is the equation for Hedges' g: $Hedges' g = \frac{M_1 + M_2}{SD_{pooled}} $
+``` {r methods}
+
+```
 

### 2020-12-07 scripts/Final_project.Rmd
Getting Closer!

@@
 With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
-(1) Do student prevalence rates of disability or disorder decrease after the supreme court ruling of 2016 against existing trends?
-(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+\begin{enumerate}
+  \item Do student prevalence rates of disability or disorder decrease after the supreme court ruling of 2016 against existing trends?
+  \item Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+ \end{enumerate}
 
 # Methods
-
 Datasets were extracted from publicly available sources.
@@
 $$
+
 # Results
-Writing on results here...
+To answer the first research question, we examine the provincial-level prevalence data by exceptionality to view trendlines over 15 years.
+It is apparent that 2016 is not a point of inflection for any of the twelve designations listed.
+Low-incidence exceptionalities remain consistent, and high-incidence exceptionalities are linearly (e.g., Learning Disabilities) or exponentially (e.g., Autism) increasing, and some high-incidence exceptionalities are linearly decreasing (e.g., Gifted).
+Due to the fact that there are no significant points of inflection observed on any of the 12 trendlines in figure 1, no further investigation (e.g., calculating effect size) was warranted.
 
-First, we see that the BC trends do not show significant changes after the 2016 supreme court ruling.
 ```{r Provinical_Data_Overview}
@@
 This is roughly in line with Oregon's designation data, showing that there were no state-wide anomalies after the year 2016.
+Similar to the BC data, we see a steady increase in students identified under the autism diagnosis, while most categories remained relatively stable over all of the years surveyed.
+The recent growth in Developmental Delays and Emotional Disturbance would be worth watching in future years, but we can reasonably predict that more rigid  categories like Deaf-Blindness and Traumatic Brain Injury will remain stable in the years to come.
+
+Looking at Oregon disability prevalence by age gives us some idea of what happens to these categories over time, and also invites some questions.
+Most visually striking is the drop-off of students identified as having a Speech or Language Impairment, which makes sense, as it is addressed early in school and highly correctable.
+By middle school, most students have been exited from the speech program.
+At a striking contrast would be things that cannot be corrected, such as Deaf-Blindness or Traumatic Brain Injury.
+As expected, these numbers are extremely static over time.
+More thought-provoking would be the numbers for Emotional Disturbance.
+These tend to grow fairly steadily during elementary and middle school, but around age 16 the numbers actually start to decrease.
+One might wonder if this drop is due to these students no longer being part of the school system, as they may be far more likely to drop out and/or enter a different system as adjudicated youth.
+It certainly would be worth an additional look in the future.
+
 ```{r}
@@
 
-Then, we examine the rural versus urban trends and notice a few things...
+Taken together, we see that at the provincial- and state-levels, there is consistency in identification of students with exceptionalities.
+It is clear that a notable inflection point in 2016 for the BC data would have represented a significant change from practice when comparing student identification practices in Oregon; however, there is no apparent inflection point in the BC data.
+
+To examine potential differences within BC, the authors chose to divide school districts by urban or rural designation according to Statistics Canada guidelines on characterizing living populations.
 
@@
 
+There are two major outcomes of interest observed with the data sets when disaggregated by rural and urban school districts.
+First, the trends of rural student exceptionality designations followed urban trends closely.
+The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (ps > 0.05).
+However, for diagnostically malleable categories (e.g., Moderate Behavioural designation) in rural school districts, 2016 was an inflection point for an increase in prevalence rates.
+When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (t(239) = -2.113, p = 0.036, g = 0.261, CI[-0.504, -0.019]), Learning Disability (t(262) = -1.984, p = 0.483, g = 0.230, CI[-0.453, -0.007]), and Chronic Health (t(255) = -2.068, p = 0.040, g = 0.247, CI[-0.479, -0.015]).
+Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges’ g > 0.2, and g < 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI[-0.383, 0.144]).
+
 # Discussion
-Regarding the first research question, it is apparent that the 2016 supreme court decision did not impact any of the 12 designation categories as hypothesized.
+Regarding the first research question, it is apparent that the 2016 Supreme Court decision did not impact any of the 12 designation categories as hypothesized.
 The trendlines for each of the exceptionality categories remained consistent post-2016 in BC and this mapped on to Oregon trends in exceptionality designation reasonably closely.
-This finding suggests that the resources and practices around supporting student
+This finding suggests that the resources and practices around supporting students with exceptionalities did not differ significantly from a provincial policy determination.
+
+Data on the second research question on rural versus urban school district response to the 2016 Supreme Court ruling suggests that the authors correctly hypothesized that there are differences for rural versus urban districts.
+Urban school districts largely followed provincial-level trends, and rural school districts had small effect sizes of increased prevalence for the post-2016 cohort of students.
+That is, when looking at all rural school districts and comparing two groups, four years preceding 2017 and three years 2017 through 2019, there is a significant increase in prevalence for designations in which clinicians have diagnostic flexibility (e.g., behavioural designations, temporary health conditions).
+This finding suggests that clinicians and administrators in rural school districts responded to the 2016 Supreme Court ruling by changing exceptionality designation practices that align with an increase in prevalence.
 
@@
 For example, Lo et al.
-(2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+(2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+
+# References

### 2020-12-07 scripts/Final_project.Rmd
Beginning of references
IODPbY(na)
Inline Oregon Disability Prevalence by Age (now added)

@@
 The recent growth in Developmental Delays and Emotional Disturbance would be worth watching in future years, but we can reasonably predict that more rigid  categories like Deaf-Blindness and Traumatic Brain Injury will remain stable in the years to come.
+```{r Oregon plot}
 
+ggplot(fresh_primo, aes(age, frequency))+
+  geom_line(aes(color=disability))+
+  facet_wrap(~year)+
+    labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
+
+```
 Looking at Oregon disability prevalence by age gives us some idea of what happens to these categories over time, and also invites some questions.
@@
 # References
+
+RStudio Team (2020). RStudio: Integrated Development Environment for R.
+  RStudio, PBC, Boston, MA URL http://www.rstudio.com/.

### 2020-12-07 scripts/Final_project.Rmd
References section round 2

@@
 library(dplyr)
+library(GGally)
 
@@
   as_tibble()
-
-install.packages("GGally")
-library(GGally)
 
@@
 
+\newpage
 # References
 
-RStudio Team (2020). RStudio: Integrated Development Environment for R.
-  RStudio, PBC, Boston, MA URL http://www.rstudio.com/.
+@book{american2013diagnostic,
+  title={Diagnostic and statistical manual of mental disorders (DSM-5{\textregistered})},
+  author={American Psychiatric Association and others},
+  year={2013},
+  publisher={American Psychiatric Pub}
+}
+
+@article{ball1998big,
+  title={Big policies/small world: An introduction to international perspectives in education policy},
+  author={Ball, Stephen J},
+  journal={Comparative education},
+  volume={34},
+  number={2},
+  pages={119--130},
+  year={1998},
+  publisher={Taylor \& Francis}
+}
+
+@Manual{R-base,
+  title = {R: A Language and Environment for Statistical Computing},
+  author = {{R Core Team}},
+  organization = {R Foundation for Statistical Computing},
+  address = {Vienna, Austria},
+  year = {2020},
+  url = {https://www.R-project.org/},
+}
+
+@article{cohen2009school,
+  title={School climate: Research, policy, practice, and teacher education},
+  author={Cohen, Jonathan and McCabe, Libby and Michelli, Nicholas M and Pickeral, Terry},
+  journal={Teachers college record},
+  volume={111},
+  number={1},
+  pages={180--213},
+  year={2009},
+  publisher={Columbia}
+}
+
+citation("dplyr")
+
+citation("esc")
+
+citation("GGally")
+
+@article{gu2018principals,
+  title={How principals of successful schools enact education policy: Perceptions and accounts from senior and middle leaders},
+  author={Gu, Qing and Sammons, Pam and Chen, Junjun},
+  journal={Leadership and Policy in Schools},
+  volume={17},
+  number={3},
+  pages={373--390},
+  year={2018},
+  publisher={Taylor \& Francis}
+}
+
+citation("here")
+
+citation("janitor")
+
+@article{kanevsky2013accelerating,
+  title={Accelerating Gifted Students in Canada: Policies and Possibilities.},
+  author={Kanevsky, Lannie S and Clelland, Debbie},
+  journal={Canadian Journal of Education},
+  volume={36},
+  number={3},
+  pages={229--271},
+  year={2013},
+  publisher={JSTOR}
+}
+
+@article{lambert2017public,
+  title={Public education: Reflections on fifteen years of public policy},
+  author={Lambert, Susan},
+  journal={BC Studies: The British Columbian Quarterly},
+  number={194},
+  pages={163--173},
+  year={2017}
+}
+
+@article{lo2019reenvisioning,
+  title={Reenvisioning Gifted Education in British Columbia: A Qualitative Research Protocol of Policy Analysis in the Context of a Redesigned Curriculum},
+  author={Lo, C Owen and Tan, Yuen Sze Michelle and Chrostowski, Megan and Hu, Shun-Fu and Chan, Diana and Sue, Deanna M and Wu, I-Chen and Li, Wei},
+  journal={International Journal of Qualitative Methods},
+  volume={18},
+  pages={1609406918822233},
+  year={2019},
+  publisher={SAGE Publications Sage CA: Los Angeles, CA}
+}
+
+@article{luke2009model,
+  title={A model for research-based state professional development policy},
+  author={Luke, Allan and McArdle, Felicity},
+  journal={Asia-Pacific Journal of Teacher Education},
+  volume={37},
+  number={3},
+  pages={231--251},
+  year={2009},
+  publisher={Taylor \& Francis}
+}
+
+@article{michelmore2017gap,
+  title={The gap within the gap: Using longitudinal data to understand income differences in educational outcomes},
+  author={Michelmore, Katherine and Dynarski, Susan},
+  journal={AERA Open},
+  volume={3},
+  number={1},
+  pages={2332858417692958},
+  year={2017},
+  publisher={SAGE Publications Sage CA: Los Angeles, CA}
+}
+
+@Manual{R-papaja,
+  author = {Frederik Aust and Marius Barth},
+  title = {{papaja}: {Create} {APA} manuscripts with {R Markdown}},
+  year = {2020},
+  note = {R package version 0.1.0.9997},
+  url = {https://github.com/crsh/papaja},
+}
+
+citation("readxl")
+
+@article{swanson2007policy,
+  title={Policy and practice: A case study of gifted education policy implementation},
+  author={Swanson, Julie Dingle},
+  journal={Journal for the Education of the Gifted},
+  volume={31},
+  number={2},
+  pages={131--164},
+  year={2007},
+  publisher={SAGE Publications Sage CA: Los Angeles, CA}
+}
+
+citation("tidyverse")

### 2020-12-07 scripts/Final_project.Rmd
ref2.bib Cross your fingers

--- 2020-12-07 scripts/Final_project.Rmd
@@
 
-bibliography      : ["r-references.bib"]
+bibliography      : "r-references.bib"
 
@@
 
+\newpage
 # References
 
-RStudio Team (2020). RStudio: Integrated Development Environment for R.
-  RStudio, PBC, Boston, MA URL http://www.rstudio.com/.
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup

### 2020-12-07 scripts/Final_project.tex
ref2.bib Cross your fingers

@@
 With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
-(1) Do student prevalence rates of disability or disorder decrease after the supreme court ruling of 2016 against existing trends?
-(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+
+\begin{enumerate}
+  \item Do student prevalence rates of disability or disorder decrease after the supreme court ruling of 2016 against existing trends?
+  \item Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
+ \end{enumerate}
 
@@
 \]
-\# Results
-Writing on results here\ldots{}
 
-First, we see that the BC trends do not show significant changes after the 2016 supreme court ruling.
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+To answer the first research question, we examine the provincial-level prevalence data by exceptionality to view trendlines over 15 years.
+It is apparent that 2016 is not a point of inflection for any of the twelve designations listed.
+Low-incidence exceptionalities remain consistent, and high-incidence exceptionalities are linearly (e.g., Learning Disabilities) or exponentially (e.g., Autism) increasing, and some high-incidence exceptionalities are linearly decreasing (e.g., Gifted).
+Due to the fact that there are no significant points of inflection observed on any of the 12 trendlines in figure 1, no further investigation (e.g., calculating effect size) was warranted.
+
 \includegraphics{Final_project_files/figure-latex/Provinical_Data_Overview-1.pdf}
@@
 This is roughly in line with Oregon's designation data, showing that there were no state-wide anomalies after the year 2016.
+Similar to the BC data, we see a steady increase in students identified under the autism diagnosis, while most categories remained relatively stable over all of the years surveyed.
+The recent growth in Developmental Delays and Emotional Disturbance would be worth watching in future years, but we can reasonably predict that more rigid categories like Deaf-Blindness and Traumatic Brain Injury will remain stable in the years to come.
+
+\begin{verbatim}
+## Warning: Removed 12 row(s) containing missing values (geom_path).
+\end{verbatim}
+
+\includegraphics{Final_project_files/figure-latex/Oregon plot-1.pdf}
+Looking at Oregon disability prevalence by age gives us some idea of what happens to these categories over time, and also invites some questions.
+Most visually striking is the drop-off of students identified as having a Speech or Language Impairment, which makes sense, as it is addressed early in school and highly correctable.
+By middle school, most students have been exited from the speech program.
+At a striking contrast would be things that cannot be corrected, such as Deaf-Blindness or Traumatic Brain Injury.
+As expected, these numbers are extremely static over time.
+More thought-provoking would be the numbers for Emotional Disturbance.
+These tend to grow fairly steadily during elementary and middle school, but around age 16 the numbers actually start to decrease.
+One might wonder if this drop is due to these students no longer being part of the school system, as they may be far more likely to drop out and/or enter a different system as adjudicated youth.
+It certainly would be worth an additional look in the future.
 
@@
 
-Then, we examine the rural versus urban trends and notice a few things\ldots{}
+Taken together, we see that at the provincial- and state-levels, there is consistency in identification of students with exceptionalities.
+It is clear that a notable inflection point in 2016 for the BC data would have represented a significant change from practice when comparing student identification practices in Oregon; however, there is no apparent inflection point in the BC data.
+
+To examine potential differences within BC, the authors chose to divide school districts by urban or rural designation according to Statistics Canada guidelines on characterizing living populations.
 
@@
 
+There are two major outcomes of interest observed with the data sets when disaggregated by rural and urban school districts.
+First, the trends of rural student exceptionality designations followed urban trends closely.
+The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (ps \textgreater{} 0.05).
+However, for diagnostically malleable categories (e.g., Moderate Behavioural designation) in rural school districts, 2016 was an inflection point for an increase in prevalence rates.
+When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (t(239) = -2.113, p = 0.036, g = 0.261, CI{[}-0.504, -0.019{]}), Learning Disability (t(262) = -1.984, p = 0.483, g = 0.230, CI{[}-0.453, -0.007{]}), and Chronic Health (t(255) = -2.068, p = 0.040, g = 0.247, CI{[}-0.479, -0.015{]}).
+Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges' g \textgreater{} 0.2, and g \textless{} 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI{[}-0.383, 0.144{]}).
+
 \hypertarget{discussion}{%
@@
 
-Regarding the first research question, it is apparent that the 2016 supreme court decision did not impact any of the 12 designation categories as hypothesized.
+Regarding the first research question, it is apparent that the 2016 Supreme Court decision did not impact any of the 12 designation categories as hypothesized.
 The trendlines for each of the exceptionality categories remained consistent post-2016 in BC and this mapped on to Oregon trends in exceptionality designation reasonably closely.
-This finding suggests that the resources and practices around supporting student
+This finding suggests that the resources and practices around supporting students with exceptionalities did not differ significantly from a provincial policy determination.
+
+Data on the second research question on rural versus urban school district response to the 2016 Supreme Court ruling suggests that the authors correctly hypothesized that there are differences for rural versus urban districts.
+Urban school districts largely followed provincial-level trends, and rural school districts had small effect sizes of increased prevalence for the post-2016 cohort of students.
+That is, when looking at all rural school districts and comparing two groups, four years preceding 2017 and three years 2017 through 2019, there is a significant increase in prevalence for designations in which clinicians have diagnostic flexibility (e.g., behavioural designations, temporary health conditions).
+This finding suggests that clinicians and administrators in rural school districts responded to the 2016 Supreme Court ruling by changing exceptionality designation practices that align with an increase in prevalence.
 
@@
 Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students.
-For example, Lo et al.~(2019) suggest that the British Columbia Ministry of Education's recent curriculum update makes amends for traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+For example, Lo et al.~(2019) suggest that the British Columbia Ministry of Education's recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+
+\newpage
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+\hypertarget{refs}{}
+
+\endgroup
 

### 2020-12-07 scripts/draft_play.tex
ref2.bib Cross your fingers

@@
 \hypersetup{
-  pdftitle={The title},
-  pdfauthor={Makayla Whitney1, Ernst-August Doelle1, Ernst-August Doelle1, \& Ernst-August Doelle1},
+  pdftitle={Trends in Pacific Northwest Special Education Population},
+  pdfauthor={David Fainstein1, Janette Avelar1, Makayla Whitney1, \& Joe Swinehart1},
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={Vancouver, British Columbia, Oregon, Special Education, SpEd},
   hidelinks,
@@
 \makeatother
-\shorttitle{Title}
-\keywords{keywords\newline\indent Word count: X}
+\shorttitle{PACNW SpEd Trends}
+\keywords{Vancouver, British Columbia, Oregon, Special Education, SpEd\newline\indent Word count: 0<n<1,000,000}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 
-\title{The title}
-\author{Makayla Whitney\textsuperscript{1}, Ernst-August Doelle\textsuperscript{1}, Ernst-August Doelle\textsuperscript{1}, \& Ernst-August Doelle\textsuperscript{1}}
+\title{Trends in Pacific Northwest Special Education Population}
+\author{David Fainstein\textsuperscript{1}, Janette Avelar\textsuperscript{1}, Makayla Whitney\textsuperscript{1}, \& Joe Swinehart\textsuperscript{1}}
 \date{}
@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+Janette Avelar, University of Oregon College of Education
+David Fainstein, University of Oregon College of Education
+Joe Swinehart, Behavioral Research \& Teaching
+Makayla Whitney, University of Oregon College of Education
 
-Enter author note here.
+Project completed using RStudio, GitKraken, GitHub, and papaja
+
+Contact: \href{mailto:dfainste@uoregon.edu}{\nolinkurl{dfainste@uoregon.edu}}
 
 The authors made the following contributions.
-Makayla Whitney: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing; Ernst-August Doelle: Writing - Review \& Editing.
+David Fainstein: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Janette Avelar: Writing - Review \& Editing; Makayla Whitney: Writing - Review \& Editing; Joe Swinehart: Writing - Review \& Editing.
 
-Correspondence concerning this article should be addressed to Makayla Whitney, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
+Correspondence concerning this article should be addressed to David Fainstein, 1 David Ave, Eugene, Oregon, USA. E-mail: \href{mailto:dfainste@uoregon.edu}{\nolinkurl{dfainste@uoregon.edu}}
 
@@
 \abstract{
-One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.
+This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
+The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministrty of Education in British Columbia.
+This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
+The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020.
+This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year.
+Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
 
@@
 Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
-The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon.
+The datasets detail the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon (OR).
 The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
@@
 
-\hypertarget{problem-statement-and-rqs}{%
-\section{Problem Statement and RQs}\label{problem-statement-and-rqs}}
-
-In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker \& Bergmann, 2013).
-While Canadian K-12 academic achievement outcomes are viewed as favourable on the world stage, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education.
-Significant events, such as the 2016 Canadian supreme court ruling that directly impacted British Columbia classrooms, have downstream effects on instructional interactions; yet a retroactive policy lens is rarely applied after such events occur.
-The current study seeks to address the knowledge gap of downstream instructional interact effects after a significant event in BC educational policy.
-
-With respect to the 2016 Supreme Court of Canada decision to revert BC classroom composition, size, and ratios for specialist teachers:
-(1) Do student prevalence rates of disability or disorder change after the supreme court ruling of 2016?
-(2) Are there different patterns for disability or disorder designation for rural versus urban school districts after the supreme court ruling in 2016?
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
+-- more introductory text here --
 Children designated with Special Needs categories have predominantly increased at different rates in British Columbia over time.
@@
 
+\begin{verbatim}
+## Warning: Problem with `mutate()` input `NUMBER_OF_STUDENTS`.
+## i NAs introduced by coercion
+## i Input `NUMBER_OF_STUDENTS` is `as.integer(NUMBER_OF_STUDENTS)`.
+\end{verbatim}
+
+\begin{verbatim}
+## Warning in mask$eval_all_mutate(dots[[i]]): NAs introduced by coercion
+\end{verbatim}
+
 \includegraphics{draft_play_files/figure-latex/sandbox-1.pdf}
+
+\begin{verbatim}
+## Warning in FUN(newX[, i], ...): NAs introduced by coercion
+\end{verbatim}
+
+\begin{verbatim}
+##  [1] "1996/1997" "1997/1998" "1998/1999" "1999/2000" "2000/2001" "2001/2002"
+##  [7] "2002/2003" "2003/2004" "2004/2005" "2005/2006" "2006/2007" "2007/2008"
+## [13] "2008/2009" "2009/2010" "2010/2011" "2011/2012" "2012/2013" "2013/2014"
+## [19] "2014/2015" "2015/2016" "2016/2017" "2017/2018" "2018/2019" "2019/2020"
+\end{verbatim}
+
+\begin{verbatim}
+## Warning in FUN(newX[, i], ...): NAs introduced by coercion
+\end{verbatim}
 
@@
 
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:District Census Population Table}District Classification by Population According to 2016 Census}
+
+\begin{tabular}{lllll}
+\toprule
+District Number & District Name & Population & Classification & Number of Students\\
+\midrule
+022 & Vernon & 61,334 & Rural & 16,768\\
+023 & Central Okanagan & 194,882 & Urban & 43,609\\
+028 & Quesnel & 23,146 & Rural & 9,036\\
+033 & Chilliwack & 101,512 & Urban & 28,997\\
+034 & Abbotsford & 180,518 & Urban & 38,369\\
+035 & Langley & 2,463,431 & Urban & 43,043\\
+036 & Surrey & 2,463,431 & Urban & 143,222\\
+037 & Delta & 2,463,431 & Urban & 43,897\\
+038 & Richmond & 2,463,431 & Urban & 40,221\\
+039 & Vancouver & 2,463,431 & Urban & 132,584\\
+040 & New Westminster & 2,463,431 & Urban & 11,714\\
+041 & Burnaby & 2,463,431 & Urban & 47,952\\
+042 & Maple Ridge-Pitt Meadows & 2,463,431 & Urban & 33,633\\
+043 & Coquitlam & 2,463,431 & Urban & 88,291\\
+044 & North Vancouver & 2,463,431 & Urban & 39,657\\
+045 & West Vancouver & 2,463,431 & Urban & 11,732\\
+046 & Sunshine Coast & 2,463,431 & Urban & 13,283\\
+047 & Powell River & 16,783 & Rural & 6,816\\
+052 & Prince Rupert & 12,687 & Rural & 7,222\\
+053 & Okanagan Similkameen & 43,432 & Rural & 6,596\\
+057 & Prince George & 86,622 & Rural & 39,443\\
+061 & Greater Victoria & 367,770 & Urban & 50,051\\
+062 & Sooke & 367,770 & Urban & 24,126\\
+063 & Saanich & 367,770 & Urban & 19,533\\
+067 & Okanagan Skaha & 43,432 & Rural & 19,156\\
+068 & Nanaimo-Ladysmith & 104,936 & Urban & 30,870\\
+070 & Port Alberni & 25,112 & Rural & 11,654\\
+073 & Kamloops-Thompson & 103,811 & Urban & 32,620\\
+075 & Mission & 33,261 & Rural & 13,858\\
+079 & Cowichan Valley & 44,451 & Rural & 21,479\\
+081 & Fort Nelson & 18,307 & Rural & 1,319\\
+083 & North Okanagan-Shuswap & 43,432 & Rural & 23,720\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Number of students is estimated and does not include masked data.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
 This table displays the census results from 2011 and 2016 for our school districts.
@@
 In 2011, they were rural, but for our dataset they have been classified as urban due to their population increase in 2016.
+
+\begin{verbatim}
+## Warning: Removed 129 row(s) containing missing values (geom_path).
+\end{verbatim}
+
+\includegraphics{draft_play_files/figure-latex/rural play-1.pdf} \includegraphics{draft_play_files/figure-latex/rural play-2.pdf} \includegraphics{draft_play_files/figure-latex/rural play-3.pdf}
+
+\begin{verbatim}
+## Warning: Removed 12 rows containing missing values (position_stack).
+\end{verbatim}
+
+\includegraphics{draft_play_files/figure-latex/rural play-4.pdf} \includegraphics{draft_play_files/figure-latex/rural play-5.pdf}
+
+\begin{verbatim}
+## geom_col: width = NULL, na.rm = FALSE
+## stat_identity: na.rm = FALSE
+## position_stack
+\end{verbatim}
+
+\includegraphics{draft_play_files/figure-latex/urban play-1.pdf}
+
+\begin{verbatim}
+## Warning: Removed 25 row(s) containing missing values (geom_path).
+\end{verbatim}
+
+\includegraphics{draft_play_files/figure-latex/urban play-2.pdf}
 
@@
 
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
+
+\hypertarget{material}{%
+\subsection{Material}\label{material}}
+
 \hypertarget{procedure}{%
@@
 
-We used R (Version 4.0.2; R Core Team, 2020) and the R-packages \emph{dplyr} (Version 1.0.2; Wickham et al., 2020), \emph{forcats} (Version 0.5.0; Wickham, 2020a), \emph{ggplot2} (Version 3.3.2; Wickham, 2016), \emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 2.0.1; Firke, 2020), \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2020), \emph{purrr} (Version 0.3.4; Henry \& Wickham, 2020), \emph{readr} (Version 1.3.1; Wickham, Hester, \& Francois, 2018), \emph{readxl} (Version 1.3.1; Wickham \& Bryan, 2019), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version 3.0.3; Müller \& Wickham, 2020), \emph{tidyr} (Version 1.1.2; Wickham, 2020b), and \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al., 2019) for all our analyses.
+We used R (Version 4.0.3; R Core Team, 2020) and the R-packages \emph{dplyr} (Version 1.0.2; Wickham et al., 2020), \emph{forcats} (Version 0.5.0; Wickham, 2020a), \emph{ggplot2} (Version 3.3.2; Wickham, 2016), \emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 2.0.1; Firke, 2020), \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2020), \emph{purrr} (Version 0.3.4; Henry \& Wickham, 2020), \emph{readr} (Version 1.4.0; Wickham \& Hester, 2020), \emph{readxl} (Version 1.3.1; Wickham \& Bryan, 2019), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tibble} (Version 3.0.4; Müller \& Wickham, 2020), \emph{tidyr} (Version 1.1.2; Wickham, 2020b), and \emph{tidyverse} (Version 1.3.0; Wickham, Averick, et al., 2019) for all our analyses.
 
-\hypertarget{results-1}{%
-\section{Results}\label{results-1}}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 \leavevmode\hypertarget{ref-R-readr}{}%
-Wickham, H., Hester, J., \& Francois, R. (2018). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
+Wickham, H., \& Hester, J. (2020). \emph{Readr: Read rectangular text data}. Retrieved from \url{https://CRAN.R-project.org/package=readr}
 

### 2020-12-07 scripts/j_play.Rmd
ref2.bib Cross your fingers

@@
 
-#Ignore
+unique(district_class$DISTRICT_NAME)
+# Table showing BC total needs designation by year
+urban_needs_data <- urban_districts %>% 
+  group_by(school_year, special_needs_code) %>% 
+  mutate(total = sum(number_of_students, na.rm = TRUE)) %>% 
+    select(-c("district_number",
+            "district_name",
+            "district_classification",
+            "number_of_students")) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = special_needs_code,
+              values_from = total)
+
+apa_table(
+  urban_needs_data,
+  caption = "Total BC Students Eligible for Special Education Services",
+  note = "NA's indicate masked data where students eligible were less than 10.",
+  col.names = c("School Year",
+                "Physically Dependent",
+                "Deafblind",
+                "Moderate to Profound Intellectual Disability",
+                "Physical Disability or Chronic Health Impairment",
+                "Visual Impairment",
+                "Deaf or Hard of Hearing",
+                "Autism Spectrum Disorder",))
 
@@
             "district_population_2016"))
-
 #Viz - Totals by Code
@@
 rural_totals <- rural_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
+  group_by(school_year, district_name) %>% 
   mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students, district_number)) %>% 
   distinct()
+
+#This table is not in final draft, it shows rural district totals by year.
+rural_totals_by_year <- rural_totals %>% 
+  pivot_wider(names_from = "district_name",
+              values_from = "students")
 
@@
     theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+apa_table(pg_gifted,
+          caption = "Prince George Gifted Student Classification 1999-2020",
+          note = "Data masked when reflecting less than 10 students.")
 ```

### 2020-12-07 scripts/Final_project.Rmd
Correlogram finalized w/ text blurb

@@
 
-#best one yet! Box-and-whisker plot
-ggpairs(g_vic,
-        upper = list(continuous = ggally_dot_and_box, combo = ggally_box_no_facet),
-        lower = list (continuous = "smooth",
-                      combo = "facetdensity",
-                      mapping = aes(color = SPECIAL_NEEDS_CODE)))
-
-#let's try bar graph
+#bar graph added
 ggpairs(g_vic,
@@
 
-ggally_dot_and_box(g_vic, 
-                   mapping = ggplot2::aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS),
-                   boxPlot = TRUE) +
-  theme(axis.text.x = element_blank())
+```
 
-```
+The above correlogram isolates a single school district - Greater Victoria - as a representation of the trends observed within the whole data set.
+This series of plots gives a snapshot introduction for the data trends to be observed.
 

### 2020-12-07 scripts/Final_project.Rmd
Text Push

@@
 abstract: |
-  This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
-The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministry of Education in British Columbia.
-This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
-The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020.
-This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year.
-Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
+  Over the last 20 years, there have been drastic changes within public school districts in regards to special education classifications.
+By examining and comparing special education data spanning over 15 years in Vancouver, British Columbia and Oregon, policy changes can be visualized in student population data.
+Descriptive statistics are used to analyze trends seen both across special education categories, school years, and districts.
+Districts are classified as urban or rural and comparative analysis tracks trends observed.
+Special attention is given to any population or categorization changes occurring past the 2016 Supreme Court of Canada decision to revert BC classroom composition.
+Data visualization allows for large datasets, such as these, to be compacted into thorough plots where longitudinal trends can be observed and analyzed.
 
 keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
-wordcount         : "0<n<1,000,000"
+wordcount         :
 
@@
 
-floatsintext      : no
-figurelist        : no
-tablelist         : no
+floatsintext      : yes
+figurelist        : yes
+tablelist         : yes
 footnotelist      : no
@@
 BCdatastripped <- BCdata %>%
-    #Makayla & Janette - may want to include more columns here with the select for our global "BCdatastripped" df
+
     select("SCHOOL_YEAR", 
@@
 Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017).
-Canadian supreme court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire.
+Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire.
 The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change.
@@
 Together, these datasets help achieve three objectives for acquiring data appropriate to the research questions:
-Headcount data at multiple levels (e.g., district and state or province) for students with exceptionalities, disaggregated by category of exceptionality, year, and location.
-Population data to fit school district locations to a geographical classification (e.g., urban, suburban, rural) as specified in national documentation (e.g., urban centers have a minimum of 10,000 inhabitants based on census data), and calculate prevalence rates.
-Data from multiple jurisdictions where educational policies do not align chronologically.
+*Headcount data at multiple levels (e.g., district and state or province) for students with exceptionalities, disaggregated by category of exceptionality, year, and location.
+*Population data to fit school district locations to a geographical classification (e.g., urban, suburban, rural) as specified in national documentation (e.g., urban centers have a minimum of 10,000 inhabitants based on census data), and calculate prevalence rates.
+*Data from multiple jurisdictions where educational policies do not align chronologically.
 For example, special education law in the United States is disjointed from BC special education law and policy changes are not the same, nor are they implemented simultaneously.
@@
 The recent growth in Developmental Delays and Emotional Disturbance would be worth watching in future years, but we can reasonably predict that more rigid  categories like Deaf-Blindness and Traumatic Brain Injury will remain stable in the years to come.
+
 ```{r Oregon plot}
@@
 ```
+
 Looking at Oregon disability prevalence by age gives us some idea of what happens to these categories over time, and also invites some questions.
@@
 
+The district classification data was scaled down to include only public schools, while excluding private institutions.
+The school districts of Southeast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gwaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga'a, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information.
+The urban vs.
+rural classifications were made based on the district's population on the 2016 census.
+If the population was above 100,000 individuals, it is classified as urban.
+If the population was below 99,999 individuals, then it is classified as rural.
+
+Disaggregating districts by urban and rural classification allowed for analysis across geographic designation and within case-study districts for further understanding of policy effects on regions with differing resources and large differences in student population sizes.
+The table below details our classification of districts, according to the 2016 census.
+Classification and population is shown in comparison to the number of students receiving special education services for the 2016/2017 school year in that school district.
+
 ``` {r District Census Population Table}
@@
 
-```{r rural play}
+A closer look at the districts in play in figure (Rainbow urban) details how overall urban areas appear more stable over time compared to urban areas, as in figure (Rainbow rural), which appear to have clustering trends.
+Between 1999 and 2004/2005 the number of students with exceptionalities appear fairly stable.
+Rural districts then suffer a noticeable drop between 2004/2005 and 2005/2006, specifically in the district of Prince George.
+The years following present a steady decline through 2016, which then begins a steady upward trajectory.
+Both urban and rural areas alike exhibit a similar dramatic rise between 1998/1999 and 1999/2000, which we attribute to data reporting changes, but may be worth a closer look.
+
+There are two major outcomes of interest observed with the data sets when disaggregated by rural and urban school districts.
+First, the trends of rural student exceptionality designations followed urban trends closely.
+The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (ps > 0.05).
+However, for diagnostically malleable categories (e.g., Moderate Behavioural designation) in rural school districts, 2016 was an inflection point for an increase in prevalence rates.
+When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (t(239) = -2.113, p = 0.036, g = 0.261, CI[-0.504, -0.019]), Learning Disability (t(262) = -1.984, p = 0.483, g = 0.230, CI[-0.453, -0.007]), and Chronic Health (t(255) = -2.068, p = 0.040, g = 0.247, CI[-0.479, -0.015]).
+Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges’ g > 0.2, and g < 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI[-0.383, 0.144]).
+
+```{r rainbow figure rural}
 #Data Prep
@@
 
-#Viz - Totals by Code
+#Viz - Rainbow bar chart of rural totals by exceptionality
 ggplot(rural_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Total Students by Exceptionality - Rural",
+       x = "Exceptionality",
+       y = "Number of Students")
+```
+
+```{r rainbow figure urban}
+#Data Prep
+urban_districts <- district_class %>% 
+  janitor::clean_names() %>% 
+  filter(district_classification == "Urban") %>% 
+  select(-c("data_level",
+            "public_or_independent",
+            "district_population_2016"))
+
+#Viz - Rainbow bar chart of urban totals by exceptionality
+ggplot(urban_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
+```
+
+Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures x and x below.
+Of note, the “Gifted” designation suffers a severe drop around 2005/2006 for urban and rural areas alike, though the change is more pronounced within rural districts.
+
+```{r rural facet bars}
+#Data for Rural Totals (No categories)
+rural_totals <- rural_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#A Vizualization of Totals/Time for Rural areas
+ggplot(rural_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
+```
+
+```{r urban facet bars}
+#Data Prep for Urban totals (no categories)
+urban_totals <- urban_districts %>% 
+  group_by(school_year, district_name, district_number) %>% 
+  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
+  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
+  distinct()
+
+#A visualization of Totals/Time for Urban areas
+ggplot(urban_districts, 
        aes(special_needs_code, number_of_students,
@@
 
-#Data for Rural Totals (No categories)
-rural_totals <- rural_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
+```
 
-#Data Viz Totals/Time
-ggplot(rural_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
+The significant drop presents a second major outcome of interest, from which a posteriori visual analysis served as a basis for understanding notable trends.
+For example, multiple rural school districts reported a head count of more than 1,000 gifted students in 2004, and then by 2008 the districts reported fewer than 10 gifted students.
+This magnitude of declining gifted student designations is significant and specific to rural school districts in BC.
+Of particular note is the district of Prince George, where gifted students appeared to have been more
+
+```{r rural play}
 
@@
 
-```{r urban play}
-#Data Prep
-urban_districts <- district_class %>% 
-  janitor::clean_names() %>% 
-  filter(district_classification == "Urban") %>% 
-  select(-c("data_level",
-            "public_or_independent",
-            "district_population_2016"))
-
-urban_totals <- urban_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#Data Viz Totals/Time
-ggplot(urban_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
-
-#Viz - Totals by Code
-ggplot(urban_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
-       x = "Exceptionality",
-       y = "Number of Students")
-```
-
-There are two major outcomes of interest observed with the data sets when disaggregated by rural and urban school districts.
-First, the trends of rural student exceptionality designations followed urban trends closely.
-The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (ps > 0.05).
-However, for diagnostically malleable categories (e.g., Moderate Behavioural designation) in rural school districts, 2016 was an inflection point for an increase in prevalence rates.
-When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (t(239) = -2.113, p = 0.036, g = 0.261, CI[-0.504, -0.019]), Learning Disability (t(262) = -1.984, p = 0.483, g = 0.230, CI[-0.453, -0.007]), and Chronic Health (t(255) = -2.068, p = 0.040, g = 0.247, CI[-0.479, -0.015]).
-Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges’ g > 0.2, and g < 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI[-0.383, 0.144]).
-
 # Discussion

### 2020-12-07 scripts/Final_project.Rmd
Editing and rearranging urban/rural chunks

@@
 ```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
+knitr::opts_chunk$set(warning = FALSE)
 library("papaja")
@@
    facet_wrap(~SPECIAL_NEEDS_CODE) +
-   labs(title = "BC Provincial Prevalence of Students with Exceptionalities by Category",
+   labs(title = "Fig. 1 BC Provincial Prevalence of Students with Exceptionalities by Category",
         x = "School Year",
@@
   facet_wrap(~year)+
-    labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
+    labs(color="disability", x= "Student Age", y="Student Count", title ="Fig. 2 Disability Prevalence By Age")
 
@@
  facet_wrap(~disability)+
-    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")
+    labs(color="frequency", x= "Year", y="Student Count", title = "Fig. 3 Oregon Headcount of Students with Exceptionalities by Category")
 
@@
   geom_smooth(data = BCAutism, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS, color = "British Columbia"), se = FALSE) +
-  labs(title = "BC and Oregon Headcount of Students with Chronic Health Conditions",
+  labs(title = "Fig. 4 BC and Oregon Headcount of Students with Chronic Health Conditions",
         x = "School Year",
@@
 
-A closer look at the districts in play in figure (Rainbow urban) details how overall urban areas appear more stable over time compared to urban areas, as in figure (Rainbow rural), which appear to have clustering trends.
+A closer look at the districts in Figure 5 details how overall urban areas appear more stable over time compared to urban areas, as in figure 6, which appear to have clustering trends.
 Between 1999 and 2004/2005 the number of students with exceptionalities appear fairly stable.
@@
 
-```{r rainbow figure rural}
-#Data Prep
+```{r urban and rural data}
+#Data for Rural Totals (No categories)
 rural_districts <- district_class %>% 
@@
 
-#Viz - Rainbow bar chart of rural totals by exceptionality
-ggplot(rural_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality - Rural",
-       x = "Exceptionality",
-       y = "Number of Students")
-```
-
-```{r rainbow figure urban}
-#Data Prep
+#Data Prep for Urban totals (no categories)
 urban_districts <- district_class %>% 
@@
             "district_population_2016"))
-
-#Viz - Rainbow bar chart of urban totals by exceptionality
-ggplot(urban_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Urban Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
-```
-
-Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures x and x below.
-Of note, the “Gifted” designation suffers a severe drop around 2005/2006 for urban and rural areas alike, though the change is more pronounced within rural districts.
-
-```{r rural facet bars}
-#Data for Rural Totals (No categories)
+#Data Prep
 rural_totals <- rural_districts %>% 
@@
 
-#A Vizualization of Totals/Time for Rural areas
-ggplot(rural_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Students with Exceptionalities in Rural Districts 1996-2020",
-       x = "School Year",
-       y = "Total Students")
-```
-
-```{r urban facet bars}
-#Data Prep for Urban totals (no categories)
+#Data Prep
 urban_totals <- urban_districts %>% 
@@
   distinct()
+```
 
+```{r rainbow figure urban}
+#Viz - Rainbow bar chart of urban totals by exceptionality
+ggplot(urban_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Fig. 5 Students with Exceptionalities in Urban Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
+```
+
+```{r rainbow figure rural}
+#Viz - Rainbow bar chart of rural totals by exceptionality
+ggplot(rural_totals, aes(school_year, students)) +
+  geom_col(aes(color = district_name)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "Fig. 6 Students with Exceptionalities in Rural Districts 1996-2020",
+       x = "School Year",
+       y = "Total Students")
+```
+
+Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures 7 and 8 below.
+Of note, the “Gifted” designation suffers a severe drop around 2005/2006 for urban and rural areas alike, though the change is more pronounced within rural districts.
+
+```{r rural facet bars}
+#A Vizualization of Totals/Time for Rural areas
+ggplot(rural_districts, 
+       aes(special_needs_code, number_of_students,
+           color = special_needs_code)) +
+  geom_line() +
+  facet_wrap(~school_year) +
+  labs(title = "Fig. 7 Total Students by Exceptionality - Rural",
+       x = "Exceptionality",
+       y = "Number of Students")
+```
+
+```{r urban facet bars}
 #A visualization of Totals/Time for Urban areas
@@
   facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
+  labs(title = "Fig. 8 Total Students by Exceptionality",
        x = "Exceptionality",
@@
 This magnitude of declining gifted student designations is significant and specific to rural school districts in BC.
-Of particular note is the district of Prince George, where gifted students appeared to have been more
+Of particular note is the district of Prince George, where gifted students appear to have accounted for the majority of overall students with exceptionalities until 2006.
 
-```{r rural play}
-
-#What's going on with Prince George? Why is there such a steep drop in 2006?
-pg <- rural_districts %>% 
-  filter(district_name == "Prince George") %>% 
-  as_tibble()
-
-pg_sums <- pg %>% 
-  group_by(school_year) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#What's happening by Dx?
-pg_dx_sum <- pg %>% 
-  group_by(school_year, special_needs_code) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, district_classification, number_of_students)) %>% 
-  distinct()
-
-ggplot(pg_dx_sum, aes(school_year, student)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  facet_wrap(~special_needs_code) +
-  labs(title = "Students with Exceptionalities in Prince George 1999-2020",
-       x = "School Year",
-       y = "Number of Students")
-# Need some explanation here for explaining facet wrap
-
-#What's going on with gifted?????
-pg_gifted <- pg %>% 
-  filter(special_needs_code == "P") %>% 
-  select(-c(district_number, district_classification))
-
-ggplot(pg_gifted, aes(school_year, number_of_students)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  labs(title = "Gifted Students in Prince George 1999-2020",
-       subtitle = "Note: Totals under 10 are masked by district in reporting.",
-       x = "School Year",
-       y = "Total Students")
+```{r rural play, echo = FALSE}
 

### 2020-12-07 scripts/Final_project.tex
Editing and rearranging urban/rural chunks

@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \shorttitle{PACNW SpEd Trends}
-\keywords{Vancouver, British Columbia, Oregon, Special Education, SpEd\newline\indent Word count: 0<n<1,000,000}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
+\keywords{Vancouver, British Columbia, Oregon, Special Education, SpEd}
 \usepackage{csquotes}
+\usepackage[titles]{tocloft}
+\cftpagenumbersoff{figure}
+\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
+\renewcommand{\cftfigaftersnum}{.\space}
+\setlength{\cftfigindent}{0pt}
+\setlength{\cftafterloftitleskip}{0pt}
+\settowidth{\cftfignumwidth}{Figure 10.\qquad}
+\cftpagenumbersoff{table}
+\renewcommand{\cfttabpresnum}{\itshape\tablename\enspace}
+\renewcommand{\cfttabaftersnum}{.\space}
+\setlength{\cfttabindent}{0pt}
+\setlength{\cftafterloftitleskip}{0pt}
+\settowidth{\cfttabnumwidth}{Table 10.\qquad}
 \ifxetex
@@
 \abstract{
-This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
-The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministry of Education in British Columbia.
-This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
-The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020.
-This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year.
-Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
+Over the last 20 years, there have been drastic changes within public school districts in regards to special education classifications.
+By examining and comparing special education data spanning over 15 years in Vancouver, British Columbia and Oregon, policy changes can be visualized in student population data.
+Descriptive statistics are used to analyze trends seen both across special education categories, school years, and districts.
+Districts are classified as urban or rural and comparative analysis tracks trends observed.
+Special attention is given to any population or categorization changes occurring past the 2016 Supreme Court of Canada decision to revert BC classroom composition.
+Data visualization allows for large datasets, such as these, to be compacted into thorough plots where longitudinal trends can be observed and analyzed.
 }
@@
 
+The above correlogram isolates a single school district - Greater Victoria - as a representation of the trends observed within the whole data set.
+This series of plots gives a snapshot introduction for the data trends to be observed.
+
 \hypertarget{current-study}{%
@@
 Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017).
-Canadian supreme court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers' that school districts are required to hire.
+Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers' that school districts are required to hire.
 The driving theory of change is that decreasing student-to-teacher ratios and capping the number of students with exceptionalities in each classroom causes student designation practices to change.
@@
 Together, these datasets help achieve three objectives for acquiring data appropriate to the research questions:
-Headcount data at multiple levels (e.g., district and state or province) for students with exceptionalities, disaggregated by category of exceptionality, year, and location.
-Population data to fit school district locations to a geographical classification (e.g., urban, suburban, rural) as specified in national documentation (e.g., urban centers have a minimum of 10,000 inhabitants based on census data), and calculate prevalence rates.
-Data from multiple jurisdictions where educational policies do not align chronologically.
+\emph{Headcount data at multiple levels (e.g., district and state or province) for students with exceptionalities, disaggregated by category of exceptionality, year, and location.
+}Population data to fit school district locations to a geographical classification (e.g., urban, suburban, rural) as specified in national documentation (e.g., urban centers have a minimum of 10,000 inhabitants based on census data), and calculate prevalence rates.
+*Data from multiple jurisdictions where educational policies do not align chronologically.
 For example, special education law in the United States is disjointed from BC special education law and policy changes are not the same, nor are they implemented simultaneously.
@@
 
-\begin{verbatim}
-## Warning: Removed 12 row(s) containing missing values (geom_path).
-\end{verbatim}
+\includegraphics{Final_project_files/figure-latex/Oregon plot-1.pdf}
 
-\includegraphics{Final_project_files/figure-latex/Oregon plot-1.pdf}
 Looking at Oregon disability prevalence by age gives us some idea of what happens to these categories over time, and also invites some questions.
@@
 It certainly would be worth an additional look in the future.
-
-\begin{verbatim}
-## Warning: Removed 190 rows containing non-finite values (stat_smooth).
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-## Warning in FUN(newX[, i], ...): NAs introduced by coercion
-\end{verbatim}
+The district classification data was scaled down to include only public schools, while excluding private institutions.
+The school districts of Southeast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gwaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga'a, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information.
+The urban vs.~rural classifications were made based on the district's population on the 2016 census.
+If the population was above 100,000 individuals, it is classified as urban.
+If the population was below 99,999 individuals, then it is classified as rural.
+
+Disaggregating districts by urban and rural classification allowed for analysis across geographic designation and within case-study districts for further understanding of policy effects on regions with differing resources and large differences in student population sizes.
+The table below details our classification of districts, according to the 2016 census.
+Classification and population is shown in comparison to the number of students receiving special education services for the 2016/2017 school year in that school district.
 
@@
 
-\begin{verbatim}
-## Warning: Removed 129 row(s) containing missing values (geom_path).
-\end{verbatim}
-
-\includegraphics{Final_project_files/figure-latex/rural play-1.pdf} \includegraphics{Final_project_files/figure-latex/rural play-2.pdf} \includegraphics{Final_project_files/figure-latex/rural play-3.pdf}
-
-\begin{verbatim}
-## Warning: Removed 12 rows containing missing values (position_stack).
-\end{verbatim}
-
-\includegraphics{Final_project_files/figure-latex/rural play-4.pdf} \includegraphics{Final_project_files/figure-latex/rural play-5.pdf}
-
-\begin{verbatim}
-## geom_col: width = NULL, na.rm = FALSE
-## stat_identity: na.rm = FALSE
-## position_stack
-\end{verbatim}
-
-\includegraphics{Final_project_files/figure-latex/urban play-1.pdf}
-
-\begin{verbatim}
-## Warning: Removed 25 row(s) containing missing values (geom_path).
-\end{verbatim}
-
-\includegraphics{Final_project_files/figure-latex/urban play-2.pdf}
+A closer look at the districts in Figure 5 details how overall urban areas appear more stable over time compared to urban areas, as in figure 6, which appear to have clustering trends.
+Between 1999 and 2004/2005 the number of students with exceptionalities appear fairly stable.
+Rural districts then suffer a noticeable drop between 2004/2005 and 2005/2006, specifically in the district of Prince George.
+The years following present a steady decline through 2016, which then begins a steady upward trajectory.
+Both urban and rural areas alike exhibit a similar dramatic rise between 1998/1999 and 1999/2000, which we attribute to data reporting changes, but may be worth a closer look.
 
@@
 Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges' g \textgreater{} 0.2, and g \textless{} 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI{[}-0.383, 0.144{]}).
+
+\includegraphics{Final_project_files/figure-latex/rainbow figure urban-1.pdf}
+
+\includegraphics{Final_project_files/figure-latex/rainbow figure rural-1.pdf}
+
+Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures x and x below.
+Of note, the \enquote{Gifted} designation suffers a severe drop around 2005/2006 for urban and rural areas alike, though the change is more pronounced within rural districts.
+
+\includegraphics{Final_project_files/figure-latex/rural facet bars-1.pdf}
+
+\includegraphics{Final_project_files/figure-latex/urban facet bars-1.pdf}
+
+The significant drop presents a second major outcome of interest, from which a posteriori visual analysis served as a basis for understanding notable trends.
+For example, multiple rural school districts reported a head count of more than 1,000 gifted students in 2004, and then by 2008 the districts reported fewer than 10 gifted students.
+This magnitude of declining gifted student designations is significant and specific to rural school districts in BC.
+Of particular note is the district of Prince George, where gifted students appear to have accounted for the majority of overall students with exceptionalities until 2006.
+
+\includegraphics{Final_project_files/figure-latex/rural play-1.pdf} \includegraphics{Final_project_files/figure-latex/rural play-2.pdf} \includegraphics{Final_project_files/figure-latex/rural play-3.pdf}
+
+\begin{verbatim}
+## geom_col: width = NULL, na.rm = FALSE
+## stat_identity: na.rm = FALSE
+## position_stack
+\end{verbatim}
 
@@
 
+\clearpage
+\renewcommand{\listfigurename}{Figure captions}
+
+\clearpage
+\renewcommand{\listtablename}{Table captions}
+
 \end{document}

### 2020-12-07 scripts/Final_project.Rmd
Final BC table by designation - select years

@@
         y = "Prevalence Rate")
+```
+
+```{r Provincial Data Table}
+# A table showing special need designation by year in BC.
+need_des <- BCdata %>% 
+  janitor::clean_names() %>% 
+  select(-c(data_level,
+            public_or_independent,
+            district_number,
+            district_name,
+            school_number,
+            school_name,
+            special_needs_code)) %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(number_of_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = school_year,
+              values_from = number_of_students) %>% 
+  select(c("1996/1997",
+           "2001/2002",
+           "2005/2006",
+           "2010/2011",
+           "2015/2016",
+           "2019/2020")) %>% 
+  rename("Special Needs" = special_needs_desc)
+
+apa_table(need_des,
+          caption = "Special Needs Designation by Select Years (BC)",
+          note = "NAs indicate masked data where students <10")
 ```

### 2020-12-07 scripts/draft_play.Rmd
Final BC table by designation - select years

@@
 #This is a work in progress no touchy
+#Forget this chunk BLAH
 
@@
 unique(count_by_year$school_year)
+```
+
+```{r Provincial Data Table}
+# A table showing special need designation by year in BC.
+need_des <- BCdata %>% 
+  janitor::clean_names() %>% 
+  select(-c(data_level,
+            public_or_independent,
+            district_number,
+            district_name,
+            school_number,
+            school_name,
+            special_needs_code)) %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(number_of_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = school_year,
+              values_from = number_of_students) %>% 
+  select(c("1996/1997",
+           "2001/2002",
+           "2005/2006",
+           "2010/2011",
+           "2015/2016",
+           "2019/2020")) %>% 
+  rename("Special Needs" = special_needs_desc)
+
+apa_table(need_des,
+          caption = "Special Needs Designation by Select Years (BC)",
+          note = "NAs indicate masked data where students <10")
 ```

### 2020-12-07 scripts/j_play.Rmd
Final BC table by designation - select years

@@
 #unique(count_by_year$school_year)
+
+# A table showing special need designation by year in BC.
+need_des <- BCdata %>% 
+  janitor::clean_names() %>% 
+  select(-c(data_level,
+            public_or_independent,
+            district_number,
+            district_name,
+            school_number,
+            school_name,
+            special_needs_code)) %>% 
+  group_by(school_year, special_needs_desc) %>% 
+  mutate(number_of_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  distinct() %>% 
+  pivot_wider(names_from = school_year,
+              values_from = number_of_students) %>% 
+  select(c("1996/1997",
+           "2001/2002",
+           "2005/2006",
+           "2010/2011",
+           "2015/2016",
+           "2019/2020")) %>% 
+  rename("Special Needs" = special_needs_desc)
+
+apa_table(need_des,
+          caption = "Special Needs Designation by Select Years (BC)",
+          note = "NAs indicate masked data where students <10")
+
+as_tibble(BCdata)
 ```
@@
 #Data Prep
-census_data <- district_class %>% 
-  janitor::clean_names() %>% 
-  group_by(district_name, district_number, school_year) %>% 
-  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c("school_year",
-            "data_level",
-            "public_or_independent",
-            "special_needs_code",
-            "number_of_students")) %>% 
-  distinct()
+#census_data <- district_class %>% 
+ # janitor::clean_names() %>% 
+  #group_by(district_name, district_number, school_year) %>% 
+  #mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
+  #select(-c("school_year",
+  #          "data_level",
+   #         "public_or_independent",
+    #        "special_needs_code",
+     #       "number_of_students")) %>% 
+  #distinct()
 
 #Table
-apa_table(
-  census_data,
-    digits = 0,
-  caption = "District Classification by Population According to 2016 Census",
-  note = "Number of students does not include masked data.",
-  col.names = c("District Number",
-            "District Name",
-            "Population",
-            "Classification",
-            "Number of Students")
-)
+#apa_table(
+#  census_data,
+#    digits = 0,
+#  caption = "District Classification by Population According to 2016 Census",
+#  note = "Number of students does not include masked data.",
+#  col.names = c("District Number",
+#            "District Name",
+#            "Population",
+#            "Classification",
+#            "Number of Students")
+#)
 

### 2020-12-07 scripts/Final_project.Rmd
References push for inline code

@@
 
-bibliography      : "r-references.bib"
+bibliography      : ["r-references.bib"]
 
@@
 # Introduction
-Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society (Ball, 1998).
+Educational policy can impact practice by reorganizing educational delivery and reconceptualizing the meaning and impact of education on society [@ball_1998].
 Educational policy exists as a mechanism to drive changes in school buildings that result in socially important outcomes (e.g., increasing literacy acquisition through improved instruction, reducing instances of bullying through prevention programs).
-However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical (Cohen et al., 2009).
-Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students (Swanson, 2007).
+However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical [@cohen_2009].
+Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students [@swanson_2007].
 
-When policy is enacted, the intended outcomes are often not realized for months, years, or decades (Gu et al., 2018).
-For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke & McArdle, 2009).
-Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest (Michelmore & Dynarski, 2017).
+When policy is enacted, the intended outcomes are often not realized for months, years, or decades [@gu-2018].
+For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores [@luke_2009].
+Longitudinal data helps to understand the impact of policy change because of the year-over-year comparisons that illustrate differences, inflection points, or anomalies of interest [@michelmore_2017].
 
@@
 The initial analyses will inform a discussion on developmental trajectories by illustrating shifts in prevalence growth from static to linearly increasing based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data (e.g., common age of onset for diagnosis per category and diagnostic patterns over time).
-In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM; American Psychological Association, 2013), in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).
+In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders [DSM, @american_2013], in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).
 
-Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017).
+Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC [see @lambert_2017].
 Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers’ that school districts are required to hire.
@@
 ## Problem Statement and RQs
-In comparison to the United States, Canadian education policy receives little attention and scholarly interest (Walker & Bergmann, 2013).
+In comparison to the United States, Canadian education policy receives little attention and scholarly interest [@walker_2013].
 While Canadian K-12 academic achievement outcomes are generally higher than a national-level comparison with the US, there are ongoing policy issues to address when provincial ministries of education are crafting legislation and procedures to impact a top-tier system of education.
@@
 With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts.
-In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky & Clelland, 2016).
+In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent [@kanevsky_2013].
 In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students.
@@
 Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students.
-For example, Lo et al.
-(2019) suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+For example, @lo_2019 suggest that the British Columbia Ministry of Education’s recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
 

### 2020-12-07 scripts/Final_project.Rmd
Numerize numbers
-Sooke

@@
   filter(SCHOOL_YEAR >= 9)
+
+ # Making student number column numeric
+x <- c(10)
+BCdata[ , x] <- apply(BCdata[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
 
@@
 
-```{r rural play, echo = FALSE}
-
-##Looking into Sooke (SD 62 - pop. 9,269) categorized under Victoria so urban but technically rural
-## In progress
-
-sooke <- district_class %>%
-  filter(DISTRICT_NAME == "Sooke") %>%
-  as_tibble()
-
-ggplot(sooke, aes(SPECIAL_NEEDS_CODE)) +
-  facet_wrap(~SCHOOL_YEAR)
-  geom_col()
-```
-
 # Discussion

### 2020-12-07 scripts/Final_project.tex
-Sooke

@@
 Educational policy exists as a mechanism to drive changes in school buildings that result in socially important outcomes (e.g., increasing literacy acquisition through improved instruction, reducing instances of bullying through prevention programs).
-However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical (Cohen et al., 2009).
+However, the relationship between policy and practice in Western formal systems of education is often unclear, misdirected, or atheoretical (Cohen, McCabe, Michelli, \& Pickeral, 2009).
 Leveraging the power of policy as a change agent for the delivery and conceptualization of education stands to benefit a significant proportion of students (Swanson, 2007).
 
-When policy is enacted, the intended outcomes are often not realized for months, years, or decades (Gu et al., 2018).
+When policy is enacted, the intended outcomes are often not realized for months, years, or decades ({\textbf{???}}).
 For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores (Luke \& McArdle, 2009).
@@
 The initial analyses will inform a discussion on developmental trajectories by illustrating shifts in prevalence growth from static to linearly increasing based on age of diagnosis for the Oregon data, which can serve as a springboard to make inferences about BC data (e.g., common age of onset for diagnosis per category and diagnostic patterns over time).
-In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM; American Psychological Association, 2013), in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).
+In studying the differences and similarities between OR and BC datasets, we will also engage in a discussion on diagnostic terminology across regions with respect to the definitions as detailed by the Diagnostic and Statistical Manual of Mental Disorders (DSM, Association \& others, 2013), in part as a response to a challenge set by differing terminology between BC and OR categorization (i.e., the United States federal guidelines for exceptionality designation are similar but not identical to exceptionality designation guidelines in BC).
 
-Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (See Lambert, 2017).
+Second, we hope to closely analyze any changes, or lack thereof, within the BC data following Canadian Supreme Court rulings between the British Columbia Teachers Federation and the province of BC (see Lambert, 2017).
 Canadian Supreme Court rulings have significant implications on educational policy in the provinces and, in 2016, resulted in a limit to the number of students with disability designations in classrooms resulting in an expansion of the number of specialist teachers' that school districts are required to hire.
@@
 \includegraphics{Final_project_files/figure-latex/Provinical_Data_Overview-1.pdf}
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:Provincial Data Table}Special Needs Designation by Select Years (BC)}
+
+\begin{tabular}{lllllll}
+\toprule
+Special Needs & \multicolumn{1}{c}{1996/1997} & \multicolumn{1}{c}{2001/2002} & \multicolumn{1}{c}{2005/2006} & \multicolumn{1}{c}{2010/2011} & \multicolumn{1}{c}{2015/2016} & \multicolumn{1}{c}{2019/2020}\\
+\midrule
+Physically Dependent & 2,080.00 & 2,233.00 & 1,953.00 & 1,725.00 & 1,521.00 & 1,636.00\\
+Deafblind & 126.00 & 92.00 & 96.00 & 131.00 & 182.00 & 163.00\\
+Mod To Profound Intellectual Disability & 6,522.00 & 7,512.00 & 7,572.00 & 6,450.00 & 5,382.00 & 4,732.00\\
+Phys Disability Or Chronic Health Impair & 6,722.00 & 14,291.00 & 19,088.00 & 23,896.00 & 30,872.00 & 35,969.00\\
+Visual Impairment & 1,203.00 & 1,136.00 & 1,057.00 & 860.00 & 844.00 & 792.00\\
+Deaf Or Hard Of Hearing & 4,416.00 & 4,652.00 & 4,322.00 & 3,578.00 & 3,505.00 & 3,805.00\\
+Autism Spectrum Disorder & 1,913.00 & 4,487.00 & 7,942.00 & 18,054.00 & 31,423.00 & 52,475.00\\
+Inten Behav Inter/Serious Mental Illness & 19,855.00 & 23,419.00 & 26,970.00 & 26,526.00 & 27,470.00 & 33,156.00\\
+Severe Learning Disabilities & NA & 55,194.00 & NA & NA & NA & NA\\
+Mild Intellectual Disability & NA & 12,701.00 & 8,965.00 & 7,152.00 & 6,966.00 & 8,341.00\\
+Behavior Disorder - Moderate & NA & 26,707.00 & NA & NA & NA & NA\\
+Behavior Disorder - Rehabilitation & NA & 18,726.00 & NA & NA & NA & NA\\
+Gifted & NA & 64,665.00 & 43,997.00 & 27,560.00 & 19,736.00 & 19,717.00\\
+Learning Disability & NA & NA & 62,913.00 & 68,944.00 & 74,709.00 & 84,703.00\\
+Moderate Behaviour Support/Mental Illnes & NA & NA & 34,311.00 & 21,949.00 & 20,396.00 & 24,867.00\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} NAs indicate masked data where students <10}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures x and x below.
+Taken side by side, a simple visualization detailing student exceptionality designation presents the overall similarities between urban and rural areas, as in figures 7 and 8 below.
 Of note, the \enquote{Gifted} designation suffers a severe drop around 2005/2006 for urban and rural areas alike, though the change is more pronounced within rural districts.
@@
 
-\includegraphics{Final_project_files/figure-latex/rural play-1.pdf} \includegraphics{Final_project_files/figure-latex/rural play-2.pdf} \includegraphics{Final_project_files/figure-latex/rural play-3.pdf}
+\includegraphics{Final_project_files/figure-latex/rural play-1.pdf}
 
@@
 With respect to the speculative nature of this assertion, the authors assume that educational policy at the provincial level impacts educational practice across school districts.
-In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky \& Clelland, 2016).
+In 2006, the British Columbia Ministry of Education released Special education services: A manual of policies, procedures and guidelines, which explicitly states the acceleration programming that should be afforded to students with exceptional academic talent (Kanevsky \& Clelland, 2013).
 In fact, this 2006 version of the manual was a revision of the original 1995 manual, and this initial version included scant guidance on service provision to gifted students.
@@
 Given the additional resources needed to properly implement gifted educational services, rural school districts may have de-designated students and sought other ways to provide instruction that meets the needs of their students.
-For example, Lo et al.~(2019) suggest that the British Columbia Ministry of Education's recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
+For example, Lo et al.
+(2019) suggest that the British Columbia Ministry of Education's recent curriculum update makes amends for the traditional test-and-place gifted student support model by giving educators the tools and inclusive framework to offer a rich, meaningful education to all students with exceptionalities.
 
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-american_2013}{}%
+Association, A. P., \& others. (2013). \emph{Diagnostic and statistical manual of mental disorders (dsm-5)}. American Psychiatric Pub.
+
+\leavevmode\hypertarget{ref-ball_1998}{}%
+Ball, S. J. (1998). Big policies/small world: An introduction to international perspectives in education policy. \emph{Comparative Education}, \emph{34}(2), 119--130.
+
+\leavevmode\hypertarget{ref-cohen_2009}{}%
+Cohen, J., McCabe, L., Michelli, N. M., \& Pickeral, T. (2009). School climate: Research, policy, practice, and teacher education. \emph{Teachers College Record}, \emph{111}(1), 180--213.
+
+\leavevmode\hypertarget{ref-kanevsky_2013}{}%
+Kanevsky, L. S., \& Clelland, D. (2013). Accelerating gifted students in canada: Policies and possibilities. \emph{Canadian Journal of Education}, \emph{36}(3), 229--271.
+
+\leavevmode\hypertarget{ref-lambert_2017}{}%
+Lambert, S. (2017). Public education: Reflections on fifteen years of public policy. \emph{BC Studies: The British Columbian Quarterly}, (194), 163--173.
+
+\leavevmode\hypertarget{ref-lo_2019}{}%
+Lo, C.
+O., Tan, Y.
+S.
+M., Chrostowski, M., Hu, S.-F., Chan, D., Sue, D.
+M., \ldots{} Li, W.
+(2019).
+Reenvisioning gifted education in british columbia: A qualitative research protocol of policy analysis in the context of a redesigned curriculum.
+\emph{International Journal of Qualitative Methods}, \emph{18}, 1609406918822233.
+
+\leavevmode\hypertarget{ref-luke_2009}{}%
+Luke, A., \& McArdle, F. (2009). A model for research-based state professional development policy. \emph{Asia-Pacific Journal of Teacher Education}, \emph{37}(3), 231--251.
+
+\leavevmode\hypertarget{ref-michelmore_2017}{}%
+Michelmore, K., \& Dynarski, S. (2017). The gap within the gap: Using longitudinal data to understand income differences in educational outcomes. \emph{AERA Open}, \emph{3}(1), 2332858417692958.
+
+\leavevmode\hypertarget{ref-swanson_2007}{}%
+Swanson, J. D. (2007). Policy and practice: A case study of gifted education policy implementation. \emph{Journal for the Education of the Gifted}, \emph{31}(2), 131--164.
+
+\leavevmode\hypertarget{ref-walker_2013}{}%
+Walker, J., \& Bergmann, H. von. (2013). Teacher education policy in canada: Beyond professionalization and deregulation. \emph{Canadian Journal of Education}, \emph{36}(4), 65--92.
 

### 2020-12-07 scripts/Final_project.Rmd
Please please please work
edited out script, edited citation, changed references

@@
 
-When policy is enacted, the intended outcomes are often not realized for months, years, or decades [@gu-2018].
+When policy is enacted, the intended outcomes are often not realized for months, years, or decades [@gu_2018].
 For example, changing school district policy on professional development to align with research-informed practices may be constructed to increase student academic achievement; though, there is insufficient evidence pointing to how long professionalized intervention with teachers will take to increase student academic scores [@luke_2009].
@@
 cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
-cleanest_primo2
+#cleanest_primo2
 
@@
 
-fresh_primo %>% 
-  group_by(year, disability) %>% 
-  summarize(total=sum(frequency, na.rm=TRUE))
+#fresh_primo %>% 
+  #group_by(year, disability) %>% 
+  #summarize(total=sum(frequency, na.rm=TRUE))
 
@@
 First, the trends of rural student exceptionality designations followed urban trends closely.
-The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (ps > 0.05).
+The vast majority of urban school districts did not differ in their post-2016 prevalence rates for exceptionality designations (*p*s > 0.05).
 However, for diagnostically malleable categories (e.g., Moderate Behavioural designation) in rural school districts, 2016 was an inflection point for an increase in prevalence rates.
-When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (t(239) = -2.113, p = 0.036, g = 0.261, CI[-0.504, -0.019]), Learning Disability (t(262) = -1.984, p = 0.483, g = 0.230, CI[-0.453, -0.007]), and Chronic Health (t(255) = -2.068, p = 0.040, g = 0.247, CI[-0.479, -0.015]).
-Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges’ g > 0.2, and g < 0.5), but did show an increasing trend post-2016 (t(195) = -0.879, p = 0.380, g = 0.120, CI[-0.383, 0.144]).
+When comparing prevalence rates between 2013-2016 (pre-2016 group) and 2017-2019 (post-2016 group) cohorts of rural students, the prevalence rate increased for the following student designation categories: High Behavioural Interventions (*t*(239) = -2.113, *p* = 0.036, *g* = 0.261, CI[-0.504, -0.019]), Learning Disability (*t*(262) = -1.984, *p* = 0.483, *g* = 0.230, CI[-0.453, -0.007]), and Chronic Health (*t*(255) = -2.068, *p* = 0.040, *g* = 0.247, CI[-0.479, -0.015]).
+Note that the exceptionality category Mild Behavioural Intervention did not result with the same small, detectable effect size (Hedges’ *g* > 0.2, and *g* < 0.5), but did show an increasing trend post-2016 (*t*(195) = -0.879, *p* = 0.380, *g* = 0.120, CI[-0.383, 0.144]).
 

### 2020-12-07 scripts/Final_project.Rmd
My turn to try to break everything
Correlogram code edits

@@
 BCdatastripped <- BCdata %>% 
-
     select("SCHOOL_YEAR", 
@@
 
-```{comprehensive display}
+```{correllogram}
 
+#cleaning data
+bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))
+
+district_class <- bc_data_district_classification %>%
+  select("SCHOOL_YEAR",
+         "DATA_LEVEL",
+         "PUBLIC_OR_INDEPENDENT",
+         "DISTRICT_NUMBER",
+         "DISTRICT_NAME",
+         "DISTRICT_POPULATION_2016",
+         "DISTRICT_CLASSIFICATION",
+         "SPECIAL_NEEDS_CODE",
+         "NUMBER_OF_STUDENTS") %>%
+  filter(DATA_LEVEL == "DISTRICT LEVEL",
+         PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
+  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
+    select(-"name")
+
+#Attempting to make columns numeric
+#view(district_class) #Need cols 6 & 8
+
+x <- c(6,8)
+district_class[ , x] <- apply(district_class[ , x], 2,
+                               function(x)as.numeric(as.character(x)))
+#creating data for cor
 g_vic <- district_class %>%

### 2020-12-07 scripts/Final_project.Rmd
Removed Citation confusion and Correllogram

@@
 ```{correllogram}
-
+#we know this doesn't show; couldn't get it to work by submission deadline.
 #cleaning data
@@
 ```
-
-The above correlogram isolates a single school district - Greater Victoria - as a representation of the trends observed within the whole data set.
-This series of plots gives a snapshot introduction for the data trends to be observed.
 
@@
 
-The three datasets were used for analysis in R (cite R) for reproducibility and data visualization.
+The three datasets were used for analysis in R for reproducibility and data visualization.
 All code for analysis and data representation is publicly available for reproducibility (https://github.com/JoeSwinehart/datascienceproject).
@@
 That is, effect sizes between two groups (e.g., group one is three years before a policy change, and group two is three years after a policy change) may be examined to detect the magnitude of differences between groups if they are hypothesized to represent a change in practice stemming from policy implementation.
-More specifically on the calculation of effect sizes, the Hedge’s g statistic will be used because of the robustness against unequal group sizes.
+More specifically on the calculation of effect sizes, the Hedges' g statistic will be used because of the robustness against unequal group sizes.
 

### 2020-12-07 scripts/draft_play.Rmd
Xiaoyu_Peer Review

--- 2020-12-02 scripts/draft_play.Rmd
@@
 ---
-title             : "Trends in Pacific Northwest Special Education Population"
-shorttitle        : "PACNW SpEd Trends"
+title             : "The title"
+shorttitle        : "Title"
 
 author: 
-  - name          : "David Fainstein"
+  - name          : "Makayla Whitney"
     affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "1 David Ave, Eugene, Oregon, USA"
-    email         : "dfainste@uoregon.edu"
+    address       : "Postal address"
+    email         : "my@email.com"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
       - Writing - Review & Editing
-  - name          : "Janette Avelar"
+  - name          : "Ernst-August Doelle"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Makayla Whitney"
+  - name          : "Ernst-August Doelle"
     affiliation   : "1"
@@
       - Writing - Review & Editing
-  - name          : "Joe Swinehart"
+  - name          : "Ernst-August Doelle"
     affiliation   : "1"
@@
 authornote: |
-  Janette Avelar, University of Oregon College of Education
-  David Fainstein, University of Oregon College of Education
-  Joe Swinehart, Behavioral Research & Teaching
-  Makayla Whitney, University of Oregon College of Education
+  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
 
-  Project completed using RStudio, GitKraken, GitHub, and papaja
+  Enter author note here.
 
-  Contact: dfainste@uoregon.edu
 abstract: |
-  This paper will examine the changes in public school special education populations in Vancouver, British Columbia and Oregon over the last 20 years.
-The British Columbia (BC) data was pulled from an open-source data catalogue, published by the education analytics division of the Ministrty of Education in British Columbia.
-This longitudinal data acquistion is part of formal student status reporting that is required within BC public schools - typically occuring in October of each school year.
-The Oregon data was pulled from the Oregon Department of Education and covers special education enrollment from 2002 - 2020.
-This data acquistion is part of formal student status reporting that is required for compliance with federal IDEA laws - typically occuring in November or February of each school year.
-Within the BC dataset, students with exceptionalities are categorized in one of 12 eligibility categories; within the Oregon dataset, this is 13 categories.
+  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
 
@@
 
-keywords          : "Vancouver, British Columbia, Oregon, Special Education, SpEd"
-wordcount         : "0<n<1,000,000"
+keywords          : "keywords"
+wordcount         : "X"
 
@@
 library(janitor)
-library(dplyr)
 
@@
          PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")
+
+# Here I learned to use the package called "readxl", which can help to read xlsx. fless. I will try this and see what is the benefit of this package! 
 ```
@@
 Our project was built around two datasets detailing head counts of students with exceptionalities eligible for special education services aged 6-21.
-The datasets detail the categorization for special education eligibility in public schools within British Columbia (BC) and Oregon.
+The datasets detail the categorization for special education eligibility in public schools within British Columbia and Oregon.
 The head counts from BC are collected from 1996/1997 to the most recent data from 2019/2020.
@@
          NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
-         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
-  filter(DATA_LEVEL == "PROVINCE LEVEL",
-         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
-         SCHOOL_YEAR >= 7,
+         SCHOOL_YEAR = as.integer(SCHOOL_YEAR)) %>% 
+  filter(SCHOOL_YEAR >= 7,
          SPECIAL_NEEDS_CODE != "M",
@@
 
- ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
+# Strengh 4: The reconding part is really smart! 
+ ggplot(BCdatastripped_spedcat, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
   geom_line(color = "cornflowerblue") +
@@
 
- # Making student number column numeric
-x <- c(10)
-BCdata[ , x] <- apply(BCdata[ , x], 2,
-                               function(x)as.numeric(as.character(x)))
-#
-```
-
-```{r table: headcount by year}
-#This is a work in progress no touchy
-
-#Data View
-#view(BCdatastripped)
-#view(BCdata)
-#as_tibble(count_by_year)
-
-#Data Prep
-count_by_year <- BCdata %>% 
-  janitor::clean_names() %>% 
-  group_by(school_year, special_needs_desc) %>% 
-  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select("school_year",
-         "special_needs_desc",
-         "total_students") %>% 
-  drop_na()
-
-#Table
-#apa_table(
-#  count_by_year,
-#  digits = 0,
-#  caption = "Headcount by Classification and District, Select Years 1996-2020",
-#  note = "Number of students is estimated and does not include masked data.",
-#  col.names = c("School Year",
-#                "Special Needs Description",
-#                "Number of Students")
-#)
-
-#Table is huuuge - do not run table code yet or your computer will be frozen for 5 minutes.
-
-unique(count_by_year$school_year)
+ # Strength 1: I would like to comment on their writing, it is very clear and straightfoward! 
 ```
@@
 
-#Attempting to make columns numeric
-view(district_class) #Need cols 6 & 8
-
-x <- c(6,8)
-district_class[ , x] <- apply(district_class[ , x], 2,
-                               function(x)as.numeric(as.character(x)))
-#
+# strength 2: I really like the data cleaning part. They used pivot_longer to cluster the district classfication. This makes the dataset more clear and clean! 
 ```
@@
 ``` {r District Census Population Table}
-#Data Prep
-census_data <- district_class %>% 
-  janitor::clean_names() %>% 
-  group_by(district_name, district_number) %>% 
-  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c("school_year",
-            "data_level",
-            "public_or_independent",
-            "special_needs_code",
-            "number_of_students")) %>% 
-  distinct()
 
-#Table
-apa_table(
-  census_data,
-    digits = 0,
-  caption = "District Classification by Population According to 2016 Census",
-  note = "Number of students is estimated and does not include masked data.",
-  col.names = c("District Number",
-            "District Name",
-            "Population",
-            "Classification",
-            "Number of Students")
-)
 ```
@@
 
-```{r rural play}
-#Data Prep
-rural_districts <- district_class %>% 
-  janitor::clean_names() %>% 
-  filter(district_classification == "Rural") %>% 
-  select(-c("data_level",
-            "public_or_independent",
-            "district_population_2016"))
-
-#Viz - Totals by Code
-ggplot(rural_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
-       x = "Exceptionality",
-       y = "Number of Students")
-
-#Data for Rural Totals (No categories)
-rural_totals <- rural_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#Data Viz Totals/Time
-ggplot(rural_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-#What's going on with Prince George? Why is there such a steep drop in 2006?
-pg <- rural_districts %>% 
-  filter(district_name == "Prince George") %>% 
-  as_tibble()
-
-pg_sums <- pg %>% 
-  group_by(school_year) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#What's happening by Dx?
-pg_dx_sum <- pg %>% 
-  group_by(school_year, special_needs_code) %>% 
-  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(district_number, district_classification, number_of_students)) %>% 
-  distinct()
-
-ggplot(pg_dx_sum, aes(school_year, student)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  facet_wrap(~special_needs_code)
-
-#What's going on with gifted?????
-pg_gifted <- pg %>% 
-  filter(special_needs_code == "P") %>% 
-  select(-c(district_number, district_classification))
-
-ggplot(pg_gifted, aes(school_year, number_of_students)) +
-  geom_col() +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-##Looking into Sooke (SD 62 - pop. 9,269) categorized under Victoria so urban but technically rural
-## In progress
-
-sooke <- district_class %>%
-  filter(DISTRICT_NAME == "Sooke") %>%
-  as_tibble()
-
-ggplot(sooke, aes(SPECIAL_NEEDS_CODE)) +
-  facet_wrap(~SCHOOL_YEAR)
-  geom_col()
-```
-
-```{r urban play}
-#Data Prep
-urban_districts <- district_class %>% 
-  janitor::clean_names() %>% 
-  filter(district_classification == "Urban") %>% 
-  select(-c("data_level",
-            "public_or_independent",
-            "district_population_2016"))
-
-urban_totals <- urban_districts %>% 
-  group_by(school_year, district_name, district_number) %>% 
-  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
-  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
-  distinct()
-
-#Data Viz Totals/Time
-ggplot(urban_totals, aes(school_year, students)) +
-  geom_col(aes(color = district_name)) +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-#Viz - Totals by Code
-ggplot(urban_districts, 
-       aes(special_needs_code, number_of_students,
-           color = special_needs_code)) +
-  geom_line() +
-  facet_wrap(~school_year) +
-  labs(title = "Total Students by Exceptionality",
-       x = "Exceptionality",
-       y = "Number of Students")
-```
-
 ```{r cleaning Oregon data}
@@
          year = parse_number(year))
-
+# Strength 4: This data cleanig is really cool and smart! 
 fresh_primo %>% 
@@
  facet_wrap(~disability)+
-    labs(color="frequency", x= "Year", y="Student Count", title = "Oregon Headcount of Students with Exceptionalities by Category")
+    labs(color="frequency", x= "Year", y="Student Count", title = "Trends in Frequency")
 
+#improvement: For better describing the data, maybe they could considering adding one or two descriptive tables with some descriptive statistics. 
 ```
-
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Participants
-
-## Material
 
